---
title: "Revision Analysis: `r params$signal_name`"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    max-width: 1280px
params:
  # Signal to analyze
  source: 
  signal: 
  signal_name: 
  # Optional path to a local CSV or Parquet directory with a version/issue column.
  # When NULL, the signal is fetched from the Delphi API with issues = "*".
  input_dir: !expr "NULL"
  # Shared settings
  geo_type:  
  time_type: 
  start_day: 
  end_day:   
  # Maximum locations shown in faceted plots / summary tables.
  max_locations_plot: 6
  max_locations_table: 15
  # Number of worst-behaving locations to highlight.
  n_worst: 18
  # Convergence threshold: how close (relative) a report must be to the
  # latest Report value to be considered "converged".
  convergence_threshold: 0.05
  # Minimum revision history required for reference dates used in diagnostics.
  # Reference dates after render_date - min_revision_history_days are excluded.
  min_revision_history_days: 365
knitr:
  opts_chunk:
    comment: "#>"
    message: FALSE
    warning: FALSE
    out.width: "100%"
    dev: "png"
    fig.width: 9
    fig.height: 5
    fig.align: center
editor_options:
  chunk_output_type: console
execute:
  cache: false
---

```{r}
#| label: setup
#| include: false
#| cache: false
# When running interactively, fall back to the parameter defaults.
if (!exists("params")) {
    params <- list(
        source                = "hospital-admissions",
        signal                = "smoothed_covid19_from_claims",
        signal_name           = "Hospital Admissions: Smoothed COVID-19 from Claims",
        input_dir             = "revision_analysis/data/revisions/",
        geo_type              = "state",
        time_type             = "day",
        start_day             = "2023-01-01",
        end_day               = "2023-02-01",
        max_locations_plot    = 6,
        max_locations_table   = 15,
        n_worst               = 18,
        convergence_threshold = 0.05,
        min_revision_history_days = 365
    )
}

options(repos = c(
    delphi = "https://cmu-delphi.r-universe.dev",
    CRAN = "https://cloud.r-project.org"
))
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    epidatr, epiprocess, covidcast,
    tidyverse, scales, patchwork,
    knitr, here, MMWRweek, glue, DT,
    fs, arrow, data.table
)

# Invalidate knitr cache when params or local input files change.
input_cache_key <- NULL
if (!is.null(params$input_dir) && dir.exists(params$input_dir)) {
    input_files <- list.files(params$input_dir, full.names = TRUE, recursive = TRUE)
    input_cache_key <- list(
        n_files = length(input_files),
        latest_mtime = if (length(input_files) > 0) max(file.info(input_files)$mtime, na.rm = TRUE) else NA
    )
}
render_date <- Sys.Date()
knitr::opts_chunk$set(cache.extra = list(params = params, input = input_cache_key, render_date = render_date))

colours <- c(
    "#172431",
    "#ffffff",
    "#C41230",
    "#941120",
    "#007BC0",
    "#002676",
    "#FDB515",
    "#002145",
    "#00A7E1",
    "#8C338F",
    "#368f33"
)

dis_pal <- colours[c(3, 5, 7, 10, 11, 9)]

theme_set(
    theme_minimal(base_family = "sans", base_size = 13) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, color = "gray45", size = 12),
            plot.caption = element_text(hjust = 0, color = "gray50", size = 11),
            axis.title = element_text(size = 12, face = "bold"),
            axis.text = element_text(size = 11),
            legend.title = element_text(size = 11, face = "bold"),
            legend.text = element_text(size = 11),
            strip.text = element_text(size = 12, face = "bold"),
            panel.grid = element_line(linetype = 2, linewidth = 0.3, color = "gray90"),
            panel.grid.minor = element_blank(),
            legend.position = "top",
            strip.background = element_rect(fill = "gray95", linetype = "blank"),
            panel.border = element_rect(color = "gray95", fill = NA)
        )
)

# Palette as global defaults for discrete and continuous scales
options(
    ggplot2.discrete.colour   = dis_pal,
    ggplot2.discrete.fill     = dis_pal,
    ggplot2.continuous.colour = function() scale_colour_gradientn(colours = dis_pal),
    ggplot2.continuous.fill   = function() scale_fill_gradientn(colours = dis_pal)
)

tibble::tribble(
    ~name,      ~aes,     ~alpha, ~is_stat,
    "line",     "colour", 1.0,    FALSE,
    "point",    "colour", 0.7,    FALSE,
    "function", "colour", 1.0,    FALSE,
    "bar",      "fill",   0.9,    FALSE,
    "col",      "fill",   0.9,    FALSE,
    "boxplot",  "fill",   0.8,    FALSE,
    "bin",      "fill",   0.8,    TRUE,
    "identity", "fill",   0.7,    TRUE
) %>%
    purrr::pwalk(function(name, aes, alpha, is_stat) {
        updater <- if (is_stat) update_stat_defaults else update_geom_defaults
        params <- list(dis_pal[1], alpha)
        names(params) <- c(aes, "alpha")
        if (name == "boxplot") params$outlier.shape <- NA
        updater(name, params)
    })

# Helper: build epirange with correct format for the signal's time_type
time_range <- function(start, end, time_type = params$time_type) {
    if (time_type == "week") {
        to_yw <- function(d) {
            d <- as.Date(d)
            m <- MMWRweek::MMWRweek(d)
            m$MMWRyear * 100L + m$MMWRweek
        }
        epirange(to_yw(start), to_yw(end))
    } else {
        epirange(start, end)
    }
}

# Set cache
cache_path <- file.path(getwd(), "epidatr_cache")
if (!dir.exists(cache_path)) dir.create(cache_path, recursive = TRUE)

set_cache(cache_dir = cache_path, days = Inf)

# Helper: format a difftime object as a plain number of days
as_days <- function(x) as.numeric(x, units = "days")
```

This notebook analyses the **revision behaviour** of a single time series stored
as an `epi_archive`. The central question is:

> At what lag does a series become reliable enough to use?

::: {.callout-important icon=false}
## Notation and terminology {.unnumbered}

- **Reference date**: The date when the event occurred. For data fetched from the Delphi Epidata API, this is the `time_value` column.

- **Report date**: The date attached to a published or released Report value for a given reference date. For data fetched from the Delphi Epidata API, this is the `issue_date` column.

- **Lag**: The elapsed time between the reference date and the report date.

- **Report**: One released record for a specific location, reference date, and report date. It contains the Report value available for that location and reference date as of that report date.

- **Latest report**: For a given location and reference date retained in the diagnostics, the most recent report available in the input data. In this render, diagnostic reference dates must have at least `r params$min_revision_history_days` days of possible revision history, so reference dates after `r render_date - params$min_revision_history_days` are excluded before the latest report is selected.

- **Epi season**: A cross-year season used to align respiratory-virus timing. In this notebook, each epi season runs from MMWR week 40 through MMWR week 39 of the following year.

**Caution**: The report date used in this notebook is determined entirely by the input data. For data fetched from the Delphi Epidata API, the report date is the provided `issue_date`. This date may reflect a publication rule, model run, file-generation process, or API version tag, rather than the first date when the underlying raw data became available.

If you upload a CSV file, confirm that the report date has the intended meaning for your indicator before interpreting minimum lag, reporting latency, or stabilization time.
:::

---

```{r}
#| label: load-archive
#| include: false
archive <- NULL

if (!is.null(params$input_dir)) {
    path <- here::here(params$input_dir)

    if (fs::is_dir(path)) {
        cli::cli_inform("Reading data from directory: {.path {params$input_dir}}")
        raw <- arrow::open_dataset(path) %>%
            select(any_of(c("geo_value", "time_value", "version", "issue", "value"))) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            ) %>%
            collect()
    } else if (grepl("\\.parquet$", path)) {
        cli::cli_inform("Reading Parquet file: {.path {params$input_dir}}")
        raw <- arrow::read_parquet(path)
    } else {
        cli::cli_inform("Reading CSV file: {.path {params$input_dir}}")
        raw <- read_csv(path)
    }

    raw <- raw %>% mutate(time_value = as.Date(time_value))

    # Accept either 'version' or 'issue' as the version column.
    if ("issue" %in% names(raw) && !"version" %in% names(raw)) {
        raw <- rename(raw, version = issue)
    }

    if (!"version" %in% names(raw)) {
        cli::cli_abort(
            "The input data must contain a 'version' or 'issue' column. ",
            "See the README for the required schema."
        )
    }

    # If it was a single file (CSV/Parquet), we still need to filter/mutate metadata
    if (!fs::is_dir(path)) {
        raw <- raw %>%
            mutate(version = as.Date(version)) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            )
    }

    archive <- raw %>%
        select(geo_value, time_value, version, value) %>%
        as_epi_archive()

    msg_verb <- "Loaded"
} else {
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE

    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                params$signal %in% meta_res[[params$source]]$signals
            },
            error = function(e) {
                FALSE
            }
        )
        if (meta_available) {
            use_new_api <- TRUE
        }
    }

    if (use_new_api) {
        df_raw <- epidata_archive(
            source = params$source,
            signals = params$signal,
            geo_type = params$geo_type
        ) %>%
            rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            )
    } else {
        start_date <- as.Date(params$start_day)
        end_date <- as.Date(params$end_day)
        interval_starts <- seq(start_date, end_date, by = "month")
        interval_ends <- pmin(
            lubridate::ceiling_date(interval_starts, "month") - lubridate::days(1),
            end_date
        )

        df_raw <- purrr::map2_dfr(interval_starts, interval_ends, function(s, e) {
            cli::cli_inform(c(">" = "Fetching API revisions for {s} to {e}"))

            pub_covidcast(
                source = params$source,
                signal = params$signal,
                geo_type = params$geo_type,
                time_type = params$time_type,
                time_values = time_range(s, e),
                issues = "*"
            )
        }) %>%
            distinct()
    }

    if ("issue" %in% colnames(df_raw) && !"version" %in% colnames(df_raw)) {
        df_raw <- rename(df_raw, version = issue)
    }

    if (nrow(df_raw) == 0 || !"version" %in% colnames(df_raw)) {
        cli::cli_abort(
            "No versioned data returned from the API. ",
            "The signal may not support revision history."
        )
    }

    archive <- df_raw %>%
        select(geo_value, time_value, version, value) %>%
        as_epi_archive()

    msg_verb <- "Fetched"
}

n_rows <- nrow(archive$DT)
n_locations <- n_distinct(archive$DT$geo_value)
n_versions <- n_distinct(archive$DT$version)
reference_start <- min(archive$DT$time_value)
reference_end <- max(archive$DT$time_value)
version_start <- min(archive$DT$version)
version_end <- max(archive$DT$version)

cli::cli_inform(c("i" = "{msg_verb} {label_comma()(n_rows)} rows | {label_comma()(n_locations)} locations | {label_comma()(n_versions)} versions ({version_start} – {version_end})"))

invisible(gc())
```

```{r}
#| label: print-config
#| echo: false
config_tbl <- tibble(
    Field = c(
        "Source",
        "Signal",
        "Label",
        "Geo resolution",
        "Temporal resolution (reference date)",
        "Reference date range",
        "Report date range",
        "Minimum revision history",
        "Number of locations",
        "Number of reports"
    ),
    Value = c(
        params$source,
        params$signal,
        params$signal_name,
        params$geo_type,
        params$time_type,
        paste(reference_start, "to", reference_end),
        paste(version_start, "to", version_end),
        paste(params$min_revision_history_days, "days"),
        label_comma()(n_locations),
        label_comma()(n_rows)
    )
)

knitr::kable(config_tbl, caption = "Analysis configuration")
```

```{r}
#| label: quarto-metadata
#| results: asis
#| echo: false
# Dynamically set Quarto metadata so we can use it in ::: {.content-visible}
# blocks later.
n_years <- n_distinct(year(archive$DT$time_value))
cat(sprintf(
    "\n---\nmulti_location: %s\nmulti_year: %s\n---\n",
    tolower(n_locations > 1),
    tolower(n_years > 1)
))
```

```{r}
#| label: archive-meta
#| include: false
archive_span <- as.numeric(version_end - version_start, units = "days")
has_history <- archive_span >= 60
all_geos <- sort(unique(archive$DT$geo_value))
stable_reference_cutoff <- render_date - params$min_revision_history_days
max_reference_date <- max(archive$DT$time_value, na.rm = TRUE)
has_recent_reference_dates <- max_reference_date > stable_reference_cutoff
diagnostic_archive <- archive$DT %>%
    filter(time_value <= stable_reference_cutoff) %>%
    as_epi_archive()
```

```{r}
#| label: archive-guard-message
#| echo: false
#| results: asis
if (!has_history) {
    cat(sprintf(
        "> **Warning**: only %d days of report history are available ",
        archive_span
    ))
    cat(sprintf(
        "(%s – %s). At least 60 days are needed for a meaningful ",
        version_start, version_end
    ))
    cat(sprintf(
        "revision analysis. The earliest this will be possible is %s.\n\n",
        version_start + 60
    ))
}
```

---

```{r}
#| label: revision-analysis
#| include: false
#| cache: true
rev_result <- epiprocess::revision_analysis(
    diagnostic_archive,
    value,
    min_waiting_period = 0
)
rev_beh <- rev_result$revision_behavior

# Track which (geo_value, time_value) pairs were included vs. excluded.
# revision_analysis() drops pairs that have no version past the waiting
# period — typically the most recent time values with only an initial report.
all_pairs <- diagnostic_archive$DT %>%
    transmute(geo_value = tolower(geo_value), time_value) %>%
    distinct()
analyzed_pairs <- rev_beh %>%
    transmute(geo_value = tolower(geo_value), time_value) %>%
    distinct()
n_excluded <- nrow(anti_join(all_pairs, analyzed_pairs, by = c("geo_value", "time_value")))
n_analyzed <- nrow(analyzed_pairs)

# Global metrics for follow-on plots/narrative
max_lag_days <- as_days(rev_beh$max_lag)
no_rev_pct <- mean(rev_beh$n_revisions == 0, na.rm = TRUE)
conv_days_all <- as_days(rev_beh$lag_near_latest)
invisible(gc())
```

```{r}
#| label: convergence-compute
#| include: false
#| cache: true
diagnostic_report_df <- diagnostic_archive$DT %>%
    transmute(
        geo_value = tolower(geo_value),
        time_value,
        version,
        value,
        publication_lag = as.numeric(version - time_value, units = "days")
    ) %>%
    filter(publication_lag >= 0, !is.na(value))

analyzed_report_df <- diagnostic_report_df %>%
    semi_join(analyzed_pairs, by = c("geo_value", "time_value"))

pair_summary <- analyzed_report_df %>%
    group_by(geo_value, time_value) %>%
    summarize(
        first_report_date = min(version, na.rm = TRUE),
        latest_report_date = max(version, na.rm = TRUE),
        first_report_value = value[which.min(version)],
        latest_report_value = value[which.max(version)],
        min_report_value = min(value, na.rm = TRUE),
        max_report_value = max(value, na.rm = TRUE),
        initial_lag = min(publication_lag, na.rm = TRUE),
        n_reports = n_distinct(version),
        n_values = n_distinct(value),
        .groups = "drop"
    ) %>%
    filter(latest_report_value != 0, !is.na(latest_report_value), !is.na(first_report_value)) %>%
    mutate(
        initial_bias = (first_report_value - latest_report_value) / abs(latest_report_value)
    )

convergence_df <- analyzed_report_df %>%
    inner_join(
        pair_summary %>% select(geo_value, time_value, value_latest = latest_report_value),
        by = c("geo_value", "time_value")
    ) %>%
    mutate(abs_rel_diff = abs(value - value_latest) / abs(value_latest))
invisible(gc())
```

```{r}
#| label: shared-plot-helpers
#| echo: false
focus_days <- 90
tail_start <- 500
early_day_breaks <- c(0, 7, 14, 30, 60, 90)
threshold_levels <- c("50%", "20%", "10%", "5%", "1%", "0.5%", "0%")
threshold_cols <- c(
    "50%" = "#dc2626",
    "20%" = "#f59e0b",
    "10%" = "#facc15",
    "5%" = "#16a34a",
    "1%" = "#2563eb",
    "0.5%" = "#7c3aed",
    "0%" = "#92400e"
)
band_cols <- c(
    "50% to 100%" = "gray82",
    "50% to 20%" = unname(threshold_cols["50%"]),
    "20% to 10%" = unname(threshold_cols["20%"]),
    "10% to 5%" = unname(threshold_cols["10%"]),
    "5% to 1%" = unname(threshold_cols["5%"]),
    "1% to 0.5%" = unname(threshold_cols["1%"]),
    "Within 0.5%" = unname(threshold_cols["0.5%"])
)
final_revision_fill <- unname(threshold_cols["0%"])

focus_day_breaks <- function(limits, endpoint = focus_days) {
    if (limits[1] <= 0 && limits[2] <= endpoint) {
        return(early_day_breaks[early_day_breaks <= endpoint])
    }
    breaks_pretty()(limits)
}
days_since_breaks <- function(limits) {
    candidate_breaks <- sort(unique(c(breaks_pretty()(limits), early_day_breaks)))
    candidate_breaks[candidate_breaks >= limits[1] & candidate_breaks <= limits[2]]
}

epi_season_label <- function(date) {
    date <- as.Date(date)
    mmwr_date <- MMWRweek(date)
    start_year <- if_else(mmwr_date$MMWRweek >= 40, mmwr_date$MMWRyear, mmwr_date$MMWRyear - 1L)
    paste0(start_year, "-", start_year + 1L)
}

season_levels <- convergence_df %>%
    mutate(epi_season = epi_season_label(time_value)) %>%
    distinct(epi_season) %>%
    arrange(epi_season) %>%
    pull(epi_season)

convergence_season_df <- convergence_df %>%
    mutate(epi_season = factor(epi_season_label(time_value), levels = season_levels))

plot_to_data_uri <- function(plot, width = 12, height = 8, dpi = 120) {
    out_file <- tempfile(fileext = ".png")
    ggsave(out_file, plot = plot, width = width, height = height, dpi = dpi, bg = "white")
    knitr::image_uri(out_file)
}

hhs_region_map <- tribble(
    ~geo_value, ~hhs_region,
    "ct", "HHS 1", "me", "HHS 1", "ma", "HHS 1", "nh", "HHS 1", "ri", "HHS 1", "vt", "HHS 1",
    "nj", "HHS 2", "ny", "HHS 2",
    "de", "HHS 3", "dc", "HHS 3", "md", "HHS 3", "pa", "HHS 3", "va", "HHS 3", "wv", "HHS 3",
    "al", "HHS 4", "fl", "HHS 4", "ga", "HHS 4", "ky", "HHS 4", "ms", "HHS 4", "nc", "HHS 4", "sc", "HHS 4", "tn", "HHS 4",
    "il", "HHS 5", "in", "HHS 5", "mi", "HHS 5", "mn", "HHS 5", "oh", "HHS 5", "wi", "HHS 5",
    "ar", "HHS 6", "la", "HHS 6", "nm", "HHS 6", "ok", "HHS 6", "tx", "HHS 6",
    "ia", "HHS 7", "ks", "HHS 7", "mo", "HHS 7", "ne", "HHS 7",
    "co", "HHS 8", "mt", "HHS 8", "nd", "HHS 8", "sd", "HHS 8", "ut", "HHS 8", "wy", "HHS 8",
    "az", "HHS 9", "ca", "HHS 9", "hi", "HHS 9", "nv", "HHS 9",
    "ak", "HHS 10", "id", "HHS 10", "or", "HHS 10", "wa", "HHS 10"
)
```

## Results at a glance {#overall .unnumbered}

The time-series plot below compares the Report value available at selected lags
using all reference dates in the analysis period.

```{r}
#| label: report-values-by-lag
#| echo: false
#| results: asis
#| cache: true
report_lags <- c(0, 7, 14, 30, 60, 90, 120, 180)
report_lag_levels <- c(paste0(report_lags, "d"), "Latest")

report_archive_df <- archive$DT %>%
    transmute(
        geo_value = tolower(geo_value),
        time_value,
        version,
        value,
        publication_lag = as.numeric(version - time_value, units = "days")
    ) %>%
    filter(publication_lag >= 0, !is.na(value))

report_archive_dt <- as.data.table(report_archive_df)
setorder(report_archive_dt, geo_value, time_value, publication_lag, version)

as_of_lag_df <- report_archive_dt[
    ,
    {
        lag_index <- findInterval(report_lags, publication_lag)
        lag_keep <- lag_index > 0

        if (!any(lag_keep)) {
            .SD[0][, report_lag := character()]
        } else {
            .SD[lag_index[lag_keep]][
                ,
                report_lag := paste0(report_lags[lag_keep], "d")
            ]
        }
    },
    by = .(geo_value, time_value)
] %>%
    as_tibble()

latest_report_df <- report_archive_dt[
    ,
    .SD[.N],
    by = .(geo_value, time_value)
] %>%
    as_tibble() %>%
    mutate(report_lag = "Latest")

report_lag_timeseries <- bind_rows(as_of_lag_df, latest_report_df) %>%
    mutate(report_lag = factor(report_lag, levels = report_lag_levels)) %>%
    group_by(geo_value, time_value, report_lag) %>%
    summarize(
        report_value = mean(value, na.rm = TRUE),
        .groups = "drop"
    )

report_lag_cols <- setNames(
    c(hue_pal()(length(report_lags)), "#111827"),
    report_lag_levels
)
report_lag_date_end <- max(report_lag_timeseries$time_value, na.rm = TRUE)

report_lag_subtitle_base <- paste(
    "Each line shows the selected state's Report value using reports available as of the selected lag.",
    "This view uses all reference dates in the analysis period.",
    sep = "\n"
)

plot_report_lag_state <- function(state_value) {
    state_data <- report_lag_timeseries %>% filter(geo_value == state_value)
    report_lag_subtitle <- report_lag_subtitle_base

    report_lag_plot <- ggplot(state_data, aes(x = time_value, y = report_value, color = report_lag))

    if (has_recent_reference_dates) {
        report_lag_subtitle <- paste(
            report_lag_subtitle,
            "Shaded dates are excluded from subsequent diagnostics.",
            sep = "\n"
        )
        report_lag_plot <- report_lag_plot +
            annotate(
                "rect",
                xmin = stable_reference_cutoff,
                xmax = report_lag_date_end,
                ymin = -Inf,
                ymax = Inf,
                fill = "gray70",
                alpha = 0.18
            ) +
            geom_vline(xintercept = stable_reference_cutoff, color = "gray35", linetype = "dashed", linewidth = 0.7) +
            annotate(
                "text",
                x = stable_reference_cutoff,
                y = Inf,
                label = "\u2702\ufe0e",
                angle = 0,
                hjust = 0.5,
                vjust = 1.45,
                size = 7,
                color = "gray25",
                family = "Apple Symbols"
            )
    }

    report_lag_plot +
        geom_line(linewidth = 0.85, alpha = 0.95, na.rm = TRUE) +
        scale_color_manual(
            values = report_lag_cols,
            breaks = levels(droplevels(state_data$report_lag)),
            drop = TRUE
        ) +
        scale_x_date(
            breaks = function(limits) {
                candidate_breaks <- breaks_pretty()(limits)
                if (has_recent_reference_dates &&
                    stable_reference_cutoff >= limits[1] &&
                    stable_reference_cutoff <= limits[2]) {
                    candidate_breaks <- sort(unique(c(candidate_breaks, stable_reference_cutoff)))
                }
                candidate_breaks
            }
        ) +
        scale_y_continuous(labels = label_comma(), breaks = breaks_pretty()) +
        labs(
            title = sprintf("Report values by lag: %s", toupper(state_value)),
            subtitle = report_lag_subtitle,
            x = "Reference date",
            y = "Report value",
            color = "Lag"
        ) +
        theme(
            legend.title = element_text(size = 22, face = "bold"),
            legend.text = element_text(size = 20),
            legend.key.width = grid::unit(2.2, "cm"),
            legend.spacing.x = grid::unit(0.45, "cm"),
            legend.box.spacing = grid::unit(0.35, "cm"),
            plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
            plot.subtitle = element_text(size = 21, hjust = 0.5, lineheight = 1.1),
            axis.title = element_text(size = 24, face = "bold"),
            axis.text = element_text(size = 20)
        )
}

report_lag_states <- sort(unique(report_lag_timeseries$geo_value))
default_report_lag_state <- if ("pa" %in% report_lag_states) "pa" else report_lag_states[1]
report_lag_panel_html <- map_chr(report_lag_states, function(state_value) {
    image_uri <- plot_to_data_uri(
        plot_report_lag_state(state_value),
        width = 18,
        height = 8.2
    )
    state_label <- toupper(state_value)
    display <- ifelse(state_value == default_report_lag_state, "block", "none")
    sprintf(
        '<div class="report-lag-state-panel" data-state="%s" style="display:%s;"><img src="%s" style="width:100%%;" alt="Report values by lag for %s"></div>',
        state_label, display, image_uri, state_label
    )
})

report_lag_buttons_html <- paste0(
    paste0(
        '<button type="button" class="report-lag-state-button',
        ifelse(report_lag_states == default_report_lag_state, " active", ""),
        '" data-state="',
        toupper(report_lag_states),
        '">',
        toupper(report_lag_states),
        '</button>',
        collapse = ""
    )
)

report_lag_control_html <- paste0(
    '<div class="report-lag-state-control">',
    '<div class="report-lag-state-buttons">',
    report_lag_buttons_html,
    '</div>',
    '<span class="report-lag-state-status" aria-live="polite">Showing ',
    toupper(default_report_lag_state),
    '</span>',
    '</div>'
)

report_lag_script_html <- paste(
    '<script>',
    '(function() {',
    '  const widget = document.currentScript ? document.currentScript.closest(".report-lag-state-widget") : document.querySelector(".report-lag-state-widget");',
    '  if (!widget) return;',
    '  const buttons = widget.querySelectorAll(".report-lag-state-button");',
    '  const panels = widget.querySelectorAll(".report-lag-state-panel");',
    '  const status = widget.querySelector(".report-lag-state-status");',
    '  if (!buttons.length || !panels.length) return;',
    '  function showState(state) {',
    '    panels.forEach(function(panel) {',
    '      panel.style.display = panel.dataset.state === state ? "block" : "none";',
    '    });',
    '    buttons.forEach(function(button) {',
    '      const isActive = button.dataset.state === state;',
    '      button.classList.toggle("active", isActive);',
    '      button.setAttribute("aria-current", isActive ? "true" : "false");',
    '    });',
    '    if (status) status.textContent = "Showing " + state;',
    '  }',
    '  buttons.forEach(function(button) {',
    '    button.addEventListener("click", function() { showState(button.dataset.state); });',
    '  });',
    '  const activeButton = widget.querySelector(".report-lag-state-button.active") || buttons[0];',
    '  showState(activeButton.dataset.state);',
    '})();',
    '</script>',
    sep = "\n"
)

report_lag_style_html <- paste(
    '<style>',
    '.report-lag-state-control {',
    '  display: flex;',
    '  align-items: center;',
    '  flex-wrap: wrap;',
    '  gap: 0.5rem;',
    '  margin: 0.75rem 0 1rem 0;',
    '}',
    '.report-lag-state-buttons {',
    '  display: grid;',
    '  grid-template-columns: repeat(10, minmax(2.5rem, 1fr));',
    '  gap: 0.25rem;',
    '  width: min(100%, 34rem);',
    '}',
    '.report-lag-state-button {',
    '  border: 1px solid #d0d7de;',
    '  background: #fff;',
    '  color: #24292f;',
    '  border-radius: 4px;',
    '  padding: 0.28rem 0.2rem;',
    '  font-size: 0.85rem;',
    '  font-weight: 700;',
    '  line-height: 1.2;',
    '  cursor: pointer;',
    '}',
    '.report-lag-state-button.active {',
    '  background: #0969da;',
    '  border-color: #0969da;',
    '  color: #fff;',
    '}',
    '.report-lag-state-status {',
    '  color: #57606a;',
    '  margin-left: 0.35rem;',
    '  font-size: 0.9rem;',
    '}',
    '</style>',
    sep = "\n"
)

cat(paste(
    '<div class="report-lag-state-widget">',
    report_lag_control_html,
    paste(report_lag_panel_html, collapse = "\n"),
    report_lag_script_html,
    report_lag_style_html,
    '</div>',
    sep = "\n"
))
```

```{r}
#| label: diagnostic-cutoff-note
#| echo: false
#| results: asis
if (has_recent_reference_dates) {
    cat(sprintf(
        "For the diagnostics below, reference dates after **%s** are excluded because they have fewer than **%d days** of possible revision history before the render date (**%s**). This helps the latest-report comparison reflect a more stable revision pattern.\n\n",
        stable_reference_cutoff,
        params$min_revision_history_days,
        render_date
    ))
}
```

The next diagnostic summarizes revision size using Relative Error. For each
current report, Relative Error compares the Report value with the latest Report value available for
the same location and reference date:

$$
\text{Relative Error} = \frac{|r_c - r_l|}{|r_l|}
$$

Here `r_c` is the Report value in the current report and `r_l` is the Report
value in the latest report for the same location and reference date.

A Relative Error of 0% means the current report matches the latest report; a Relative Error of 10%
means it is still 10% away from the latest report.

This heatmap shows the Relative Error surface over daily reference date and lag.
Each cell summarizes the mean Relative Error across all states for reports with
the same reference date and lag. Epi seasons are shown one page at a time, with the
most recent epi season first. White cells mean that no report was available for that
reference-date and lag combination, so they should be read as missing data rather
than zero Relative Error.

```{r}
#| label: relative-error-heatmap
#| echo: false
#| results: asis
#| cache: true
season_start_date <- function(season_label) {
    start_year <- as.integer(str_extract(as.character(season_label), "^\\d{4}"))
    as.Date(sprintf("%d-10-01", start_year))
}

season_month_starts <- seq(as.Date("2000-10-01"), by = "month", length.out = 12)
season_month_breaks <- as.numeric(season_month_starts - season_month_starts[1])
season_month_labels <- format(season_month_starts, "%m")
season_axis_end <- as.numeric(as.Date("2001-10-01") - as.Date("2000-10-01"))
heatmap_focus_days <- 180
heatmap_day_breaks <- c(0, 7, 14, 30, 60, 90, 120, 150, 180)

heatmap_df <- convergence_season_df %>%
    filter(publication_lag >= 0, publication_lag <= heatmap_focus_days) %>%
    filter(!is.na(epi_season)) %>%
    mutate(
        season_start = season_start_date(epi_season),
        season_day = as.numeric(time_value - season_start)
    ) %>%
    filter(season_day >= 0, season_day < season_axis_end) %>%
    group_by(epi_season, season_day, publication_lag) %>%
    summarize(
        mean_relative_error = mean(abs_rel_diff, na.rm = TRUE),
        n_reports = n(),
        .groups = "drop"
    )

relative_error_step_cols <- c(
    alpha(unname(threshold_cols["0.5%"]), 0.52),
    alpha(unname(threshold_cols["1%"]), 0.52),
    alpha(unname(threshold_cols["5%"]), 0.52),
    alpha(unname(threshold_cols["10%"]), 0.52),
    alpha(unname(threshold_cols["20%"]), 0.52),
    alpha(unname(threshold_cols["50%"]), 0.52),
    alpha("gray82", 0.52)
)

relative_error_step_breaks <- c(0.005, 0.01, 0.05, 0.1, 0.2, 0.5)
relative_error_step_midpoints <- c(0.0025, 0.0075, 0.03, 0.075, 0.15, 0.35, 0.75)

plot_relative_error_heatmap <- function(season_value) {
    season_data <- heatmap_df %>% filter(epi_season == season_value)

    ggplot(season_data, aes(x = season_day, y = publication_lag, fill = mean_relative_error)) +
        geom_tile(width = 1, height = 1) +
        scale_x_continuous(
            breaks = season_month_breaks,
            labels = season_month_labels,
            limits = c(0, season_axis_end),
            expand = expansion(mult = 0)
        ) +
        scale_y_continuous(
            breaks = heatmap_day_breaks,
            limits = c(0, heatmap_focus_days),
            expand = expansion(mult = 0)
        ) +
        scale_fill_stepsn(
            colors = relative_error_step_cols,
            values = rescale(relative_error_step_midpoints, from = c(0, 1)),
            breaks = relative_error_step_breaks,
            labels = c("0.5%", "1%", "5%", "10%", "20%", "50%"),
            limits = c(0, 1),
            oob = squish,
            guide = guide_colorsteps(
                direction = "horizontal",
                even.steps = TRUE,
                barheight = grid::unit(0.75, "cm"),
                barwidth = grid::unit(31.2, "cm"),
                title.position = "top",
                title.hjust = 0.5,
                label.position = "bottom",
                ticks = TRUE
            )
        ) +
        labs(
            title = sprintf("Relative Error heatmap: %s", season_value),
            subtitle = "Cells show mean Relative Error across all states; white cells indicate no available report.",
            x = "Reference month",
            y = "Lag (days)",
            fill = "Relative Error"
        ) +
        theme(
            legend.position = "bottom",
            legend.box = "vertical",
            legend.title = element_text(hjust = 0.5, size = 22, face = "bold"),
            legend.text = element_text(size = 20),
            legend.key.width = grid::unit(1.9, "cm"),
            legend.key.height = grid::unit(0.75, "cm"),
            plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
            plot.subtitle = element_text(size = 22, hjust = 0.5),
            axis.title = element_text(size = 24, face = "bold"),
            axis.text = element_text(size = 20),
            plot.margin = margin(10, 14, 10, 14)
        )
}

heatmap_season_levels <- rev(season_levels)
heatmap_panel_html <- map_chr(heatmap_season_levels, function(season_value) {
    image_uri <- plot_to_data_uri(
        plot_relative_error_heatmap(season_value),
        width = 18,
        height = 8.2
    )
    display <- ifelse(season_value == heatmap_season_levels[1], "block", "none")
    sprintf(
        '<div class="relative-error-season-panel" data-season="%s" style="display:%s;"><img src="%s" style="width:100%%;" alt="Relative Error heatmap for %s"></div>',
        season_value, display, image_uri, season_value
    )
})

heatmap_control_html <- paste0(
    '<div class="relative-error-season-control">',
    '<button type="button" class="relative-error-season-step" data-direction="-1">Previous</button>',
    '<div class="relative-error-season-buttons">',
    paste0(
        '<button type="button" class="relative-error-season-button" data-season="',
        heatmap_season_levels,
        '">',
        heatmap_season_levels,
        '</button>',
        collapse = ""
    ),
    '</div>',
    '<button type="button" class="relative-error-season-step" data-direction="1">Next</button>',
    '<span class="relative-error-season-status" aria-live="polite"></span>',
    '</div>'
)

heatmap_script_html <- paste(
    '<script>',
    '(function() {',
    '  const panels = document.querySelectorAll(".relative-error-season-panel");',
    '  const buttons = document.querySelectorAll(".relative-error-season-button");',
    '  const steps = document.querySelectorAll(".relative-error-season-step");',
    '  const status = document.querySelector(".relative-error-season-status");',
    '  if (!panels.length || !buttons.length) return;',
    '  let currentPage = 0;',
    '  function updatePanels(pageIndex) {',
    '    currentPage = Math.max(0, Math.min(pageIndex, buttons.length - 1));',
    '    const season = buttons[currentPage].dataset.season;',
    '    panels.forEach(function(panel) {',
    '      panel.style.display = panel.dataset.season === season ? "block" : "none";',
    '    });',
    '    buttons.forEach(function(button, index) {',
    '      const isActive = index === currentPage;',
    '      button.classList.toggle("active", isActive);',
    '      button.setAttribute("aria-current", isActive ? "page" : "false");',
    '    });',
    '    steps.forEach(function(button) {',
    '      const direction = Number(button.dataset.direction);',
    '      button.disabled = (direction < 0 && currentPage === 0) || (direction > 0 && currentPage === buttons.length - 1);',
    '    });',
    '    if (status) status.textContent = "Page " + (currentPage + 1) + " of " + buttons.length + ": " + season;',
    '  }',
    '  buttons.forEach(function(button, index) {',
    '    button.addEventListener("click", function() { updatePanels(index); });',
    '  });',
    '  steps.forEach(function(button) {',
    '    button.addEventListener("click", function() {',
    '      updatePanels(currentPage + Number(button.dataset.direction));',
    '    });',
    '  });',
    '  updatePanels(0);',
    '})();',
    '</script>',
    sep = "\n"
)

heatmap_style_html <- paste(
    '<style>',
    '.relative-error-season-control {',
    '  display: flex;',
    '  align-items: center;',
    '  flex-wrap: wrap;',
    '  gap: 0.35rem;',
    '  margin: 0.75rem 0 1rem 0;',
    '}',
    '.relative-error-season-buttons {',
    '  display: flex;',
    '  flex-wrap: wrap;',
    '  gap: 0.25rem;',
    '}',
    '.relative-error-season-button, .relative-error-season-step {',
    '  border: 1px solid #d0d7de;',
    '  background: #fff;',
    '  color: #24292f;',
    '  border-radius: 4px;',
    '  padding: 0.25rem 0.55rem;',
    '  font-size: 0.9rem;',
    '  line-height: 1.2;',
    '  cursor: pointer;',
    '}',
    '.relative-error-season-button.active {',
    '  background: #0969da;',
    '  border-color: #0969da;',
    '  color: #fff;',
    '}',
    '.relative-error-season-step:disabled {',
    '  color: #8c959f;',
    '  background: #f6f8fa;',
    '  cursor: default;',
    '}',
    '.relative-error-season-status {',
    '  color: #57606a;',
    '  margin-left: 0.35rem;',
    '  font-size: 0.9rem;',
    '}',
    '</style>',
    sep = "\n"
)

cat(paste(
    heatmap_control_html,
    paste(heatmap_panel_html, collapse = "\n"),
    heatmap_script_html,
    heatmap_style_html,
    sep = "\n"
))
```

---

## Is the first published Report value biased? {#initial-bias}

Here we ask whether the first published Report value tends to be below or above the
latest available Report value.

For each location-reference-date pair we compute Initial Bias:

$$\text{Initial Bias} = \frac{r_f - r_l}{r_l}$$

Here `r_f` is the Report value in the first report and `r_l` is the Report value
in the latest report for the same location and reference date.

A negative Initial Bias means the first published Report value was below the latest Report value
(under-reporting); a positive value means it was above the latest Report value
(over-reporting).

The boxplot below shows the distribution of Initial Bias across reference dates
for each location. The box spans the 25th-75th percentile range, the middle line
is the median, and the whiskers use the standard 1.5 IQR rule. The color is based
on each location's median Initial Bias direction, so individual pairs
inside a colored box can still fall on either side of zero.

```{r}
#| label: bias-compute
bias_df <- pair_summary %>%
    transmute(
        geo_value,
        time_value,
        value_first = first_report_value,
        value_latest = latest_report_value,
        initial_bias
    )

bias_by_loc <- bias_df %>%
    group_by(geo_value) %>%
    summarize(
        median_bias = median(initial_bias, na.rm = TRUE),
        pct_under   = mean(initial_bias < 0, na.rm = TRUE),
        pct_over    = mean(initial_bias > 0, na.rm = TRUE),
        .groups     = "drop"
    ) %>%
    arrange(median_bias)
```

```{r}
#| label: bias-overview-plot
#| fig-height: 5
bias_df %>%
    left_join(bias_by_loc %>% select(geo_value, median_bias), by = "geo_value") %>%
    mutate(
        geo_value = factor(geo_value, levels = bias_by_loc$geo_value),
        direction = case_when(
            median_bias < 0 ~ "Under-reporting",
            median_bias > 0 ~ "Over-reporting",
            TRUE ~ "No Initial Bias"
        )
    ) %>%
    ggplot(aes(x = geo_value, y = initial_bias, fill = direction)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7) +
    geom_hline(yintercept = 0, linewidth = 0.4, color = "gray40") +
    scale_x_discrete(labels = toupper) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    coord_cartesian(ylim = c(
        min(0, quantile(bias_df$initial_bias, 0.025, na.rm = TRUE)),
        max(0, quantile(bias_df$initial_bias, 0.975, na.rm = TRUE))
    )) +
    labs(
        title = "Initial Bias distribution by location",
        subtitle = "Distribution of Initial Bias by location; color indicates median direction.",
        x = NULL, y = "Initial Bias", fill = NULL
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 11, face = "bold"))
```

The table below summarizes the median Initial Bias by location, along with the share of
location-reference-date pairs that were initially below or above the latest Report value.

```{r}
#| label: bias-table
#| echo: false
DT::datatable(
    bias_by_loc %>% mutate(geo_value = toupper(geo_value)),
    colnames = c(
        "Location", "Median Initial Bias",
        "% Under-reporting", "% Over-reporting"
    ),
    caption = sprintf(
        "First published Report value compared with latest Report value by location for %s. Negative median Initial Bias = systematic under-reporting.",
        params$signal_name
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, pageLength = params$max_locations_table, dom = "tp"
    )
) %>%
    DT::formatRound(columns = 2, digits = 3) %>%
    DT::formatPercentage(columns = 3:4, digits = 1)
```

## How soon can we trust the signal? {#maturation}

This section asks which lag is needed before Report values are close
enough to their latest Report value. The thresholds use Relative Error, defined in
Results at a glance, and describe persistent stability: once a
location-reference-date pair enters the threshold, it must stay there.

```{r}
#| label: stability-compute
#| cache: true
thresholds <- c(0.5, 0.2, 0.1, 0.05, 0.01, 0.005)
threshold_label <- function(x) {
    ifelse(x < 0.01, label_percent(accuracy = 0.1)(x), label_percent(accuracy = 1)(x))
}

stability_multi <- convergence_df %>%
    arrange(geo_value, time_value, publication_lag) %>%
    group_by(geo_value, time_value) %>%
    group_modify(function(df, key) {
        map_dfr(thresholds, function(tau) {
            over <- which(df$abs_rel_diff > tau)
            lag_stable <- if (length(over) == 0) {
                min(df$publication_lag, na.rm = TRUE)
            } else {
                df$publication_lag[min(max(over) + 1, nrow(df))]
            }

            tibble(
                lag_stable = lag_stable,
                threshold = threshold_label(tau),
                threshold_num = tau
            )
        })
    }) %>%
    ungroup()

rev_window_df <- rev_beh %>%
    filter(n_revisions > 0) %>%
    transmute(
        geo_value, time_value,
        lag_stable = as_days(max_lag),
        threshold = "0%",
        threshold_num = 0
    )

stability_combined <- bind_rows(stability_multi, rev_window_df) %>%
    mutate(threshold = factor(threshold, levels = threshold_levels))
invisible(gc())
```

### Cumulative stabilization by Relative Error threshold

The curve answers: as lag increases, what fraction of location-reference-date pairs have settled
within the specified Relative Error threshold and never left it? Higher
thresholds are easier to satisfy, so the curves and stacked bands are nested:
the 50% threshold should mature earlier than stricter thresholds such as 5%,
0.5%, or 0%.

```{r}
#| label: revision-duration-ecdf-multi
#| fig-height: 5
max_ecdf_day <- max(stability_combined$lag_stable, na.rm = TRUE)

stability_ecdf <- stability_combined %>%
    mutate(
        threshold = factor(threshold, levels = threshold_levels)
    ) %>%
    group_by(threshold) %>%
    group_modify(function(df, key) {
        days <- seq(0, max_ecdf_day)
        tibble(
            lag_stable = days,
            cumulative_fraction = map_dbl(days, ~ mean(df$lag_stable <= .x, na.rm = TRUE))
        )
    }) %>%
    ungroup()

stability_wide <- stability_ecdf %>%
    pivot_wider(names_from = threshold, values_from = cumulative_fraction)

stability_bands <- bind_rows(
    stability_wide %>% transmute(lag_stable, band = "50% to 100%", ymin = `50%`, ymax = 1),
    stability_wide %>% transmute(lag_stable, band = "50% to 20%", ymin = `20%`, ymax = `50%`),
    stability_wide %>% transmute(lag_stable, band = "20% to 10%", ymin = `10%`, ymax = `20%`),
    stability_wide %>% transmute(lag_stable, band = "10% to 5%", ymin = `5%`, ymax = `10%`),
    stability_wide %>% transmute(lag_stable, band = "5% to 1%", ymin = `1%`, ymax = `5%`),
    stability_wide %>% transmute(lag_stable, band = "1% to 0.5%", ymin = `0.5%`, ymax = `1%`),
    stability_wide %>% transmute(lag_stable, band = "Within 0.5%", ymin = 0, ymax = `0.5%`)
) %>%
    mutate(
        lower = pmin(ymin, ymax, na.rm = TRUE),
        upper = pmax(ymin, ymax, na.rm = TRUE),
        ymin = lower,
        ymax = upper,
        band = factor(
            band,
            levels = c(
                "Within 0.5%", "1% to 0.5%", "5% to 1%",
                "10% to 5%", "20% to 10%", "50% to 20%",
                "50% to 100%"
            )
        )
    ) %>%
    select(-lower, -upper)

final_revision_ecdf <- stability_ecdf %>%
    filter(threshold == "0%")

max_day <- max(stability_ecdf$lag_stable, na.rm = TRUE)
right_start <- if (max_day > tail_start) tail_start else focus_days

stability_plot <- function(x_limits, show_y = TRUE) {
    ggplot() +
        geom_ribbon(
            data = stability_bands,
            aes(x = lag_stable, ymin = ymin, ymax = ymax, fill = band),
            alpha = 0.52,
            show.legend = FALSE
        ) +
        geom_ribbon(
            data = final_revision_ecdf,
            aes(x = lag_stable, ymin = 0, ymax = cumulative_fraction),
            fill = final_revision_fill,
            alpha = 0.45,
            show.legend = FALSE
        ) +
        geom_line(
            data = stability_ecdf,
            aes(x = lag_stable, y = cumulative_fraction, color = threshold),
            linewidth = 0.9
        ) +
        geom_hline(yintercept = 1, color = "black", linewidth = 0.7) +
        scale_x_continuous(limits = x_limits, breaks = focus_day_breaks(x_limits), expand = expansion(mult = 0)) +
        scale_y_continuous(labels = label_percent(), limits = c(0, 1), expand = expansion(mult = 0)) +
        scale_color_manual(values = threshold_cols) +
        scale_fill_manual(values = band_cols) +
        guides(color = guide_legend(byrow = TRUE)) +
        labs(
            x = "Lag (days)",
            y = if (show_y) "Cumulative fraction of pairs" else NULL,
            color = "Threshold"
        ) +
        theme(
            axis.title.y = if (show_y) element_text() else element_blank(),
            axis.text.y = if (show_y) element_text() else element_blank(),
            axis.ticks.y = if (show_y) element_line() else element_blank()
        )
}

make_cumulative_stabilization_plot <- function(data, title = NULL, show_legend = FALSE) {
    data <- data %>%
        filter(!is.na(lag_stable), !is.na(threshold)) %>%
        mutate(threshold = factor(threshold, levels = threshold_levels))

    if (nrow(data) == 0) {
        return(ggplot() + labs(title = title) + theme_void())
    }

    cumulative_data <- data %>%
        group_by(threshold) %>%
        group_modify(function(df, key) {
            tibble(
                lag_stable = seq(0, focus_days),
                cumulative_fraction = map_dbl(lag_stable, ~ mean(df$lag_stable <= .x, na.rm = TRUE))
            )
        }) %>%
        ungroup()

    cumulative_wide <- cumulative_data %>%
        pivot_wider(names_from = threshold, values_from = cumulative_fraction)

    cumulative_bands <- bind_rows(
        cumulative_wide %>% transmute(lag_stable, band = "50% to 100%", ymin = `50%`, ymax = 1),
        cumulative_wide %>% transmute(lag_stable, band = "50% to 20%", ymin = `20%`, ymax = `50%`),
        cumulative_wide %>% transmute(lag_stable, band = "20% to 10%", ymin = `10%`, ymax = `20%`),
        cumulative_wide %>% transmute(lag_stable, band = "10% to 5%", ymin = `5%`, ymax = `10%`),
        cumulative_wide %>% transmute(lag_stable, band = "5% to 1%", ymin = `1%`, ymax = `5%`),
        cumulative_wide %>% transmute(lag_stable, band = "1% to 0.5%", ymin = `0.5%`, ymax = `1%`),
        cumulative_wide %>% transmute(lag_stable, band = "Within 0.5%", ymin = 0, ymax = `0.5%`)
    ) %>%
        mutate(
            lower = pmin(ymin, ymax, na.rm = TRUE),
            upper = pmax(ymin, ymax, na.rm = TRUE),
            ymin = lower,
            ymax = upper,
            band = factor(
                band,
                levels = c(
                    "Within 0.5%", "1% to 0.5%", "5% to 1%",
                    "10% to 5%", "20% to 10%", "50% to 20%",
                    "50% to 100%"
                )
            )
        ) %>%
        select(-lower, -upper)

    cumulative_final <- cumulative_data %>%
        filter(threshold == "0%")

    ggplot() +
        geom_ribbon(
            data = cumulative_bands,
            aes(x = lag_stable, ymin = ymin, ymax = ymax, fill = band),
            alpha = 0.52,
            show.legend = FALSE
        ) +
        geom_ribbon(
            data = cumulative_final,
            aes(x = lag_stable, ymin = 0, ymax = cumulative_fraction),
            fill = final_revision_fill,
            alpha = 0.45,
            show.legend = FALSE
        ) +
        geom_line(
            data = cumulative_data,
            aes(x = lag_stable, y = cumulative_fraction, color = threshold),
            linewidth = 0.8
        ) +
        geom_hline(yintercept = 1, color = "black", linewidth = 0.5) +
        scale_x_continuous(limits = c(0, focus_days), breaks = focus_day_breaks(c(0, focus_days)), expand = expansion(mult = 0)) +
        scale_y_continuous(labels = label_percent(), limits = c(0, 1), expand = expansion(mult = 0)) +
        scale_color_manual(values = threshold_cols) +
        scale_fill_manual(values = band_cols) +
        guides(color = if (show_legend) guide_legend(byrow = TRUE) else "none") +
        labs(
            title = title,
            x = "Lag (days)",
            y = "Cumulative fraction",
            color = "Threshold"
        ) +
        theme(
            plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
            axis.title = element_text(size = 12, face = "bold"),
            axis.text = element_text(size = 11),
            legend.title = element_text(size = 11, face = "bold"),
            legend.text = element_text(size = 11)
        )
}

left_plot <- stability_plot(c(0, min(focus_days, max_day)), show_y = TRUE)
right_plot <- stability_plot(c(right_start, max_day), show_y = FALSE)

(left_plot + right_plot + plot_layout(widths = c(3, 1), guides = "collect")) +
    plot_annotation(
        title = "Cumulative stabilization by Relative Error threshold",
        subtitle = sprintf(
            "Axis break after %d days; right panel shows the long tail from day %d onward.",
            focus_days, right_start
        )
    )
```

The table below gives the key readouts from the cumulative curves. For each
threshold, the columns answer how many days are needed before a given fraction
of location-reference-date pairs have reached persistent stability.

```{r}
#| label: stability-summary-table
#| echo: false
stability_combined %>%
    group_by(threshold) %>%
    summarize(
        median = median(lag_stable, na.rm = TRUE),
        p75 = quantile(lag_stable, 0.75, na.rm = TRUE),
        p90 = quantile(lag_stable, 0.90, na.rm = TRUE),
        p95 = quantile(lag_stable, 0.95, na.rm = TRUE),
        p99 = quantile(lag_stable, 0.99, na.rm = TRUE),
        max = max(lag_stable, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(threshold = factor(threshold, levels = threshold_levels)) %>%
    arrange(threshold) %>%
    DT::datatable(
        colnames = c(
            "Threshold",
            "P50 days",
            "P75 days",
            "P90 days",
            "P95 days",
            "P99 days",
            "Max days"
        ),
        caption = "Key wait-time readouts from the cumulative stabilization curves",
        rownames = FALSE,
        options = list(dom = "t", paging = FALSE)
    ) %>%
    DT::formatRound(columns = 2:7, digits = 1)
```

### Relative Error by lag

This view summarizes Relative Error at each lag. The line is the median across
reports, and the band shows the 25th-75th percentile range.
The x-axis is split: the left panel shows early lags, and the right panel shows
the long tail. Days between the two panels are intentionally omitted and should
not be read as continuous spacing.

```{r}
#| label: convergence-overall-plot
#| fig-height: 6
#| cache: true
overall_conv <- convergence_df %>%
    group_by(publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_reports = n(),
        .groups = "drop"
    ) %>%
    filter(n_reports > 5)

convergence_focus_days <- focus_days
max_conv_day <- max(overall_conv$publication_lag, na.rm = TRUE)
right_conv_start <- if (max_conv_day > tail_start) tail_start else convergence_focus_days

revision_distance_plot <- function(x_limits, show_y = TRUE) {
    ggplot(overall_conv, aes(x = publication_lag)) +
        geom_ribbon(
            aes(ymin = q25_rel_diff, ymax = q75_rel_diff),
            fill = dis_pal[1],
            alpha = 0.22
        ) +
        geom_line(aes(y = median_rel_diff), color = dis_pal[1], linewidth = 1) +
        scale_x_continuous(
            limits = x_limits,
            breaks = focus_day_breaks(x_limits, convergence_focus_days),
            expand = expansion(mult = 0)
        ) +
        scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
        labs(
            x = "Lag (days)",
            y = if (show_y) "Relative Error" else NULL
        ) +
        theme(
            axis.title.y = if (show_y) element_text() else element_blank(),
            axis.text.y = if (show_y) element_text() else element_blank(),
            axis.ticks.y = if (show_y) element_line() else element_blank()
        )
}

revision_distance_left <- revision_distance_plot(c(0, min(convergence_focus_days, max_conv_day)), show_y = TRUE)
revision_distance_right <- revision_distance_plot(c(right_conv_start, max_conv_day), show_y = FALSE)

(revision_distance_left + revision_distance_right + plot_layout(widths = c(3, 1))) +
    plot_annotation(
        title = "Relative Error by lag",
        subtitle = sprintf(
            "Axis break: days %d-%d are omitted. Line: median; band: 25th-75th percentile range.",
            convergence_focus_days, right_conv_start
        )
    )
```

### Stabilization time distribution by Relative Error threshold

This plot shows the shape of the wait-time distribution behind the cumulative
curves above. Bars show the empirical distribution within each threshold; the
smooth line is a density guide on the same approximate proportion scale.
This is the overall stabilization-time distribution. Later sections reuse the
same stabilization-time quantity, first split by location and then by
reference date.

```{r}
#| label: stability-histograms
#| fig-height: 8
#| cache: true
stability_distribution <- stability_combined %>%
    filter(!is.na(lag_stable)) %>%
    mutate(threshold = factor(threshold, levels = threshold_levels))

distribution_bin_width <- 14
distribution_max_day <- ceiling(max(stability_distribution$lag_stable, na.rm = TRUE) / distribution_bin_width) *
    distribution_bin_width
distribution_bin_mids <- seq(
    distribution_bin_width / 2,
    distribution_max_day + distribution_bin_width / 2,
    by = distribution_bin_width
)

stability_distribution_bars <- stability_distribution %>%
    mutate(
        bin_start = floor(lag_stable / distribution_bin_width) * distribution_bin_width,
        bin_mid = bin_start + distribution_bin_width / 2
    ) %>%
    count(threshold, bin_mid, name = "n") %>%
    complete(threshold, bin_mid = distribution_bin_mids, fill = list(n = 0)) %>%
    left_join(count(stability_distribution, threshold, name = "threshold_total"), by = "threshold") %>%
    mutate(proportion = n / threshold_total)

stability_distribution_density <- stability_distribution %>%
    group_by(threshold) %>%
    group_modify(function(df, key) {
        if (n_distinct(df$lag_stable) < 2) {
            return(tibble(x = numeric(), y = numeric()))
        }
        dens <- density(
            df$lag_stable,
            from = 0,
            to = distribution_max_day,
            adjust = 1.1,
            n = 512
        )
        tibble(
            x = dens$x,
            y = dens$y * distribution_bin_width
        )
    }) %>%
    ungroup()

ggplot() +
    geom_col(
        data = stability_distribution_bars,
        aes(x = bin_mid, y = proportion, fill = threshold),
        width = distribution_bin_width * 0.95,
        color = "white",
        linewidth = 0.15,
        alpha = 0.72
    ) +
    geom_line(
        data = stability_distribution_density,
        aes(x = x, y = y, color = threshold),
        linewidth = 0.9,
        na.rm = TRUE
    ) +
    facet_wrap(~threshold, ncol = 1, scales = "free_y") +
    scale_x_continuous(breaks = days_since_breaks(c(0, distribution_max_day)), expand = expansion(mult = 0)) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    scale_fill_manual(values = threshold_cols, guide = "none") +
    scale_color_manual(values = threshold_cols) +
    guides(color = "none") +
    labs(
        title = "Stabilization time distribution by Relative Error threshold",
        subtitle = sprintf(
            "Bars show binned proportions; smoothed lines show density scaled to the same %d-day bin width.",
            distribution_bin_width
        ),
        x = "Lag at stabilization (days)",
        y = "Proportion of pairs"
    )
```

---

## Is revision behavior different by location? {#by-location}

Revision behavior can vary by geography. This section first compares the total
amount of revision activity and lag by location, then compares
how long each location takes to reach the stability thresholds used above.
Locations are grouped by HHS region (U.S. Department of Health and Human
Services region).

```{r}
#| label: rev-beh-by-loc-compute
hhs_geo_order <- hhs_region_map %>%
    mutate(
        hhs_region = factor(hhs_region, levels = paste("HHS", 1:10)),
        geo_value = tolower(geo_value)
    ) %>%
    filter(geo_value %in% unique(tolower(stability_combined$geo_value))) %>%
    arrange(hhs_region, geo_value) %>%
    mutate(
        x_pos = row_number(),
        geo_label = toupper(geo_value),
        hhs_label = str_replace(as.character(hhs_region), "HHS ", "HHS\nRegion ")
    )

unmapped_geos <- setdiff(unique(tolower(stability_combined$geo_value)), hhs_geo_order$geo_value)
if (length(unmapped_geos) > 0) {
    hhs_geo_order <- bind_rows(
        hhs_geo_order,
        tibble(
            geo_value = sort(unmapped_geos),
            hhs_region = factor("Other", levels = c(paste("HHS", 1:10), "Other")),
            x_pos = seq(nrow(hhs_geo_order) + 1, nrow(hhs_geo_order) + length(unmapped_geos)),
            geo_label = toupper(sort(unmapped_geos)),
            hhs_label = "Other"
        )
    )
}

geo_sorting <- hhs_geo_order$geo_value

hhs_region_bounds <- hhs_geo_order %>%
    group_by(hhs_region, hhs_label) %>%
    summarize(
        xmin = min(x_pos),
        xmax = max(x_pos),
        xmid = mean(x_pos),
        .groups = "drop"
    )

hhs_bracket_plot <- function() {
    ggplot(hhs_region_bounds) +
        geom_segment(
            aes(x = xmin, xend = xmax, y = 0.72, yend = 0.72),
            color = "gray45",
            linewidth = 0.55
        ) +
        geom_segment(
            aes(x = xmin, xend = xmin, y = 0.72, yend = 0.95),
            color = "gray45",
            linewidth = 0.55
        ) +
        geom_segment(
            aes(x = xmax, xend = xmax, y = 0.72, yend = 0.95),
            color = "gray45",
            linewidth = 0.55
        ) +
        geom_text(
            aes(x = xmid, y = 0.22, label = hhs_label),
            size = 3.5,
            fontface = "bold",
            color = "gray20"
        ) +
        scale_x_continuous(
            limits = range(hhs_geo_order$x_pos) + c(-0.5, 0.5),
            expand = expansion(mult = c(0.01, 0.01))
        ) +
        coord_cartesian(ylim = c(0, 1), clip = "off") +
        labs(x = "Location ordered by HHS region") +
        theme_void() +
        theme(
            axis.title.x = element_text(size = 13, face = "bold", margin = margin(t = 14)),
            plot.margin = margin(-4, 5.5, 16, 5.5)
        )
}

rev_by_loc <- rev_beh %>%
    mutate(geo_value = tolower(geo_value)) %>%
    group_by(geo_value) %>%
    summarize(
        n_pairs = n(),
        total_revisions = sum(n_revisions, na.rm = TRUE),
        avg_revisions = mean(n_revisions, na.rm = TRUE),
        median_window_days = median(as_days(max_lag), na.rm = TRUE),
        p90_window_days = quantile(as_days(max_lag), 0.9, na.rm = TRUE),
        avg_rel_spread = mean(rel_spread, na.rm = TRUE),
        median_conv_days = median(as_days(lag_near_latest), na.rm = TRUE),
        .groups = "drop"
    )
```

### Cumulative stabilization by Relative Error threshold by HHS region

For each state, the curve shows the fraction of location-reference-date pairs that have settled
within each Relative Error threshold by a given lag and never left it. The view
is paged by HHS region (U.S. Department of Health and Human Services region).
Each small panel is one state. Lines and stacked bands use the same nested
threshold colors as Section 2.1; compare panels to see whether states in the
same HHS region stabilize at similar lags.

```{r}
#| label: cumulative-stabilization-by-hhs-state
#| echo: false
#| results: asis
#| cache: true
hhs_regions <- paste("HHS", 1:10)

hhs_panel_html <- map_chr(hhs_regions, function(region_name) {
    region_geos <- hhs_geo_order %>%
        filter(as.character(hhs_region) == region_name) %>%
        pull(geo_value)

    state_plots <- map(region_geos, function(geo) {
        make_cumulative_stabilization_plot(
            stability_combined %>% filter(tolower(geo_value) == geo),
            title = toupper(geo),
            show_legend = TRUE
        )
    })

    region_panels <- wrap_plots(state_plots, ncol = 2)

    region_plot <- (guide_area() / region_panels) +
        plot_layout(heights = c(0.22, 1), guides = "collect") +
        plot_annotation(
            title = sprintf("Cumulative stabilization by Relative Error threshold: %s", region_name),
            subtitle = sprintf("Each panel is one state; colors show nested Relative Error thresholds. Lags 0-%d days.", focus_days)
        )

    image_uri <- plot_to_data_uri(
        region_plot,
        width = 14,
        height = max(6, 3.6 * ceiling(length(region_geos) / 2) + 1.6)
    )

    display <- ifelse(region_name == hhs_regions[1], "block", "none")
    sprintf(
        '<div class="hhs-cumulative-panel" data-region="%s" style="display:%s;"><img src="%s" style="width:100%%;" alt="%s cumulative stabilization by state"></div>',
        region_name, display, image_uri, region_name
    )
})

control_html <- paste0(
    '<div class="hhs-cumulative-control">',
    '<button type="button" class="hhs-page-step" data-direction="-1">Previous</button>',
    '<div class="hhs-page-buttons">',
    paste0(
        '<button type="button" class="hhs-page-button" data-region="',
        hhs_regions,
        '" data-page="',
        seq_along(hhs_regions),
        '">',
        seq_along(hhs_regions),
        '</button>',
        collapse = ""
    ),
    '</div>',
    '<button type="button" class="hhs-page-step" data-direction="1">Next</button>',
    '<span class="hhs-page-status" aria-live="polite"></span>',
    '</div>'
)

script_html <- paste(
    '<script>',
    '(function() {',
    '  const panels = document.querySelectorAll(".hhs-cumulative-panel");',
    '  const pageButtons = document.querySelectorAll(".hhs-page-button");',
    '  const stepButtons = document.querySelectorAll(".hhs-page-step");',
    '  const status = document.querySelector(".hhs-page-status");',
    '  if (!panels.length || !pageButtons.length) return;',
    '  let currentPage = 0;',
    '  function updatePanels(pageIndex) {',
    '    currentPage = Math.max(0, Math.min(pageIndex, pageButtons.length - 1));',
    '    const region = pageButtons[currentPage].dataset.region;',
    '    panels.forEach(function(panel) {',
    '      panel.style.display = panel.dataset.region === region ? "block" : "none";',
    '    });',
    '    pageButtons.forEach(function(button, index) {',
    '      const isActive = index === currentPage;',
    '      button.classList.toggle("active", isActive);',
    '      button.setAttribute("aria-current", isActive ? "page" : "false");',
    '    });',
    '    stepButtons.forEach(function(button) {',
    '      const direction = Number(button.dataset.direction);',
    '      button.disabled = (direction < 0 && currentPage === 0) || (direction > 0 && currentPage === pageButtons.length - 1);',
    '    });',
    '    if (status) status.textContent = "Page " + (currentPage + 1) + " of " + pageButtons.length + ": " + region;',
    '  }',
    '  pageButtons.forEach(function(button, index) {',
    '    button.addEventListener("click", function() { updatePanels(index); });',
    '  });',
    '  stepButtons.forEach(function(button) {',
    '    button.addEventListener("click", function() {',
    '      updatePanels(currentPage + Number(button.dataset.direction));',
    '    });',
    '  });',
    '  updatePanels(0);',
    '})();',
    '</script>',
    sep = "\n"
)

style_html <- paste(
    '<style>',
    '.hhs-cumulative-control {',
    '  display: flex;',
    '  align-items: center;',
    '  flex-wrap: wrap;',
    '  gap: 0.35rem;',
    '  margin: 0.75rem 0 1rem 0;',
    '  position: relative;',
    '  z-index: 10;',
    '}',
    '.hhs-page-buttons {',
    '  display: flex;',
    '  flex-wrap: wrap;',
    '  gap: 0.25rem;',
    '}',
    '.hhs-page-button, .hhs-page-step {',
    '  border: 1px solid #d0d7de;',
    '  background: #fff;',
    '  color: #24292f;',
    '  border-radius: 4px;',
    '  padding: 0.25rem 0.55rem;',
    '  font-size: 0.9rem;',
    '  line-height: 1.2;',
    '  cursor: pointer;',
    '  pointer-events: auto;',
    '}',
    '.hhs-page-button.active {',
    '  background: #0969da;',
    '  border-color: #0969da;',
    '  color: #fff;',
    '}',
    '.hhs-page-step:disabled {',
    '  color: #8c959f;',
    '  background: #f6f8fa;',
    '  cursor: default;',
    '}',
    '.hhs-page-status {',
    '  color: #57606a;',
    '  margin-left: 0.35rem;',
    '  font-size: 0.9rem;',
    '}',
    '</style>',
    sep = "\n"
)

cat(paste(
    control_html,
    paste(hhs_panel_html, collapse = "\n"),
    script_html,
    style_html,
    sep = "\n"
))
```

### Relative Error by HHS region

This view shows how Relative Error changes with lag within each HHS region.
Each panel is one HHS region, each line is a state median, and each shaded band
is the 25th-75th percentile range for that state.

```{r}
#| label: convergence-location-plot
#| fig-height: 22
#| cache: true
region_conv <- convergence_df %>%
    mutate(geo_value = tolower(geo_value)) %>%
    inner_join(hhs_region_map, by = "geo_value") %>%
    group_by(hhs_region, geo_value, publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_reports = n(),
        .groups = "drop"
    ) %>%
    filter(n_reports > 1) %>%
    filter(publication_lag <= convergence_focus_days) %>%
    mutate(
        hhs_region = factor(hhs_region, levels = paste("HHS", 1:10)),
        geo_value = factor(geo_value, levels = sort(unique(geo_value)))
    )

plot_hhs_region <- function(region_name) {
    region_data <- region_conv %>%
        filter(hhs_region == region_name) %>%
        mutate(geo_value = droplevels(geo_value))

    state_levels <- levels(region_data$geo_value)
    state_cols <- setNames(hue_pal()(length(state_levels)), state_levels)
    state_labels <- setNames(toupper(state_levels), state_levels)

    ggplot(region_data, aes(x = publication_lag, group = geo_value)) +
        geom_ribbon(
            aes(ymin = q25_rel_diff, ymax = q75_rel_diff, fill = geo_value),
            alpha = 0.12,
            color = NA
        ) +
        geom_line(aes(y = median_rel_diff, color = geo_value), linewidth = 0.75, alpha = 0.95) +
        scale_x_continuous(
            limits = c(0, convergence_focus_days),
            breaks = focus_day_breaks(c(0, convergence_focus_days), convergence_focus_days),
            expand = expansion(mult = 0)
        ) +
        scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
        scale_color_manual(values = state_cols, labels = state_labels) +
        scale_fill_manual(values = state_cols, guide = "none") +
        labs(
            title = region_name,
            x = "Lag (days)",
            y = "Relative Error",
            color = NULL
        ) +
        theme(
            plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
            axis.title = element_text(size = 12, face = "bold"),
            axis.text = element_text(size = 11),
            legend.position = c(0.98, 0.98),
            legend.justification = c(1, 1),
            legend.background = element_rect(fill = scales::alpha("white", 0.78), color = "gray80"),
            legend.text = element_text(size = 11),
            legend.key.height = unit(0.8, "lines"),
            legend.key.width = unit(1.0, "lines")
        ) +
        guides(color = guide_legend(nrow = 2, byrow = TRUE))
}

region_plots <- lapply(paste("HHS", 1:10), plot_hhs_region)

wrap_plots(region_plots, ncol = 2) +
    plot_annotation(
        title = "Relative Error by HHS region",
        subtitle = sprintf(
            "Relative Error across lags 0-%d days; lines show state-level medians and shaded bands show 25th-75th percentile ranges.",
            convergence_focus_days
        )
    )
```

### Stabilization time distribution by location

This is the location-level version of the stabilization-time distribution from
Section 2.2. Each point is the median time to reach a threshold for one
location; each vertical line is the 25th-75th percentile range. The main reading
is across states: compare whether some locations consistently need longer lags
to reach the same threshold. The `0%` threshold means the Report value has
matched the latest Report value, while the other thresholds allow nonzero
Relative Error.

```{r}
#| label: rev-by-location-maturation
#| fig-width: 12
#| fig-height: 7
#| eval: !expr n_locations > 3
#| cache: true
location_maturation_df <- stability_combined %>%
    filter(!is.na(lag_stable), !is.na(threshold)) %>%
    mutate(geo_value = tolower(geo_value)) %>%
    inner_join(hhs_geo_order, by = "geo_value") %>%
    mutate(
        geo_value = factor(geo_value, levels = geo_sorting)
    ) %>%
    group_by(geo_value, x_pos, geo_label, threshold) %>%
    summarize(
        median_lag = median(lag_stable, na.rm = TRUE),
        q25_lag = quantile(lag_stable, 0.25, na.rm = TRUE),
        q75_lag = quantile(lag_stable, 0.75, na.rm = TRUE),
        .groups = "drop"
    )

location_maturation_plot <- ggplot(
    location_maturation_df,
    aes(x = x_pos, y = median_lag, ymin = q25_lag, ymax = q75_lag, color = threshold)
) +
    geom_pointrange(position = position_dodge(width = 0.7), linewidth = 0.45, size = 0.9) +
    scale_x_continuous(
        breaks = hhs_geo_order$x_pos,
        labels = hhs_geo_order$geo_label,
        expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_y_continuous(
        limits = c(0, NA),
        breaks = days_since_breaks(c(0, max(location_maturation_df$q75_lag, na.rm = TRUE)))
    ) +
    scale_color_manual(values = threshold_cols) +
    guides(color = guide_legend(byrow = TRUE)) +
    labs(
        title = "Stabilization time distribution by location",
        subtitle = "Lag needed to reach each stability threshold (dots = median, lines = IQR)",
        x = NULL, y = "Lag at stabilization (days)",
        color = "Threshold"
    ) +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 11, face = "bold")
    )

location_maturation_plot / hhs_bracket_plot() + plot_layout(heights = c(12, 2.1))
```

---

## Are there patterns over reference date? {#time-patterns}

This section asks whether revision behavior changes across reference dates. It
first summarizes the amount of revision activity over time, then repeats the
same two-part reading pattern used above: cumulative stabilization first, then
the Relative Error curve. The final plot turns back to stabilization time,
matching the quantity used in Sections 2.2 and 3.3.

Rather than using calendar years, these views use epi seasons
(cross-year periods used to align respiratory-virus timing). CDC
respiratory virus surveillance is commonly organized by MMWR weeks and named as
cross-year seasons. In this report, each season follows a flu-surveillance-style
MMWR season: MMWR week 40 through MMWR week 39 of the following year. Seasons are
labeled by the two calendar years they span (for example, `2022-2023`).

```{r}
#| label: season-definitions
#| echo: false
stability_season_df <- stability_combined %>%
    mutate(epi_season = factor(epi_season_label(time_value), levels = season_levels))
```

### Cumulative stabilization by Relative Error threshold and epi season

For each epi season, the curve shows the fraction of location-reference-date pairs that
have settled within each Relative Error threshold by a given lag and never left
it. Higher thresholds are easier to satisfy, so the nested color pattern should
be interpreted the same way as in the overall cumulative stabilization plot.

```{r}
#| label: cumulative-stabilization-by-season
#| fig-width: 12
#| fig-height: 7
#| cache: true
season_stabilization_plots <- map(season_levels, function(season_value) {
    make_cumulative_stabilization_plot(
        stability_season_df %>% filter(epi_season == season_value),
        title = season_value,
        show_legend = TRUE
    )
})

season_stabilization_panels <- wrap_plots(season_stabilization_plots, ncol = 2)

(guide_area() / season_stabilization_panels) +
    plot_layout(heights = c(0.22, 1), guides = "collect") +
    plot_annotation(
        title = "Cumulative stabilization by Relative Error threshold and epi season",
        subtitle = sprintf(
            "Shown separately for each MMWR week 40-39 epi season. Lags 0-%d days.",
            focus_days
        )
    )
```

### Relative Error by epi season

This view shows how Relative Error changes with lag across epi seasons.
Lines are medians, and shaded bands show the 25th-75th percentile range within
each season.

```{r}
#| label: relative-error-by-season
#| fig-height: 5
#| cache: true
season_conv <- convergence_season_df %>%
    group_by(epi_season, publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_reports = n(),
        .groups = "drop"
    ) %>%
    filter(n_reports > 5) %>%
    filter(publication_lag <= convergence_focus_days)

ggplot(season_conv, aes(x = publication_lag, color = epi_season, fill = epi_season)) +
    geom_ribbon(
        aes(ymin = q25_rel_diff, ymax = q75_rel_diff),
        alpha = 0.14,
        color = NA
    ) +
    geom_line(aes(y = median_rel_diff), linewidth = 0.9) +
    scale_x_continuous(
        limits = c(0, convergence_focus_days),
        breaks = focus_day_breaks(c(0, convergence_focus_days), convergence_focus_days),
        expand = expansion(mult = 0)
    ) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    labs(
        title = "Relative Error by epi season",
        subtitle = sprintf(
            "Seasons run MMWR week 40-39; lags 0-%d days.\nLines show medians and shaded bands show 25th-75th percentile ranges.",
            convergence_focus_days
        ),
        x = "Lag (days)",
        y = "Relative Error",
        color = "Epi season",
        fill = "Epi season"
    )
```

### Stabilization time distribution over reference date

This is the time-level version of the stabilization-time distribution from
Sections 2.2 and 3.3. Instead of showing the full distribution or splitting by
location, it asks whether the time to reach each threshold changes across
reference-date periods. Lines show the median stabilization time, and shaded bands
show the 25th-75th percentile range. Recent reference dates may be censored because
there has not yet been enough time to observe their full revision history. The
censoring boundary is not an epidemiological cutoff; it is the observable
maximum lag implied by the available report dates. Apparent jumps or short
discontinuities near that boundary are data-support effects rather than an
additional axis break.

```{r}
#| label: diagnostic-censoring
#| fig-height: 4.5
#| cache: true
max_v <- max(diagnostic_archive$DT$version)

date_range_days <- as.numeric(max(stability_combined$time_value) - min(stability_combined$time_value))
agg_unit <- case_when(
    date_range_days > 365 * 2 ~ "quarter",
    date_range_days > 365 ~ "month",
    date_range_days > 100 ~ "2-week",
    date_range_days > 40 ~ "week",
    date_range_days > 20 ~ "3-day",
    TRUE ~ "day"
)

floor_time_bucket <- function(date, unit) {
    date <- as.Date(date)
    if (unit == "2-week") {
        origin <- min(date, na.rm = TRUE)
        return(origin + 14 * floor(as.numeric(date - origin) / 14))
    }
    if (unit == "3-day") {
        origin <- min(date, na.rm = TRUE)
        return(origin + 3 * floor(as.numeric(date - origin) / 3))
    }
    floor_date(date, unit = unit)
}

stability_time_summary <- stability_combined %>%
    filter(!is.na(lag_stable), !is.na(threshold)) %>%
    mutate(
        threshold = factor(threshold, levels = threshold_levels),
        time_bucket = floor_time_bucket(time_value, agg_unit)
    ) %>%
    group_by(threshold, time_bucket) %>%
    summarize(
        median_lag = median(lag_stable, na.rm = TRUE),
        q25_lag = quantile(lag_stable, 0.25, na.rm = TRUE),
        q75_lag = quantile(lag_stable, 0.75, na.rm = TRUE),
        .groups = "drop"
    )

ggplot(stability_time_summary, aes(x = time_bucket, color = threshold, fill = threshold)) +
    geom_ribbon(
        aes(ymin = q25_lag, ymax = q75_lag),
        alpha = 0.22,
        color = NA
    ) +
    geom_line(aes(y = median_lag), linewidth = 0.85) +
    geom_function(
        fun = function(x) as.numeric(max_v - as.Date(x, origin = "1970-01-01")),
        linetype = "dashed", color = "gray50", linewidth = 0.5
    ) +
    annotate(
        "text",
        x = min(stability_time_summary$time_bucket),
        y = as_days(max_v - min(stability_combined$time_value)),
        label = "Observable maximum lag", color = "gray50",
        vjust = 1, hjust = -1, size = 3
    ) +
    scale_x_date(breaks = breaks_pretty()) +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty()) +
    scale_color_manual(values = threshold_cols) +
    scale_fill_manual(values = threshold_cols) +
    labs(
        title = "Stabilization time distribution over reference date",
        subtitle = sprintf(
            "Reference dates grouped by %s. Lines show medians; shaded bands show the 25th-75th percentile range.",
            agg_unit
        ),
        x = "Reference date",
        y = "Lag at stabilization (days)",
        color = "Threshold",
        fill = "Threshold"
    )
```

---

## Are revision patterns related to other signal features? {#signal-features}

After checking first-release behavior, maturation time, geography, and
epi season, this section asks whether other observable features of the
signal are related to revision behavior. These views are exploratory: they help
identify where revision patterns may be driven by signal magnitude, initial
lag (the lag of the first report observed for a location-reference-date pair),
or finer reference-date structure.

### Latest Report value vs. Relative Error spread

This plot compares signal magnitude with the range of Relative Error observed
over the report history for the same location and reference date.

- Point: one location-reference-date pair.
- X-axis: the latest Report value for that pair.
- Y-axis: Relative Error spread, defined as maximum Relative Error minus minimum Relative Error across reports.
- Interpretation: a larger y-value means the pair moved through a wider range of distances from its latest Report value during the revision process.

```{r}
#| label: diagnostic-magnitude
#| fig-height: 4.5
#| cache: true
relative_error_spread_df <- convergence_df %>%
    group_by(geo_value, time_value) %>%
    summarize(
        relative_error_spread = max(abs_rel_diff, na.rm = TRUE) - min(abs_rel_diff, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    inner_join(
        pair_summary %>% select(geo_value, time_value, latest_report_value),
        by = c("geo_value", "time_value")
    ) %>%
    filter(
        latest_report_value > 0,
        relative_error_spread > 0,
        is.finite(latest_report_value),
        is.finite(relative_error_spread)
    )

relative_error_spread_df %>%
    ggplot(aes(x = latest_report_value, y = relative_error_spread)) +
    geom_point(alpha = 0.2, size = 1) +
    geom_smooth() +
    scale_x_log10(labels = label_comma()) +
    scale_y_log10(labels = label_percent()) +
    labs(
        title = "Latest Report value vs. Relative Error spread",
        x = "Latest Report value (log scale)",
        y = "Relative Error spread (log scale)"
    )
```

### Relative Error by initial lag

Early publication can be useful, but it may also leave more room for later
revision. This view groups location-reference-date pairs by initial lag (the lag of the first
report observed for each location-reference-date pair), then compares how
Relative Error changes across lags.

```{r}
#| label: relative-error-by-initial-lag
#| fig-height: 5
#| cache: true
initial_lag_df <- convergence_df %>%
    group_by(geo_value, time_value) %>%
    summarize(
        initial_publication_lag = min(publication_lag, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(
        initial_lag_group = cut(
            initial_publication_lag,
            breaks = c(-Inf, 7, 14, 30, Inf),
            labels = c("<=7 days", "8-14 days", "15-30 days", ">30 days"),
            right = TRUE
        )
    )

initial_lag_revision_df <- convergence_df %>%
    inner_join(
        initial_lag_df %>% select(geo_value, time_value, initial_lag_group),
        by = c("geo_value", "time_value")
    ) %>%
    filter(!is.na(initial_lag_group), publication_lag <= convergence_focus_days) %>%
    group_by(initial_lag_group, publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_reports = n(),
        .groups = "drop"
    ) %>%
    filter(n_reports > 5)

ggplot(initial_lag_revision_df, aes(x = publication_lag, color = initial_lag_group, fill = initial_lag_group)) +
    geom_ribbon(aes(ymin = q25_rel_diff, ymax = q75_rel_diff), alpha = 0.14, color = NA) +
    geom_line(aes(y = median_rel_diff), linewidth = 0.95) +
    scale_x_continuous(
        limits = c(0, convergence_focus_days),
        breaks = focus_day_breaks(c(0, convergence_focus_days), convergence_focus_days),
        expand = expansion(mult = 0)
    ) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    labs(
        title = "Relative Error by initial lag",
        subtitle = "Groups are based on the first observed lag for each reference date.",
        x = "Lag (days)",
        y = "Relative Error",
        color = "Initial lag",
        fill = "Initial lag"
    )
```

---

## Additional revision diagnostics {#additional-diagnostics}

These diagnostics are useful for checking data support and broad
revision patterns, but they are secondary to the main interpretation sections
above.

### Overall revision diagnostics

This section collects secondary diagnostics for data support and revision
activity. These quantities are useful checks, but the main interpretation should
come from the Relative Error, Initial Bias, and stabilization results above.

```{r}
#| label: revision-analysis-print
#| echo: false
#| results: asis
quick_threshold <- 7
few_threshold <- 3
relative_error_spread_thr <- 0.1
abs_spread_thr <- 0.05 * rev_result$max_val

total_obs        <- nrow(rev_beh)
no_rev_pct       <- mean(rev_beh$n_revisions == 0, na.rm = TRUE)
quick_rev_pct    <- mean(as_days(rev_beh$max_lag) <= quick_threshold, na.rm = TRUE)
few_rev_pct      <- mean(rev_beh$n_revisions <= few_threshold, na.rm = TRUE)

revised_beh      <- rev_beh %>% filter(n_revisions > 0)
low_relative_error_spread <- mean(relative_error_spread_df$relative_error_spread < relative_error_spread_thr, na.rm = TRUE)
high_abs_spread  <- mean(revised_beh$spread > abs_spread_thr, na.rm = TRUE)

lag_stats  <- summary(as_days(rev_beh$min_lag))
conv_stats <- summary(as_days(rev_beh$lag_near_latest))
fmt_stats  <- function(s) paste(names(s), round(as.numeric(s), 1), sep = ": ", collapse = ", ")

if (n_excluded > 0) {
    cat(sprintf(
        "::: {.callout-important}\n`revision_analysis()` requires at least one report after the waiting period to assess revision behavior. %d location-reference-date pair(s) were excluded because they only have an initial report with no subsequent reports yet. The %d remaining pairs are used for all analyses below.\n:::\n\n",
        n_excluded, n_analyzed
    ))
}
cat(sprintf(
    "- **Time range**: %s to %s\n",
    rev_result$range_time_values[1], rev_result$range_time_values[2]
))
cat(sprintf("- **Min lag (days to first report)**: %s\n", fmt_stats(lag_stats)))
cat(sprintf("- **No revisions**: %s\n", label_percent()(no_rev_pct)))
cat(sprintf("- **Quick revisions (≤ %d days)**: %s\n", quick_threshold, label_percent()(quick_rev_pct)))
cat(sprintf("- **Few revisions (≤ %d)**: %s\n", few_threshold, label_percent()(few_rev_pct)))
if (nrow(revised_beh) > 0) {
    cat(sprintf(
        "- **Minor Relative Error spread (< %.1f)**: %s\n",
        relative_error_spread_thr, label_percent()(low_relative_error_spread)
    ))
    cat(sprintf(
        "- **Major absolute revisions (> %.3f)**: %s\n",
        abs_spread_thr, label_percent()(high_abs_spread)
    ))
}
cat(sprintf(
    "- **Days until within %d%% of latest Report value**: %s\n\n",
    rev_result$within_latest * 100, fmt_stats(conv_stats)
))
```

::: {.content-visible when-meta="multi_location"}

The location table summarizes average revision activity, first-report timing,
latest-report lag, absolute spread in Report values, Relative Error spread, and
the average lag until the Report value stays within the default convergence
threshold of the latest Report value.

```{r}
#| label: location-stats-compute
#| include: false
relative_error_spread_by_location <- relative_error_spread_df %>%
    group_by(geo_value) %>%
    summarize(
        rel_error_spread = mean(relative_error_spread, na.rm = TRUE),
        median_rel_error_spread = median(relative_error_spread, na.rm = TRUE),
        .groups = "drop"
    )

location_stats <- rev_beh %>%
    group_by(geo_value) %>%
    summarize(
        n_rev = mean(n_revisions),
        min_lag = min(min_lag),
        max_lag = max(max_lag),
        spread = mean(spread),
        median_window = median(as_days(max_lag), na.rm = TRUE),
        lag_near_latest = mean(lag_near_latest),
        .groups = "drop"
    ) %>%
    left_join(relative_error_spread_by_location, by = "geo_value")

# Find highest/lowest for the commentary
sorted_conv <- location_stats %>% arrange(desc(lag_near_latest))
highest_loc <- sorted_conv$geo_value[1]
lowest_loc <- sorted_conv$geo_value[nrow(sorted_conv)]
```

```{r}
#| label: location-stats-table
#| echo: false
#| eval: !expr n_locations > 1
DT::datatable(
    location_stats %>%
        mutate(geo_value = toupper(geo_value)) %>%
        select(
            geo_value, n_rev, min_lag, max_lag, spread,
            rel_error_spread, median_rel_error_spread,
            median_window, lag_near_latest
        ),
    colnames = c(
        "Location", "Avg # revisions", "Min lag", "Max lag",
        "Avg spread", "Avg Relative Error spread", "Med. Relative Error spread",
        "Med. window (d)", "Avg conv. (d)"
    ),
    caption = sprintf(
        "Revision behavior statistics by location for %s.",
        params$signal_name
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, pageLength = params$max_locations_table, dom = "tp"
    )
) %>%
    DT::formatRound(columns = c(2, 5, 6, 7, 8, 9), digits = 3)
```

:::

### Number of revisions per location

This figure counts revision events by location across all reference dates. It is
intended as a data-support diagnostic: locations with many revisions have deeper
report histories, while locations with few revisions may provide less stable
evidence about long-lag behavior.

```{r}
#| label: revisions-by-location-barplot
#| fig-width: 12
#| fig-height: 5
#| eval: !expr n_locations > 3
revisions_by_location_plot <- rev_by_loc %>%
    inner_join(hhs_geo_order, by = "geo_value") %>%
    ggplot(aes(x = x_pos, y = total_revisions, fill = hhs_region)) +
    geom_col(width = 0.82, alpha = 0.88) +
    scale_x_continuous(
        breaks = hhs_geo_order$x_pos,
        labels = hhs_geo_order$geo_label,
        expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_y_continuous(labels = label_comma(), breaks = breaks_pretty()) +
    labs(
        title = "Number of revisions per location",
        subtitle = "Total revision events across all reference dates",
        x = NULL,
        y = "Number of revisions",
        fill = "HHS region"
    ) +
    guides(fill = "none") +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 11, face = "bold"),
        legend.position = "none"
    )

revisions_by_location_plot / hhs_bracket_plot() + plot_layout(heights = c(12, 2.1))
```

### Lag distribution by location

The diagnostic below shows the distribution of lag for all states in
the selected HHS region. Here, lag is the number of days between the reference
date and the report date. The split between first reports and later reports
helps separate initial reporting timing from later revisions.

```{r}
#| label: backfill-gap-compute
#| cache: true
backfill_df <- diagnostic_report_df %>%
    group_by(geo_value, time_value) %>%
    mutate(
        backfill_gap = publication_lag,
        is_initial_release = version == min(version, na.rm = TRUE)
    ) %>%
    ungroup()
```

```{r}
#| label: backfill-gap-geography
#| echo: false
#| results: asis
#| cache: true
lag_hhs_regions <- paste("HHS", 1:10)

lag_panel_html <- map_chr(lag_hhs_regions, function(region_name) {
    region_geos <- hhs_geo_order %>%
        filter(as.character(hhs_region) == region_name) %>%
        pull(geo_value)

    lag_plot <- backfill_df %>%
        mutate(geo_value = tolower(geo_value)) %>%
        filter(geo_value %in% region_geos) %>%
        mutate(
            geo_value = factor(geo_value, levels = region_geos, labels = toupper(region_geos))
        ) %>%
        ggplot(aes(x = backfill_gap, fill = is_initial_release)) +
        geom_histogram(bins = 40, position = "identity", alpha = 0.7) +
        facet_wrap(~geo_value, ncol = 2, scales = "free_y") +
        scale_fill_manual(
            values = c("TRUE" = dis_pal[2], "FALSE" = dis_pal[1]),
            labels = c("TRUE" = "First report", "FALSE" = "Later report"),
            name = "Report type"
        ) +
        scale_x_continuous(breaks = breaks_pretty()) +
        labs(
            title = sprintf("Lag distribution: %s", region_name),
            subtitle = "First-report timing vs. later revision depth",
            x = "Lag (days)",
            y = "Count"
        ) +
        theme(legend.position = "top", strip.text.x = element_text(size = 12, face = "bold"))

    image_uri <- plot_to_data_uri(
        lag_plot,
        width = 14,
        height = max(6, 3.4 * ceiling(length(region_geos) / 2) + 1.6)
    )

    display <- ifelse(region_name == lag_hhs_regions[1], "block", "none")
    sprintf(
        '<div class="lag-hhs-panel" data-region="%s" style="display:%s;"><img src="%s" style="width:100%%;" alt="%s lag by state"></div>',
        region_name, display, image_uri, region_name
    )
})

lag_control_html <- paste0(
    '<div class="lag-hhs-control">',
    '<button type="button" class="lag-page-step" data-direction="-1">Previous</button>',
    '<div class="lag-page-buttons">',
    paste0(
        '<button type="button" class="lag-page-button" data-region="',
        lag_hhs_regions,
        '" data-page="',
        seq_along(lag_hhs_regions),
        '">',
        seq_along(lag_hhs_regions),
        '</button>',
        collapse = ""
    ),
    '</div>',
    '<button type="button" class="lag-page-step" data-direction="1">Next</button>',
    '<span class="lag-page-status" aria-live="polite"></span>',
    '</div>'
)

lag_script_html <- paste(
    '<script>',
    '(function() {',
    '  const panels = document.querySelectorAll(".lag-hhs-panel");',
    '  const pageButtons = document.querySelectorAll(".lag-page-button");',
    '  const stepButtons = document.querySelectorAll(".lag-page-step");',
    '  const status = document.querySelector(".lag-page-status");',
    '  if (!panels.length || !pageButtons.length) return;',
    '  let currentPage = 0;',
    '  function updatePanels(pageIndex) {',
    '    currentPage = Math.max(0, Math.min(pageIndex, pageButtons.length - 1));',
    '    const region = pageButtons[currentPage].dataset.region;',
    '    panels.forEach(function(panel) {',
    '      panel.style.display = panel.dataset.region === region ? "block" : "none";',
    '    });',
    '    pageButtons.forEach(function(button, index) {',
    '      const isActive = index === currentPage;',
    '      button.classList.toggle("active", isActive);',
    '      button.setAttribute("aria-current", isActive ? "page" : "false");',
    '    });',
    '    stepButtons.forEach(function(button) {',
    '      const direction = Number(button.dataset.direction);',
    '      button.disabled = (direction < 0 && currentPage === 0) || (direction > 0 && currentPage === pageButtons.length - 1);',
    '    });',
    '    if (status) status.textContent = "Page " + (currentPage + 1) + " of " + pageButtons.length + ": " + region;',
    '  }',
    '  pageButtons.forEach(function(button, index) {',
    '    button.addEventListener("click", function() { updatePanels(index); });',
    '  });',
    '  stepButtons.forEach(function(button) {',
    '    button.addEventListener("click", function() {',
    '      updatePanels(currentPage + Number(button.dataset.direction));',
    '    });',
    '  });',
    '  updatePanels(0);',
    '})();',
    '</script>',
    sep = "\n"
)

lag_style_html <- paste(
    '<style>',
    '.lag-hhs-control {',
    '  display: flex;',
    '  align-items: center;',
    '  flex-wrap: wrap;',
    '  gap: 0.35rem;',
    '  margin: 0.75rem 0 1rem 0;',
    '  position: relative;',
    '  z-index: 10;',
    '}',
    '.lag-page-buttons {',
    '  display: flex;',
    '  flex-wrap: wrap;',
    '  gap: 0.25rem;',
    '}',
    '.lag-page-button, .lag-page-step {',
    '  border: 1px solid #d0d7de;',
    '  background: #fff;',
    '  color: #24292f;',
    '  border-radius: 4px;',
    '  padding: 0.25rem 0.55rem;',
    '  font-size: 0.9rem;',
    '  line-height: 1.2;',
    '  cursor: pointer;',
    '  pointer-events: auto;',
    '}',
    '.lag-page-button.active {',
    '  background: #0969da;',
    '  border-color: #0969da;',
    '  color: #fff;',
    '}',
    '.lag-page-step:disabled {',
    '  color: #8c959f;',
    '  background: #f6f8fa;',
    '  cursor: default;',
    '}',
    '.lag-page-status {',
    '  color: #57606a;',
    '  margin-left: 0.35rem;',
    '  font-size: 0.9rem;',
    '}',
    '</style>',
    sep = "\n"
)

cat(paste(
    lag_control_html,
    paste(lag_panel_html, collapse = "\n"),
    lag_script_html,
    lag_style_html,
    sep = "\n"
))
```

### Number of revisions over reference date

This figure summarizes how the number of revisions per location changes over
reference dates. The line shows the median across locations, and the shaded band
shows the 25th-75th percentile range.

```{r}
#| label: revision-count-plot
#| echo: false
#| fig-height: 4.5
#| cache: true
rev_count_df <- diagnostic_report_df %>%
    group_by(geo_value, time_value) %>%
    summarize(n_revisions = n_distinct(version) - 1L, .groups = "drop")

rev_count_summary <- rev_count_df %>%
    group_by(time_value) %>%
    summarize(
        q25_revisions = quantile(n_revisions, 0.25, na.rm = TRUE),
        median_revisions = median(n_revisions, na.rm = TRUE),
        q75_revisions = quantile(n_revisions, 0.75, na.rm = TRUE),
        n_locations = n_distinct(geo_value),
        .groups = "drop"
    )

ggplot(rev_count_summary, aes(x = time_value)) +
    geom_ribbon(
        aes(ymin = q25_revisions, ymax = q75_revisions, fill = "25th-75th percentile"),
        alpha = 0.22,
        color = NA
    ) +
    geom_line(aes(y = median_revisions, color = "Median"), linewidth = 0.9) +
    scale_color_manual(values = c("Median" = "steelblue4")) +
    scale_fill_manual(values = c("25th-75th percentile" = "steelblue")) +
    scale_x_date(breaks = breaks_pretty()) +
    scale_y_continuous(breaks = breaks_pretty()) +
    labs(
        title = "Number of revisions by reference date",
        subtitle = "Line shows the median across locations; shaded band shows the 25th-75th percentile range.",
        x = "Reference date",
        y = "Number of revisions per location",
        color = NULL,
        fill = NULL
    )
```

### Lag distribution by epi season

The diagnostic below shows lag by epi season. Each panel is one MMWR
week 40-39 season, allowing the seasonal timing pattern to be compared directly.

```{r}
#| label: report-lag-by-season
#| fig-height: 7
backfill_df %>%
    mutate(epi_season = factor(epi_season_label(time_value), levels = season_levels)) %>%
    filter(!is.na(epi_season)) %>%
    ggplot(aes(x = backfill_gap, fill = is_initial_release)) +
    geom_histogram(bins = 40, position = "identity", alpha = 0.7, color = "white") +
    facet_wrap(~epi_season, ncol = 2, scales = "free_y") +
        scale_fill_manual(
            values = c("TRUE" = dis_pal[2], "FALSE" = dis_pal[1]),
            labels = c("TRUE" = "First report", "FALSE" = "Later report"),
            name = "Report type"
        ) +
    scale_x_continuous(breaks = breaks_pretty()) +
    labs(
        title = "Lag distribution by epi season",
        subtitle = "MMWR week 40-39 epi seasons; first-report timing vs. later revision depth",
        x = "Lag (days)",
        y = "Count"
    ) +
    theme(strip.text.x = element_text(size = 12, face = "bold"))
```

## Summary {#summary}

```{r}
#| label: final-summary-narrative
#| echo: false
#| results: asis
# Statistical context for the narrative
final_rev_data <- stability_combined %>% filter(threshold == "0%")
med_window <- round(median(final_rev_data$lag_stable, na.rm = TRUE))
p90_window <- round(quantile(final_rev_data$lag_stable, 0.9, na.rm = TRUE))

p_no_rev <- label_percent()(no_rev_pct)
pct_under <- label_percent()(mean(bias_by_loc$median_bias < 0))

cat(sprintf(
    "For **%s** (`%s / %s`), the diagnostic reference dates covering %s to %s\nshow the following revision and backfill profile:\n\n",
    params$signal_name, params$source, params$signal,
    rev_result$range_time_values[1], rev_result$range_time_values[2]
))

cat(sprintf("- **%s** of location-reference-date pairs were never revised after their first published Report value.\n", p_no_rev))
cat(sprintf(
    "- The **median lag to persistent stability is %d days**: half of all revised ",
    med_window
))
cat(sprintf(
    "location-reference-date pairs reached their latest Report value by lag %d.\n",
    med_window
))
cat(sprintf(
    "- The 90th percentile for maturity is **%d days** (~%d weeks).\n",
    p90_window, ceiling(p90_window / 7)
))
cat(sprintf(
    "- **%s of locations** exhibit systematic under-reporting in their first published Report value.\n",
    pct_under
))

# Threshold maturation summary
stab_summary_meds <- stability_combined %>%
    group_by(threshold) %>%
    summarize(median = round(median(lag_stable, na.rm = TRUE)), .groups = "drop")

cat(sprintf("- **Median days to reach stability**: "))
cat(paste0(
    "**", stab_summary_meds$threshold, "**: ", stab_summary_meds$median, " days",
    collapse = " | "
))
cat(".\n\n")
```
