---
title: "Revision Analysis: `r params$signal_name`"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    max-width: 1280px
params:
  # Signal to analyze
  source: 
  signal: 
  signal_name: 
  # Optional path to a local CSV or Parquet directory with a version/issue column.
  # When NULL, the signal is fetched from the Delphi API with issues = "*".
  input_dir: !expr "NULL"
  # Shared settings
  geo_type:  
  time_type: 
  start_day: 
  end_day:   
  # Maximum locations shown in faceted plots / summary tables.
  max_locations_plot: 6
  max_locations_table: 15
  # Number of worst-behaving locations to highlight.
  n_worst: 18
  # Convergence threshold: how close (relative) a version must be to the
  # latest known value to be considered "converged".
  convergence_threshold: 0.05
knitr:
  opts_chunk:
    comment: "#>"
    message: FALSE
    warning: FALSE
    out.width: "100%"
    dev: "png"
    fig.width: 9
    fig.height: 5
    fig.align: center
editor_options:
  chunk_output_type: console
execute:
  cache: false
---

```{r}
#| label: setup
#| include: false
#| cache: false
# When running interactively, fall back to the parameter defaults.
if (!exists("params")) {
    params <- list(
        source                = "hospital-admissions",
        signal                = "smoothed_covid19_from_claims",
        signal_name           = "Hospital Admissions: Smoothed COVID-19 from Claims",
        input_dir             = "revision_analysis/data/revisions/",
        geo_type              = "state",
        time_type             = "day",
        start_day             = "2023-01-01",
        end_day               = "2023-02-01",
        max_locations_plot    = 6,
        max_locations_table   = 15,
        n_worst               = 18,
        convergence_threshold = 0.05
    )
}

options(repos = c(
    delphi = "https://cmu-delphi.r-universe.dev",
    CRAN = "https://cloud.r-project.org"
))
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    epidatr, epiprocess, covidcast,
    tidyverse, scales, patchwork,
    knitr, here, MMWRweek, glue, DT,
    fs, arrow
)

# Invalidate knitr cache when params change
knitr::opts_chunk$set(cache.extra = list(params))

colours <- c(
    "#172431",
    "#ffffff",
    "#C41230",
    "#941120",
    "#007BC0",
    "#002676",
    "#FDB515",
    "#002145",
    "#00A7E1",
    "#8C338F",
    "#368f33"
)

dis_pal <- colours[c(3, 5, 7, 10, 11, 9)]

theme_set(
    theme_minimal(base_family = "sans", base_size = 13) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, color = "gray45", size = 12),
            plot.caption = element_text(hjust = 0, color = "gray50", size = 11),
            axis.title = element_text(size = 12, face = "bold"),
            axis.text = element_text(size = 11),
            legend.title = element_text(size = 11, face = "bold"),
            legend.text = element_text(size = 11),
            strip.text = element_text(size = 12, face = "bold"),
            panel.grid = element_line(linetype = 2, linewidth = 0.3, color = "gray90"),
            panel.grid.minor = element_blank(),
            legend.position = "top",
            strip.background = element_rect(fill = "gray95", linetype = "blank"),
            panel.border = element_rect(color = "gray95", fill = NA)
        )
)

# Palette as global defaults for discrete and continuous scales
options(
    ggplot2.discrete.colour   = dis_pal,
    ggplot2.discrete.fill     = dis_pal,
    ggplot2.continuous.colour = function() scale_colour_gradientn(colours = dis_pal),
    ggplot2.continuous.fill   = function() scale_fill_gradientn(colours = dis_pal)
)

tibble::tribble(
    ~name,      ~aes,     ~alpha, ~is_stat,
    "line",     "colour", 1.0,    FALSE,
    "point",    "colour", 0.7,    FALSE,
    "function", "colour", 1.0,    FALSE,
    "bar",      "fill",   0.9,    FALSE,
    "col",      "fill",   0.9,    FALSE,
    "boxplot",  "fill",   0.8,    FALSE,
    "bin",      "fill",   0.8,    TRUE,
    "identity", "fill",   0.7,    TRUE
) %>%
    purrr::pwalk(function(name, aes, alpha, is_stat) {
        updater <- if (is_stat) update_stat_defaults else update_geom_defaults
        params <- list(dis_pal[1], alpha)
        names(params) <- c(aes, "alpha")
        if (name == "boxplot") params$outlier.shape <- NA
        updater(name, params)
    })

# Helper: build epirange with correct format for the signal's time_type
time_range <- function(start, end, time_type = params$time_type) {
    if (time_type == "week") {
        to_yw <- function(d) {
            d <- as.Date(d)
            m <- MMWRweek::MMWRweek(d)
            m$MMWRyear * 100L + m$MMWRweek
        }
        epirange(to_yw(start), to_yw(end))
    } else {
        epirange(start, end)
    }
}

# Set cache
cache_path <- file.path(getwd(), "epidatr_cache")
if (!dir.exists(cache_path)) dir.create(cache_path, recursive = TRUE)

set_cache(cache_dir = cache_path, days = Inf)

# Helper: format a difftime object as a plain number of days
as_days <- function(x) as.numeric(x, units = "days")
```

This notebook analyses the **revision behaviour** of a single time series stored
as an `epi_archive`. The central question is:

> At what published-version lag does a series become reliable enough to use?

::: {.callout-important}
## Interpreting `version`, `issue`, and lag {.unnumbered}

`time_value` is the **reference date**: the date when the event or observation
occurred. The result for that reference date may be reported later, and those
successive report dates are stored as `version` or `issue`.

Lag is measured from the reference date to the corresponding report date. In the
data, the reference date is `time_value`, and the report date is the archive
`version` column or, in some inputs, `issue`.

```text
lag = report date - reference date
    = version - time_value
```

Interpret that version/issue date as the timestamp attached to the released or
uploaded signal value. It may reflect a publication rule, model run,
file-generation process, or API version tag rather than the first date when the
underlying raw data were available. Because this report uses daily data, lag
axes are labeled in days.

Before interpreting minimum lag, reporting latency, or stabilization time,
confirm what `version` / `issue` represents for this signal.
:::

```{r}
#| label: print-config
#| echo: false
config_tbl <- tibble(
    Field = c("Source", "Signal", "Label", "Geography", "Period"),
    Value = c(
        params$source, params$signal, params$signal_name,
        params$geo_type,
        paste(params$start_day, "to", params$end_day)
    )
)

knitr::kable(config_tbl, caption = "Analysis configuration")
```

---

## Data used in this report {#data .unnumbered}

```{r}
#| label: load-archive
#| results: asis
#| message: true
archive <- NULL

if (!is.null(params$input_dir)) {
    path <- here::here(params$input_dir)

    if (fs::is_dir(path)) {
        cli::cli_inform("Reading data from directory: {.path {params$input_dir}}")
        raw <- arrow::open_dataset(path) %>%
            select(any_of(c("geo_value", "time_value", "version", "issue", "value"))) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            ) %>%
            collect()
    } else if (grepl("\\.parquet$", path)) {
        cli::cli_inform("Reading Parquet file: {.path {params$input_dir}}")
        raw <- arrow::read_parquet(path)
    } else {
        cli::cli_inform("Reading CSV file: {.path {params$input_dir}}")
        raw <- read_csv(path)
    }

    raw <- raw %>% mutate(time_value = as.Date(time_value))

    # Accept either 'version' or 'issue' as the version column.
    if ("issue" %in% names(raw) && !"version" %in% names(raw)) {
        raw <- rename(raw, version = issue)
    }

    if (!"version" %in% names(raw)) {
        cli::cli_abort(
            "The input data must contain a 'version' or 'issue' column. ",
            "See the README for the required schema."
        )
    }

    # If it was a single file (CSV/Parquet), we still need to filter/mutate metadata
    if (!fs::is_dir(path)) {
        raw <- raw %>%
            mutate(version = as.Date(version)) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            )
    }

    archive <- raw %>%
        select(geo_value, time_value, version, value) %>%
        as_epi_archive()

    msg_verb <- "Loaded"
} else {
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE

    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                params$signal %in% meta_res[[params$source]]$signals
            },
            error = function(e) {
                FALSE
            }
        )
        if (meta_available) {
            use_new_api <- TRUE
        }
    }

    if (use_new_api) {
        df_raw <- epidata_archive(
            source = params$source,
            signals = params$signal,
            geo_type = params$geo_type
        ) %>%
            rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            )
    } else {
        start_date <- as.Date(params$start_day)
        end_date <- as.Date(params$end_day)
        interval_starts <- seq(start_date, end_date, by = "month")
        interval_ends <- pmin(
            lubridate::ceiling_date(interval_starts, "month") - lubridate::days(1),
            end_date
        )

        df_raw <- purrr::map2_dfr(interval_starts, interval_ends, function(s, e) {
            cli::cli_inform(c(">" = "Fetching API revisions for {s} to {e}"))

            pub_covidcast(
                source = params$source,
                signal = params$signal,
                geo_type = params$geo_type,
                time_type = params$time_type,
                time_values = time_range(s, e),
                issues = "*"
            )
        }) %>%
            distinct()
    }

    if ("issue" %in% colnames(df_raw) && !"version" %in% colnames(df_raw)) {
        df_raw <- rename(df_raw, version = issue)
    }

    if (nrow(df_raw) == 0 || !"version" %in% colnames(df_raw)) {
        cli::cli_abort(
            "No versioned data returned from the API. ",
            "The signal may not support revision history."
        )
    }

    archive <- df_raw %>%
        select(geo_value, time_value, version, value) %>%
        as_epi_archive()

    msg_verb <- "Fetched"
}

n_rows <- nrow(archive$DT)
n_locations <- n_distinct(archive$DT$geo_value)
n_versions <- n_distinct(archive$DT$version)
version_start <- min(archive$DT$version)
version_end <- max(archive$DT$version)

cli::cli_inform(c("i" = "{msg_verb} {label_comma()(n_rows)} rows | {label_comma()(n_locations)} locations | {label_comma()(n_versions)} versions ({version_start} – {version_end})"))

invisible(gc())
```

```{r}
#| label: quarto-metadata
#| results: asis
#| echo: false
# Dynamically set Quarto metadata so we can use it in ::: {.content-visible}
# blocks later.
n_years <- n_distinct(year(archive$DT$time_value))
cat(sprintf(
    "\n---\nmulti_location: %s\nmulti_year: %s\n---\n",
    tolower(n_locations > 1),
    tolower(n_years > 1)
))
```

```{r}
#| label: archive-meta
#| include: false
archive_span <- as.numeric(version_end - version_start, units = "days")
has_history <- archive_span >= 60
all_geos <- sort(unique(archive$DT$geo_value))
```

```{r}
#| label: archive-guard-message
#| echo: false
#| results: asis
if (!has_history) {
    cat(sprintf(
        "> **Warning**: only %d days of version history are available ",
        archive_span
    ))
    cat(sprintf(
        "(%s – %s). At least 60 days are needed for a meaningful ",
        version_start, version_end
    ))
    cat(sprintf(
        "revision analysis. The earliest this will be possible is %s.\n\n",
        version_start + 60
    ))
}
```

---

## Results at a glance {#overall .unnumbered}

This section computes the shared revision metrics used throughout the report.
The key quantities are the number of revisions, the size of revisions, and the
published-version lag at which values become stable or close to their latest
value. Later sections use these metrics to answer the interpretation questions
above.

```{r}
#| label: revision-analysis
rev_result <- epiprocess::revision_analysis(
    archive,
    value,
    min_waiting_period = 0
)
rev_beh <- rev_result$revision_behavior

# Track which (geo_value, time_value) pairs were included vs. excluded.
# revision_analysis() drops pairs that have no version past the waiting
# period — typically the most recent time values with only an initial report.
all_pairs <- archive$DT %>% distinct(geo_value, time_value)
analyzed_pairs <- rev_beh %>% distinct(geo_value, time_value)
n_excluded <- nrow(anti_join(all_pairs, analyzed_pairs, by = c("geo_value", "time_value")))
n_analyzed <- nrow(analyzed_pairs)

# Global metrics for follow-on plots/narrative
max_lag_days <- as_days(rev_beh$max_lag)
no_rev_pct <- mean(rev_beh$n_revisions == 0, na.rm = TRUE)
conv_days_all <- as_days(rev_beh$lag_near_latest)
invisible(gc())
```

```{r}
#| label: convergence-compute
convergence_df <- archive$DT %>%
    semi_join(analyzed_pairs, by = c("geo_value", "time_value")) %>%
    group_by(geo_value, time_value) %>%
    mutate(
        value_latest = value[which.max(version)]
    ) %>%
    ungroup() %>%
    filter(value_latest != 0, !is.na(value_latest), !is.na(value)) %>%
    mutate(
        publication_lag = as.numeric(version - time_value, units = "days"),
        abs_rel_diff = abs(value - value_latest) / abs(value_latest)
    ) %>%
    filter(publication_lag >= 0)
invisible(gc())
```

```{r}
#| label: revision-analysis-print
#| echo: false
#| results: asis
quick_threshold <- 7
few_threshold <- 3
rel_spread_thr <- 0.1
abs_spread_thr <- 0.05 * rev_result$max_val

total_obs        <- nrow(rev_beh)
no_rev_pct       <- mean(rev_beh$n_revisions == 0, na.rm = TRUE)
quick_rev_pct    <- mean(as_days(rev_beh$max_lag) <= quick_threshold, na.rm = TRUE)
few_rev_pct      <- mean(rev_beh$n_revisions <= few_threshold, na.rm = TRUE)

revised_beh      <- rev_beh %>% filter(n_revisions > 0)
low_rel_spread   <- mean(revised_beh$rel_spread < rel_spread_thr, na.rm = TRUE)
high_abs_spread  <- mean(revised_beh$spread > abs_spread_thr, na.rm = TRUE)

lag_stats  <- summary(as_days(rev_beh$min_lag))
conv_stats <- summary(as_days(rev_beh$lag_near_latest))
fmt_stats  <- function(s) paste(names(s), round(as.numeric(s), 1), sep = ": ", collapse = ", ")

if (n_excluded > 0) {
    cat(sprintf(
        "::: {.callout-important}\n`revision_analysis()` requires at least one version past the waiting period to assess revision behavior. %d time×location pair(s) were excluded because they only have an initial report with no subsequent versions yet. The %d remaining pairs are used for all analyses below.\n:::\n\n",
        n_excluded, n_analyzed
    ))
}
cat(sprintf(
    "- **Time range**: %s to %s\n",
    rev_result$range_time_values[1], rev_result$range_time_values[2]
))
cat(sprintf("- **Min lag (days to first published version)**: %s\n", fmt_stats(lag_stats)))
cat(sprintf("- **No revisions**: %s\n", label_percent()(no_rev_pct)))
cat(sprintf("- **Quick revisions (≤ %d days)**: %s\n", quick_threshold, label_percent()(quick_rev_pct)))
cat(sprintf("- **Few revisions (≤ %d)**: %s\n", few_threshold, label_percent()(few_rev_pct)))
if (nrow(revised_beh) > 0) {
    cat(sprintf(
        "- **Minor relative revisions (< %.1f)**: %s\n",
        rel_spread_thr, label_percent()(low_rel_spread)
    ))
    cat(sprintf(
        "- **Major absolute revisions (> %.3f)**: %s\n",
        abs_spread_thr, label_percent()(high_abs_spread)
    ))
}
cat(sprintf(
    "- **Days until within %d%% of latest**: %s\n\n",
    rev_result$within_latest * 100, fmt_stats(conv_stats)
))
```


::: {.content-visible when-meta="multi_location"}

Here we collect a number of statistics for each location. This includes:

- Avg # revisions. The average number of times a value was updated across all reference dates for that location.
- Min lag. The minimum number of days between the observation date and the first published version.
- Max lag. The maximum number of days between the event date and the final revision.
- Avg spread. The average absolute difference between the minimum and maximum published values.
- Avg rel. spread. The average relative difference between the minimum and maximum published values.
- Med. rel. spread. The median relative spread across all observation dates.
- Med. window (d). The median lag of the final revision.
- Avg conv. (d). The average days until the value stays within a certain percentage (default 20%) of the final value and never leaves it.

```{r}
#| label: location-stats-compute
#| include: false
location_stats <- rev_beh %>%
    group_by(geo_value) %>%
    summarize(
        n_rev = mean(n_revisions),
        min_lag = min(min_lag),
        max_lag = max(max_lag),
        spread = mean(spread),
        rel_spread = mean(rel_spread),
        median_rel_spread = median(rel_spread, na.rm = TRUE),
        median_window = median(as_days(max_lag), na.rm = TRUE),
        lag_near_latest = mean(lag_near_latest),
        .groups = "drop"
    )

# Find highest/lowest for the commentary
sorted_conv <- location_stats %>% arrange(desc(lag_near_latest))
highest_loc <- sorted_conv$geo_value[1]
lowest_loc <- sorted_conv$geo_value[nrow(sorted_conv)]
```

```{r}
#| label: location-stats-table
#| echo: false
#| eval: !expr n_locations > 1
DT::datatable(
    location_stats,
    colnames = c(
        "Location", "Avg # revisions", "Min lag", "Max lag",
        "Avg spread", "Avg rel. spread", "Med. rel. spread",
        "Med. window (d)", "Avg conv. (d)"
    ),
    caption = sprintf(
        "Revision behavior statistics by location for %s.",
        params$signal_name
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, pageLength = params$max_locations_table, dom = "tp"
    )
) %>%
    DT::formatRound(columns = c(2, 5, 6, 7, 8, 9), digits = 3)
```

:::

---

## Is the first published value biased? {#initial-bias}

### First published value vs. latest value

Here we ask whether the first published value tends to be below or above the
latest available value.

For each observation we compute:

$$\text{initial bias} = \frac{v_\text{first} - v_\text{latest}}{v_\text{latest}}$$

A negative value means the first published value was below the latest value
(under-reporting); a positive value means it was above the latest value
(over-reporting).

```{r}
#| label: bias-compute
first_vals <- archive$DT %>%
    group_by(geo_value, time_value) %>%
    slice_min(version, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    select(geo_value, time_value, value_first = value)

latest_vals <- archive$DT %>%
    group_by(geo_value, time_value) %>%
    slice_max(version, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    select(geo_value, time_value, value_latest = value)

bias_df <- first_vals %>%
    inner_join(latest_vals, by = c("geo_value", "time_value")) %>%
    select(geo_value, time_value, value_first, value_latest) %>%
    filter(value_latest != 0, !is.na(value_latest), !is.na(value_first)) %>%
    mutate(initial_bias = (value_first - value_latest) / abs(value_latest))

bias_by_loc <- bias_df %>%
    group_by(geo_value) %>%
    summarize(
        median_bias = median(initial_bias, na.rm = TRUE),
        pct_under   = mean(initial_bias < 0, na.rm = TRUE),
        pct_over    = mean(initial_bias > 0, na.rm = TRUE),
        .groups     = "drop"
    ) %>%
    arrange(median_bias)
```

```{r}
#| label: bias-overview-plot
#| fig-height: 5
bias_df %>%
    left_join(bias_by_loc %>% select(geo_value, median_bias), by = "geo_value") %>%
    mutate(
        geo_value = factor(geo_value, levels = bias_by_loc$geo_value),
        direction = case_when(
            median_bias < 0 ~ "Under-reporting",
            median_bias > 0 ~ "Over-reporting",
            TRUE ~ "No bias"
        )
    ) %>%
    ggplot(aes(x = geo_value, y = initial_bias, fill = direction)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7) +
    geom_hline(yintercept = 0, linewidth = 0.4, color = "gray40") +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    coord_cartesian(ylim = c(
        min(0, quantile(bias_df$initial_bias, 0.025, na.rm = TRUE)),
        max(0, quantile(bias_df$initial_bias, 0.975, na.rm = TRUE))
    )) +
    labs(
        title = "Initial-release bias distribution by location",
        subtitle = "Distribution of (first_value − latest_value) / latest_value",
        x = NULL, y = "Relative bias", fill = NULL
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10, face = "bold"))
```

The table below summarizes the median direction of the first published value by
location, along with the share of observations that were initially below or
above the latest value.

```{r}
#| label: bias-table
#| echo: false
DT::datatable(
    bias_by_loc,
    colnames = c(
        "Location", "Median rel. bias",
        "% Under-reporting", "% Over-reporting"
    ),
    caption = sprintf(
        "First published value compared with latest value by location for %s. Negative median bias = systematic under-reporting.",
        params$signal_name
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, pageLength = params$max_locations_table, dom = "tp"
    )
) %>%
    DT::formatRound(columns = 2, digits = 3) %>%
    DT::formatPercentage(columns = 3:4, digits = 1)
```

## How soon can we trust the signal? {#maturation}

This section asks which published-version lag is needed before values are close
enough to their latest value. The thresholds describe persistent stability: once
an observation enters the threshold, it must stay there.

Both plots in this section use the same distance-to-latest definition:

$$
\text{revision distance} =
\frac{|v_\text{published} - v_\text{latest}|}{|v_\text{latest}|}
$$

A value of 0% means the published version matches the latest value; a value of
10% means it is still 10% away from the latest value. We call this quantity
**revision distance**. The first plot summarizes when observations cross
practical revision-distance thresholds. The second plot shows the same distance
directly as a median curve with an interquartile band over lag. Later sections
reuse this same revision-distance curve by location and by observation season.

### Version date frequency

Before interpreting revision lag, it is helpful to check how often new archived
versions appear. This cadence affects which lags can be observed in the first
place.

```{r}
#| label: pipeline-issue-cadence
#| fig-height: 3.5
issue_dates <- sort(unique(archive$DT$version))
gaps <- as.numeric(diff(issue_dates), units = "days")

tibble(gap = gaps) %>%
    ggplot(aes(x = gap)) +
    geom_histogram(bins = 30, color = "white") +
    scale_x_continuous(breaks = breaks_pretty()) +
    labs(
        title = "Cadence of archive updates",
        subtitle = sprintf(
            "%d unique issue dates | median gap: %.0f d",
            n_versions, median(gaps)
        ),
        x = "Gap (days)", y = "Count"
    )
```

```{r}
#| label: stability-compute
thresholds <- c(0.5, 0.2, 0.1, 0.05, 0.01, 0.005)
threshold_label <- function(x) {
    ifelse(x < 0.01, label_percent(accuracy = 0.1)(x), label_percent(accuracy = 1)(x))
}

stability_multi <- map_dfr(thresholds, function(tau) {
    convergence_df %>%
        group_by(geo_value, time_value) %>%
        arrange(desc(version)) %>%
        summarize(
            lag_stable = {
                over <- which(abs_rel_diff > tau)
                if (length(over) == 0) {
                    min(publication_lag, na.rm = TRUE)
                } else {
                    lag_order <- order(publication_lag)
                    last_over <- max(match(over, lag_order), na.rm = TRUE)
                    publication_lag[lag_order[min(last_over + 1, length(lag_order))]]
                }
            },
            threshold = threshold_label(tau),
            threshold_num = tau,
            .groups = "drop"
        )
})

rev_window_df <- rev_beh %>%
    filter(n_revisions > 0) %>%
    transmute(
        geo_value, time_value,
        lag_stable = as_days(max_lag),
        threshold = "Final revision",
        threshold_num = 0
    )

stability_combined <- bind_rows(stability_multi, rev_window_df) %>%
    mutate(threshold = factor(threshold, levels = c("50%", "20%", "10%", "5%", "1%", "0.5%", "Final revision")))
invisible(gc())
```

### Cumulative stabilization by revision-distance threshold

The curve answers: as lag increases, what fraction of observations have settled
within the specified revision-distance threshold and never left it?

```{r}
#| label: revision-duration-ecdf-multi
#| fig-height: 5
threshold_levels <- c("50%", "20%", "10%", "5%", "1%", "0.5%", "Final revision")
threshold_cols <- c(
    "50%" = "#dc2626",
    "20%" = "#f59e0b",
    "10%" = "#facc15",
    "5%" = "#16a34a",
    "1%" = "#2563eb",
    "0.5%" = "#7c3aed",
    "Final revision" = "#92400e"
)
band_cols <- c(
    "50% to 100%" = "gray82",
    "50% to 20%" = unname(threshold_cols["50%"]),
    "20% to 10%" = unname(threshold_cols["20%"]),
    "10% to 5%" = unname(threshold_cols["10%"]),
    "5% to 1%" = unname(threshold_cols["5%"]),
    "1% to 0.5%" = unname(threshold_cols["1%"]),
    "Within 0.5%" = unname(threshold_cols["0.5%"])
)
final_revision_fill <- unname(threshold_cols["Final revision"])
max_ecdf_day <- max(stability_combined$lag_stable, na.rm = TRUE)

stability_ecdf <- stability_combined %>%
    mutate(
        threshold = factor(threshold, levels = threshold_levels)
    ) %>%
    group_by(threshold) %>%
    group_modify(function(df, key) {
        days <- seq(0, max_ecdf_day)
        tibble(
            lag_stable = days,
            cumulative_fraction = map_dbl(days, ~ mean(df$lag_stable <= .x, na.rm = TRUE))
        )
    }) %>%
    ungroup()

stability_wide <- stability_ecdf %>%
    pivot_wider(names_from = threshold, values_from = cumulative_fraction)

stability_bands <- bind_rows(
    stability_wide %>% transmute(lag_stable, band = "50% to 100%", ymin = `50%`, ymax = 1),
    stability_wide %>% transmute(lag_stable, band = "50% to 20%", ymin = `20%`, ymax = `50%`),
    stability_wide %>% transmute(lag_stable, band = "20% to 10%", ymin = `10%`, ymax = `20%`),
    stability_wide %>% transmute(lag_stable, band = "10% to 5%", ymin = `5%`, ymax = `10%`),
    stability_wide %>% transmute(lag_stable, band = "5% to 1%", ymin = `1%`, ymax = `5%`),
    stability_wide %>% transmute(lag_stable, band = "1% to 0.5%", ymin = `0.5%`, ymax = `1%`),
    stability_wide %>% transmute(lag_stable, band = "Within 0.5%", ymin = 0, ymax = `0.5%`)
) %>%
    mutate(
        lower = pmin(ymin, ymax, na.rm = TRUE),
        upper = pmax(ymin, ymax, na.rm = TRUE),
        ymin = lower,
        ymax = upper,
        band = factor(
            band,
            levels = c(
                "Within 0.5%", "1% to 0.5%", "5% to 1%",
                "10% to 5%", "20% to 10%", "50% to 20%",
                "50% to 100%"
            )
        )
    ) %>%
    select(-lower, -upper)

final_revision_ecdf <- stability_ecdf %>%
    filter(threshold == "Final revision")

focus_days <- 90
tail_start <- 500
early_day_breaks <- c(0, 7, 14, 30, 60, 90)
focus_day_breaks <- function(limits, endpoint = focus_days) {
    if (limits[1] <= 0 && limits[2] <= endpoint) {
        return(early_day_breaks[early_day_breaks <= endpoint])
    }
    breaks_pretty()(limits)
}
days_since_breaks <- function(limits) {
    candidate_breaks <- sort(unique(c(breaks_pretty()(limits), early_day_breaks)))
    candidate_breaks[candidate_breaks >= limits[1] & candidate_breaks <= limits[2]]
}
max_day <- max(stability_ecdf$lag_stable, na.rm = TRUE)
right_start <- if (max_day > tail_start) tail_start else focus_days

stability_plot <- function(x_limits, show_y = TRUE) {
    ggplot() +
        geom_ribbon(
            data = stability_bands,
            aes(x = lag_stable, ymin = ymin, ymax = ymax, fill = band),
            alpha = 0.52,
            show.legend = FALSE
        ) +
        geom_ribbon(
            data = final_revision_ecdf,
            aes(x = lag_stable, ymin = 0, ymax = cumulative_fraction),
            fill = final_revision_fill,
            alpha = 0.45,
            show.legend = FALSE
        ) +
        geom_line(
            data = stability_ecdf,
            aes(x = lag_stable, y = cumulative_fraction, color = threshold),
            linewidth = 0.9
        ) +
        geom_hline(yintercept = 1, color = "black", linewidth = 0.7) +
        scale_x_continuous(limits = x_limits, breaks = focus_day_breaks(x_limits), expand = expansion(mult = 0)) +
        scale_y_continuous(labels = label_percent(), limits = c(0, 1), expand = expansion(mult = 0)) +
        scale_color_manual(values = threshold_cols) +
        scale_fill_manual(values = band_cols) +
        guides(color = guide_legend(byrow = TRUE)) +
        labs(
            x = "Lag (days)",
            y = if (show_y) "Cumulative fraction of observations" else NULL,
            color = "Threshold"
        ) +
        theme(
            axis.title.y = if (show_y) element_text() else element_blank(),
            axis.text.y = if (show_y) element_text() else element_blank(),
            axis.ticks.y = if (show_y) element_line() else element_blank()
        )
}

make_cumulative_stabilization_plot <- function(data, title = NULL, show_legend = FALSE) {
    data <- data %>%
        filter(!is.na(lag_stable), !is.na(threshold)) %>%
        mutate(threshold = factor(threshold, levels = threshold_levels))

    if (nrow(data) == 0) {
        return(ggplot() + labs(title = title) + theme_void())
    }

    cumulative_data <- data %>%
        group_by(threshold) %>%
        group_modify(function(df, key) {
            tibble(
                lag_stable = seq(0, focus_days),
                cumulative_fraction = map_dbl(lag_stable, ~ mean(df$lag_stable <= .x, na.rm = TRUE))
            )
        }) %>%
        ungroup()

    cumulative_wide <- cumulative_data %>%
        pivot_wider(names_from = threshold, values_from = cumulative_fraction)

    cumulative_bands <- bind_rows(
        cumulative_wide %>% transmute(lag_stable, band = "50% to 100%", ymin = `50%`, ymax = 1),
        cumulative_wide %>% transmute(lag_stable, band = "50% to 20%", ymin = `20%`, ymax = `50%`),
        cumulative_wide %>% transmute(lag_stable, band = "20% to 10%", ymin = `10%`, ymax = `20%`),
        cumulative_wide %>% transmute(lag_stable, band = "10% to 5%", ymin = `5%`, ymax = `10%`),
        cumulative_wide %>% transmute(lag_stable, band = "5% to 1%", ymin = `1%`, ymax = `5%`),
        cumulative_wide %>% transmute(lag_stable, band = "1% to 0.5%", ymin = `0.5%`, ymax = `1%`),
        cumulative_wide %>% transmute(lag_stable, band = "Within 0.5%", ymin = 0, ymax = `0.5%`)
    ) %>%
        mutate(
            lower = pmin(ymin, ymax, na.rm = TRUE),
            upper = pmax(ymin, ymax, na.rm = TRUE),
            ymin = lower,
            ymax = upper,
            band = factor(
                band,
                levels = c(
                    "Within 0.5%", "1% to 0.5%", "5% to 1%",
                    "10% to 5%", "20% to 10%", "50% to 20%",
                    "50% to 100%"
                )
            )
        ) %>%
        select(-lower, -upper)

    cumulative_final <- cumulative_data %>%
        filter(threshold == "Final revision")

    ggplot() +
        geom_ribbon(
            data = cumulative_bands,
            aes(x = lag_stable, ymin = ymin, ymax = ymax, fill = band),
            alpha = 0.52,
            show.legend = FALSE
        ) +
        geom_ribbon(
            data = cumulative_final,
            aes(x = lag_stable, ymin = 0, ymax = cumulative_fraction),
            fill = final_revision_fill,
            alpha = 0.45,
            show.legend = FALSE
        ) +
        geom_line(
            data = cumulative_data,
            aes(x = lag_stable, y = cumulative_fraction, color = threshold),
            linewidth = 0.8
        ) +
        geom_hline(yintercept = 1, color = "black", linewidth = 0.5) +
        scale_x_continuous(limits = c(0, focus_days), breaks = focus_day_breaks(c(0, focus_days)), expand = expansion(mult = 0)) +
        scale_y_continuous(labels = label_percent(), limits = c(0, 1), expand = expansion(mult = 0)) +
        scale_color_manual(values = threshold_cols) +
        scale_fill_manual(values = band_cols) +
        guides(color = if (show_legend) guide_legend(byrow = TRUE) else "none") +
        labs(
            title = title,
            x = "Lag (days)",
            y = "Cumulative fraction",
            color = "Threshold"
        ) +
        theme(
            plot.title = element_text(size = 10, face = "bold"),
            axis.title = element_text(size = 8),
            axis.text = element_text(size = 7)
        )
}

plot_to_data_uri <- function(plot, width = 12, height = 8) {
    out_file <- tempfile(fileext = ".png")
    ggsave(out_file, plot = plot, width = width, height = height, dpi = 150, bg = "white")
    knitr::image_uri(out_file)
}

left_plot <- stability_plot(c(0, min(focus_days, max_day)), show_y = TRUE)
right_plot <- stability_plot(c(right_start, max_day), show_y = FALSE)

(left_plot + right_plot + plot_layout(widths = c(3, 1), guides = "collect")) +
    plot_annotation(
        title = "Cumulative stabilization by revision-distance threshold",
        subtitle = sprintf(
            "Axis break after %d days; right panel shows the long tail from day %d onward.",
            focus_days, right_start
        )
    )
```

The next view uses the same formula, but instead of asking when each observation
crosses a threshold, it shows how the distance to the latest value changes with
lag.

### Revision distance by lag

```{r}
#| label: convergence-overall-plot
#| fig-height: 6
overall_conv <- convergence_df %>%
    group_by(publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_obs = n(),
        .groups = "drop"
    ) %>%
    filter(n_obs > 5)

convergence_focus_days <- focus_days
max_conv_day <- max(overall_conv$publication_lag, na.rm = TRUE)
right_conv_start <- if (max_conv_day > tail_start) tail_start else convergence_focus_days

revision_distance_plot <- function(x_limits, show_y = TRUE) {
    ggplot(overall_conv, aes(x = publication_lag)) +
        geom_ribbon(
            aes(ymin = q25_rel_diff, ymax = q75_rel_diff),
            fill = dis_pal[1],
            alpha = 0.22
        ) +
        geom_line(aes(y = median_rel_diff), color = dis_pal[1], linewidth = 1) +
        scale_x_continuous(
            limits = x_limits,
            breaks = focus_day_breaks(x_limits, convergence_focus_days),
            expand = expansion(mult = 0)
        ) +
        scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
        labs(
            x = "Lag (days)",
            y = if (show_y) "Revision distance" else NULL
        ) +
        theme(
            axis.title.y = if (show_y) element_text() else element_blank(),
            axis.text.y = if (show_y) element_text() else element_blank(),
            axis.ticks.y = if (show_y) element_line() else element_blank()
        )
}

revision_distance_left <- revision_distance_plot(c(0, min(convergence_focus_days, max_conv_day)), show_y = TRUE)
revision_distance_right <- revision_distance_plot(c(right_conv_start, max_conv_day), show_y = FALSE)

(revision_distance_left + revision_distance_right + plot_layout(widths = c(3, 1))) +
    plot_annotation(
        title = "Revision distance by lag",
        subtitle = sprintf(
            "Axis break after %d days; right panel shows the long tail from day %d onward. Line: median; band: 25th-75th percentile range.",
            convergence_focus_days, right_conv_start
        )
    )
```

The table below gives the key readouts from the cumulative curves. For each
threshold, the columns answer how many days are needed before a given fraction
of observations have reached persistent stability.

```{r}
#| label: stability-summary-table
#| echo: false
stability_combined %>%
    group_by(threshold) %>%
    summarize(
        median = median(lag_stable, na.rm = TRUE),
        p75 = quantile(lag_stable, 0.75, na.rm = TRUE),
        p90 = quantile(lag_stable, 0.90, na.rm = TRUE),
        p95 = quantile(lag_stable, 0.95, na.rm = TRUE),
        p99 = quantile(lag_stable, 0.99, na.rm = TRUE),
        max = max(lag_stable, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(threshold = factor(threshold, levels = threshold_levels)) %>%
    arrange(threshold) %>%
    DT::datatable(
        colnames = c(
            "Threshold",
            "P50 days",
            "P75 days",
            "P90 days",
            "P95 days",
            "P99 days",
            "Max days"
        ),
        caption = "Key wait-time readouts from the cumulative stabilization curves",
        rownames = FALSE,
        options = list(dom = "t", paging = FALSE)
    ) %>%
    DT::formatRound(columns = 2:7, digits = 1)
```

### Stabilization time distribution by revision-distance threshold

This plot shows the shape of the wait-time distribution behind the cumulative
curves above. Bars show the empirical distribution within each threshold; the
smooth line is a density guide on the same approximate proportion scale.
This is the overall stabilization-time distribution. Later sections reuse the
same stabilization-time quantity, first split by location and then by
observation time.

```{r}
#| label: stability-histograms
#| fig-height: 8
stability_distribution <- stability_combined %>%
    filter(!is.na(lag_stable)) %>%
    mutate(threshold = factor(threshold, levels = threshold_levels))

distribution_bin_width <- 14
distribution_max_day <- ceiling(max(stability_distribution$lag_stable, na.rm = TRUE) / distribution_bin_width) *
    distribution_bin_width
distribution_bin_mids <- seq(
    distribution_bin_width / 2,
    distribution_max_day + distribution_bin_width / 2,
    by = distribution_bin_width
)

stability_distribution_bars <- stability_distribution %>%
    mutate(
        bin_start = floor(lag_stable / distribution_bin_width) * distribution_bin_width,
        bin_mid = bin_start + distribution_bin_width / 2
    ) %>%
    count(threshold, bin_mid, name = "n") %>%
    complete(threshold, bin_mid = distribution_bin_mids, fill = list(n = 0)) %>%
    left_join(count(stability_distribution, threshold, name = "threshold_total"), by = "threshold") %>%
    mutate(proportion = n / threshold_total)

stability_distribution_density <- stability_distribution %>%
    group_by(threshold) %>%
    group_modify(function(df, key) {
        if (n_distinct(df$lag_stable) < 2) {
            return(tibble(x = numeric(), y = numeric()))
        }
        dens <- density(
            df$lag_stable,
            from = 0,
            to = distribution_max_day,
            adjust = 1.1,
            n = 512
        )
        tibble(
            x = dens$x,
            y = dens$y * distribution_bin_width
        )
    }) %>%
    ungroup()

ggplot() +
    geom_col(
        data = stability_distribution_bars,
        aes(x = bin_mid, y = proportion, fill = threshold),
        width = distribution_bin_width * 0.95,
        color = "white",
        linewidth = 0.15,
        alpha = 0.72
    ) +
    geom_line(
        data = stability_distribution_density,
        aes(x = x, y = y, color = threshold),
        linewidth = 0.9,
        na.rm = TRUE
    ) +
    facet_wrap(~threshold, ncol = 1, scales = "free_y") +
    scale_x_continuous(breaks = days_since_breaks(c(0, distribution_max_day)), expand = expansion(mult = 0)) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    scale_fill_manual(values = threshold_cols, guide = "none") +
    scale_color_manual(values = threshold_cols) +
    guides(color = "none") +
    labs(
        title = "Stabilization time distribution by revision-distance threshold",
        subtitle = sprintf(
            "Bars show binned proportions; smoothed lines show density scaled to the same %d-day bin width.",
            distribution_bin_width
        ),
        x = "Lag at stabilization (days)",
        y = "Proportion of observations"
    )
```

---

## Is revision behavior different by location? {#by-location}

Revision behavior can vary by geography. This section first compares the total
amount of revision activity and published-version lag by location, then compares
how long each location takes to reach the stability thresholds used above.

```{r}
#| label: rev-beh-by-loc-compute
hhs_region_map <- tribble(
    ~geo_value, ~hhs_region,
    "ct", "HHS 1", "me", "HHS 1", "ma", "HHS 1", "nh", "HHS 1", "ri", "HHS 1", "vt", "HHS 1",
    "nj", "HHS 2", "ny", "HHS 2",
    "de", "HHS 3", "dc", "HHS 3", "md", "HHS 3", "pa", "HHS 3", "va", "HHS 3", "wv", "HHS 3",
    "al", "HHS 4", "fl", "HHS 4", "ga", "HHS 4", "ky", "HHS 4", "ms", "HHS 4", "nc", "HHS 4", "sc", "HHS 4", "tn", "HHS 4",
    "il", "HHS 5", "in", "HHS 5", "mi", "HHS 5", "mn", "HHS 5", "oh", "HHS 5", "wi", "HHS 5",
    "ar", "HHS 6", "la", "HHS 6", "nm", "HHS 6", "ok", "HHS 6", "tx", "HHS 6",
    "ia", "HHS 7", "ks", "HHS 7", "mo", "HHS 7", "ne", "HHS 7",
    "co", "HHS 8", "mt", "HHS 8", "nd", "HHS 8", "sd", "HHS 8", "ut", "HHS 8", "wy", "HHS 8",
    "az", "HHS 9", "ca", "HHS 9", "hi", "HHS 9", "nv", "HHS 9",
    "ak", "HHS 10", "id", "HHS 10", "or", "HHS 10", "wa", "HHS 10"
)

hhs_geo_order <- hhs_region_map %>%
    mutate(
        hhs_region = factor(hhs_region, levels = paste("HHS", 1:10)),
        geo_value = tolower(geo_value)
    ) %>%
    filter(geo_value %in% unique(tolower(stability_combined$geo_value))) %>%
    arrange(hhs_region, geo_value) %>%
    mutate(
        x_pos = row_number(),
        geo_label = toupper(geo_value),
        hhs_label = str_replace(as.character(hhs_region), "HHS ", "HHS\nRegion ")
    )

unmapped_geos <- setdiff(unique(tolower(stability_combined$geo_value)), hhs_geo_order$geo_value)
if (length(unmapped_geos) > 0) {
    hhs_geo_order <- bind_rows(
        hhs_geo_order,
        tibble(
            geo_value = sort(unmapped_geos),
            hhs_region = factor("Other", levels = c(paste("HHS", 1:10), "Other")),
            x_pos = seq(nrow(hhs_geo_order) + 1, nrow(hhs_geo_order) + length(unmapped_geos)),
            geo_label = toupper(sort(unmapped_geos)),
            hhs_label = "Other"
        )
    )
}

geo_sorting <- hhs_geo_order$geo_value

hhs_region_bounds <- hhs_geo_order %>%
    group_by(hhs_region, hhs_label) %>%
    summarize(
        xmin = min(x_pos),
        xmax = max(x_pos),
        xmid = mean(x_pos),
        .groups = "drop"
    )

hhs_bracket_plot <- function() {
    ggplot(hhs_region_bounds) +
        geom_segment(
            aes(x = xmin, xend = xmax, y = 0.72, yend = 0.72),
            color = "gray45",
            linewidth = 0.55
        ) +
        geom_segment(
            aes(x = xmin, xend = xmin, y = 0.72, yend = 0.95),
            color = "gray45",
            linewidth = 0.55
        ) +
        geom_segment(
            aes(x = xmax, xend = xmax, y = 0.72, yend = 0.95),
            color = "gray45",
            linewidth = 0.55
        ) +
        geom_text(
            aes(x = xmid, y = 0.22, label = hhs_label),
            size = 3.5,
            fontface = "bold",
            color = "gray20"
        ) +
        scale_x_continuous(
            limits = range(hhs_geo_order$x_pos) + c(-0.5, 0.5),
            expand = expansion(mult = c(0.01, 0.01))
        ) +
        coord_cartesian(ylim = c(0, 1), clip = "off") +
        labs(x = "Location ordered by HHS region") +
        theme_void() +
        theme(
            axis.title.x = element_text(size = 13, face = "bold", margin = margin(t = 14)),
            plot.margin = margin(-4, 5.5, 16, 5.5)
        )
}

rev_by_loc <- rev_beh %>%
    mutate(geo_value = tolower(geo_value)) %>%
    group_by(geo_value) %>%
    summarize(
        n_obs = n(),
        total_revisions = sum(n_revisions, na.rm = TRUE),
        avg_revisions = mean(n_revisions, na.rm = TRUE),
        median_window_days = median(as_days(max_lag), na.rm = TRUE),
        p90_window_days = quantile(as_days(max_lag), 0.9, na.rm = TRUE),
        avg_rel_spread = mean(rel_spread, na.rm = TRUE),
        median_conv_days = median(as_days(lag_near_latest), na.rm = TRUE),
        .groups = "drop"
    )
```

### Number of revisions per location

```{r}
#| label: revisions-by-location-barplot
#| fig-width: 12
#| fig-height: 5
#| eval: !expr n_locations > 3
revisions_by_location_plot <- rev_by_loc %>%
    inner_join(hhs_geo_order, by = "geo_value") %>%
    ggplot(aes(x = x_pos, y = total_revisions, fill = hhs_region)) +
    geom_col(width = 0.82, alpha = 0.88) +
    scale_x_continuous(
        breaks = hhs_geo_order$x_pos,
        labels = hhs_geo_order$geo_label,
        expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_y_continuous(labels = label_comma(), breaks = breaks_pretty()) +
    labs(
        title = "Number of revisions per location",
        subtitle = "Total revision events across all observation dates",
        x = NULL,
        y = "Number of revisions",
        fill = "HHS region"
    ) +
    guides(fill = "none") +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10, face = "bold"),
        legend.position = "none"
    )

revisions_by_location_plot / hhs_bracket_plot() + plot_layout(heights = c(12, 2.1))
```

The companion diagnostic below shows the distribution of `version - time_value`
for all states in the selected HHS region. The split between first published
values and subsequent versions helps separate initial publication timing from
later back-revisions.

```{r}
#| label: backfill-gap-compute
backfill_df <- archive$DT %>%
    filter(!is.na(value)) %>%
    group_by(geo_value, time_value) %>%
    mutate(
        backfill_gap = as.numeric(version - time_value, units = "days"),
        is_initial_release = version == min(version, na.rm = TRUE)
    ) %>%
    ungroup()
```

```{r}
#| label: backfill-gap-geography
#| echo: false
#| results: asis
lag_hhs_regions <- paste("HHS", 1:10)

lag_panel_html <- map_chr(lag_hhs_regions, function(region_name) {
    region_geos <- hhs_geo_order %>%
        filter(as.character(hhs_region) == region_name) %>%
        pull(geo_value)

    lag_plot <- backfill_df %>%
        mutate(geo_value = tolower(geo_value)) %>%
        filter(geo_value %in% region_geos) %>%
        mutate(
            geo_value = factor(geo_value, levels = region_geos, labels = toupper(region_geos))
        ) %>%
        ggplot(aes(x = backfill_gap, fill = is_initial_release)) +
        geom_histogram(bins = 40, position = "identity", alpha = 0.7) +
        facet_wrap(~geo_value, ncol = 2, scales = "free_y") +
        scale_fill_manual(
            values = c("TRUE" = dis_pal[2], "FALSE" = dis_pal[1]),
            labels = c("TRUE" = "First published value", "FALSE" = "Subsequent version"),
            name = "Version type"
        ) +
        scale_x_continuous(breaks = breaks_pretty()) +
        labs(
            title = sprintf("Published-version lag distribution: %s", region_name),
            subtitle = "First published value timing vs. subsequent revision depth",
            x = "Lag (days)",
            y = "Count"
        ) +
        theme(legend.position = "top", strip.text.x = element_text(size = 9))

    image_uri <- plot_to_data_uri(
        lag_plot,
        width = 12,
        height = max(4.5, 2.6 * ceiling(length(region_geos) / 2) + 1.2)
    )

    display <- ifelse(region_name == lag_hhs_regions[1], "block", "none")
    sprintf(
        '<div class="lag-hhs-panel" data-region="%s" style="display:%s;"><img src="%s" style="width:100%%;" alt="%s published-version lag by state"></div>',
        region_name, display, image_uri, region_name
    )
})

lag_control_html <- paste0(
    '<div class="lag-hhs-control">',
    '<button type="button" class="lag-page-step" data-direction="-1">Previous</button>',
    '<div class="lag-page-buttons">',
    paste0(
        '<button type="button" class="lag-page-button" data-region="',
        lag_hhs_regions,
        '" data-page="',
        seq_along(lag_hhs_regions),
        '">',
        seq_along(lag_hhs_regions),
        '</button>',
        collapse = ""
    ),
    '</div>',
    '<button type="button" class="lag-page-step" data-direction="1">Next</button>',
    '<span class="lag-page-status" aria-live="polite"></span>',
    '</div>'
)

lag_script_html <- paste(
    '<script>',
    '(function() {',
    '  const panels = document.querySelectorAll(".lag-hhs-panel");',
    '  const pageButtons = document.querySelectorAll(".lag-page-button");',
    '  const stepButtons = document.querySelectorAll(".lag-page-step");',
    '  const status = document.querySelector(".lag-page-status");',
    '  if (!panels.length || !pageButtons.length) return;',
    '  let currentPage = 0;',
    '  function updatePanels(pageIndex) {',
    '    currentPage = Math.max(0, Math.min(pageIndex, pageButtons.length - 1));',
    '    const region = pageButtons[currentPage].dataset.region;',
    '    panels.forEach(function(panel) {',
    '      panel.style.display = panel.dataset.region === region ? "block" : "none";',
    '    });',
    '    pageButtons.forEach(function(button, index) {',
    '      const isActive = index === currentPage;',
    '      button.classList.toggle("active", isActive);',
    '      button.setAttribute("aria-current", isActive ? "page" : "false");',
    '    });',
    '    stepButtons.forEach(function(button) {',
    '      const direction = Number(button.dataset.direction);',
    '      button.disabled = (direction < 0 && currentPage === 0) || (direction > 0 && currentPage === pageButtons.length - 1);',
    '    });',
    '    if (status) status.textContent = "Page " + (currentPage + 1) + " of " + pageButtons.length + ": " + region;',
    '  }',
    '  pageButtons.forEach(function(button, index) {',
    '    button.addEventListener("click", function() { updatePanels(index); });',
    '  });',
    '  stepButtons.forEach(function(button) {',
    '    button.addEventListener("click", function() {',
    '      updatePanels(currentPage + Number(button.dataset.direction));',
    '    });',
    '  });',
    '  updatePanels(0);',
    '})();',
    '</script>',
    sep = "\n"
)

lag_style_html <- paste(
    '<style>',
    '.lag-hhs-control {',
    '  display: flex;',
    '  align-items: center;',
    '  flex-wrap: wrap;',
    '  gap: 0.35rem;',
    '  margin: 0.75rem 0 1rem 0;',
    '  position: relative;',
    '  z-index: 10;',
    '}',
    '.lag-page-buttons {',
    '  display: flex;',
    '  flex-wrap: wrap;',
    '  gap: 0.25rem;',
    '}',
    '.lag-page-button, .lag-page-step {',
    '  border: 1px solid #d0d7de;',
    '  background: #fff;',
    '  color: #24292f;',
    '  border-radius: 4px;',
    '  padding: 0.25rem 0.55rem;',
    '  font-size: 0.9rem;',
    '  line-height: 1.2;',
    '  cursor: pointer;',
    '  pointer-events: auto;',
    '}',
    '.lag-page-button.active {',
    '  background: #0969da;',
    '  border-color: #0969da;',
    '  color: #fff;',
    '}',
    '.lag-page-step:disabled {',
    '  color: #8c959f;',
    '  background: #f6f8fa;',
    '  cursor: default;',
    '}',
    '.lag-page-status {',
    '  color: #57606a;',
    '  margin-left: 0.35rem;',
    '  font-size: 0.9rem;',
    '}',
    '</style>',
    sep = "\n"
)

cat(paste(
    lag_control_html,
    paste(lag_panel_html, collapse = "\n"),
    lag_script_html,
    lag_style_html,
    sep = "\n"
))
```

### Cumulative stabilization by revision-distance threshold by HHS region

The curve answers the same question as Section 2.1: as lag increases, what
fraction of observations have settled within each revision-distance threshold
and never left it? Here the same view is drawn separately for each state within
the selected HHS region.

```{r}
#| label: cumulative-stabilization-by-hhs-state
#| echo: false
#| results: asis
hhs_regions <- paste("HHS", 1:10)

hhs_panel_html <- map_chr(hhs_regions, function(region_name) {
    region_geos <- hhs_geo_order %>%
        filter(as.character(hhs_region) == region_name) %>%
        pull(geo_value)

    state_plots <- map(region_geos, function(geo) {
        make_cumulative_stabilization_plot(
            stability_combined %>% filter(tolower(geo_value) == geo),
            title = toupper(geo)
        )
    })

    region_plot <- wrap_plots(state_plots, ncol = 2) +
        plot_annotation(
            title = sprintf("Cumulative stabilization by revision-distance threshold: %s", region_name),
            subtitle = sprintf("Same view as Section 2.1, shown separately for each state. Lags 0-%d days.", focus_days)
        )

    image_uri <- plot_to_data_uri(
        region_plot,
        width = 12,
        height = max(4.5, 2.7 * ceiling(length(region_geos) / 2) + 1.2)
    )

    display <- ifelse(region_name == hhs_regions[1], "block", "none")
    sprintf(
        '<div class="hhs-cumulative-panel" data-region="%s" style="display:%s;"><img src="%s" style="width:100%%;" alt="%s cumulative stabilization by state"></div>',
        region_name, display, image_uri, region_name
    )
})

control_html <- paste0(
    '<div class="hhs-cumulative-control">',
    '<button type="button" class="hhs-page-step" data-direction="-1">Previous</button>',
    '<div class="hhs-page-buttons">',
    paste0(
        '<button type="button" class="hhs-page-button" data-region="',
        hhs_regions,
        '" data-page="',
        seq_along(hhs_regions),
        '">',
        seq_along(hhs_regions),
        '</button>',
        collapse = ""
    ),
    '</div>',
    '<button type="button" class="hhs-page-step" data-direction="1">Next</button>',
    '<span class="hhs-page-status" aria-live="polite"></span>',
    '</div>'
)

script_html <- paste(
    '<script>',
    '(function() {',
    '  const panels = document.querySelectorAll(".hhs-cumulative-panel");',
    '  const pageButtons = document.querySelectorAll(".hhs-page-button");',
    '  const stepButtons = document.querySelectorAll(".hhs-page-step");',
    '  const status = document.querySelector(".hhs-page-status");',
    '  if (!panels.length || !pageButtons.length) return;',
    '  let currentPage = 0;',
    '  function updatePanels(pageIndex) {',
    '    currentPage = Math.max(0, Math.min(pageIndex, pageButtons.length - 1));',
    '    const region = pageButtons[currentPage].dataset.region;',
    '    panels.forEach(function(panel) {',
    '      panel.style.display = panel.dataset.region === region ? "block" : "none";',
    '    });',
    '    pageButtons.forEach(function(button, index) {',
    '      const isActive = index === currentPage;',
    '      button.classList.toggle("active", isActive);',
    '      button.setAttribute("aria-current", isActive ? "page" : "false");',
    '    });',
    '    stepButtons.forEach(function(button) {',
    '      const direction = Number(button.dataset.direction);',
    '      button.disabled = (direction < 0 && currentPage === 0) || (direction > 0 && currentPage === pageButtons.length - 1);',
    '    });',
    '    if (status) status.textContent = "Page " + (currentPage + 1) + " of " + pageButtons.length + ": " + region;',
    '  }',
    '  pageButtons.forEach(function(button, index) {',
    '    button.addEventListener("click", function() { updatePanels(index); });',
    '  });',
    '  stepButtons.forEach(function(button) {',
    '    button.addEventListener("click", function() {',
    '      updatePanels(currentPage + Number(button.dataset.direction));',
    '    });',
    '  });',
    '  updatePanels(0);',
    '})();',
    '</script>',
    sep = "\n"
)

style_html <- paste(
    '<style>',
    '.hhs-cumulative-control {',
    '  display: flex;',
    '  align-items: center;',
    '  flex-wrap: wrap;',
    '  gap: 0.35rem;',
    '  margin: 0.75rem 0 1rem 0;',
    '  position: relative;',
    '  z-index: 10;',
    '}',
    '.hhs-page-buttons {',
    '  display: flex;',
    '  flex-wrap: wrap;',
    '  gap: 0.25rem;',
    '}',
    '.hhs-page-button, .hhs-page-step {',
    '  border: 1px solid #d0d7de;',
    '  background: #fff;',
    '  color: #24292f;',
    '  border-radius: 4px;',
    '  padding: 0.25rem 0.55rem;',
    '  font-size: 0.9rem;',
    '  line-height: 1.2;',
    '  cursor: pointer;',
    '  pointer-events: auto;',
    '}',
    '.hhs-page-button.active {',
    '  background: #0969da;',
    '  border-color: #0969da;',
    '  color: #fff;',
    '}',
    '.hhs-page-step:disabled {',
    '  color: #8c959f;',
    '  background: #f6f8fa;',
    '  cursor: default;',
    '}',
    '.hhs-page-status {',
    '  color: #57606a;',
    '  margin-left: 0.35rem;',
    '  font-size: 0.9rem;',
    '}',
    '</style>',
    sep = "\n"
)

cat(paste(
    control_html,
    paste(hhs_panel_html, collapse = "\n"),
    script_html,
    style_html,
    sep = "\n"
))
```

The next view uses the same revision-distance formula, but instead of asking
when each observation crosses a threshold, it shows how that distance changes
with lag within each HHS region.

### Revision distance by HHS region

```{r}
#| label: convergence-location-plot
#| fig-height: 18
region_conv <- convergence_df %>%
    mutate(geo_value = tolower(geo_value)) %>%
    inner_join(hhs_region_map, by = "geo_value") %>%
    group_by(hhs_region, geo_value, publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_obs = n(),
        .groups = "drop"
    ) %>%
    filter(n_obs > 1) %>%
    filter(publication_lag <= convergence_focus_days) %>%
    mutate(
        hhs_region = factor(hhs_region, levels = paste("HHS", 1:10)),
        geo_value = factor(geo_value, levels = sort(unique(geo_value)))
    )

plot_hhs_region <- function(region_name) {
    region_data <- region_conv %>%
        filter(hhs_region == region_name) %>%
        mutate(geo_value = droplevels(geo_value))

    state_levels <- levels(region_data$geo_value)
    state_cols <- setNames(hue_pal()(length(state_levels)), state_levels)
    state_labels <- setNames(toupper(state_levels), state_levels)

    ggplot(region_data, aes(x = publication_lag, group = geo_value)) +
        geom_ribbon(
            aes(ymin = q25_rel_diff, ymax = q75_rel_diff, fill = geo_value),
            alpha = 0.12,
            color = NA
        ) +
        geom_line(aes(y = median_rel_diff, color = geo_value), linewidth = 0.75, alpha = 0.95) +
        scale_x_continuous(
            limits = c(0, convergence_focus_days),
            breaks = focus_day_breaks(c(0, convergence_focus_days), convergence_focus_days),
            expand = expansion(mult = 0)
        ) +
        scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
        scale_color_manual(values = state_cols, labels = state_labels) +
        scale_fill_manual(values = state_cols, guide = "none") +
        labs(
            title = region_name,
            x = "Lag (days)",
            y = "Revision distance",
            color = NULL
        ) +
        theme(
            plot.title = element_text(size = 10, face = "bold"),
            legend.position = c(0.98, 0.98),
            legend.justification = c(1, 1),
            legend.background = element_rect(fill = scales::alpha("white", 0.78), color = "gray80"),
            legend.text = element_text(size = 8),
            legend.key.height = unit(0.65, "lines"),
            legend.key.width = unit(0.9, "lines")
        ) +
        guides(color = guide_legend(nrow = 2, byrow = TRUE))
}

region_plots <- lapply(paste("HHS", 1:10), plot_hhs_region)

wrap_plots(region_plots, ncol = 2) +
    plot_annotation(
        title = "Revision distance by HHS region",
        subtitle = sprintf(
            "Same definition as Section 2. Lags 0-%d days; lines show state-level medians and shaded bands show 25th-75th percentile ranges.",
            convergence_focus_days
        )
    )
```

### Stabilization time distribution by location

This is the location-level version of the stabilization-time distribution from
Section 2.2. Each point is the median time to reach a threshold for one
location; each vertical line is the 25th-75th percentile range.

```{r}
#| label: rev-by-location-maturation
#| fig-width: 12
#| fig-height: 7
#| eval: !expr n_locations > 3
location_maturation_df <- stability_combined %>%
    filter(!is.na(lag_stable), !is.na(threshold)) %>%
    mutate(geo_value = tolower(geo_value)) %>%
    inner_join(hhs_geo_order, by = "geo_value") %>%
    mutate(
        geo_value = factor(geo_value, levels = geo_sorting)
    ) %>%
    group_by(geo_value, x_pos, geo_label, threshold) %>%
    summarize(
        median_lag = median(lag_stable, na.rm = TRUE),
        q25_lag = quantile(lag_stable, 0.25, na.rm = TRUE),
        q75_lag = quantile(lag_stable, 0.75, na.rm = TRUE),
        .groups = "drop"
    )

location_maturation_plot <- ggplot(
    location_maturation_df,
    aes(x = x_pos, y = median_lag, ymin = q25_lag, ymax = q75_lag, color = threshold)
) +
    geom_pointrange(position = position_dodge(width = 0.7), linewidth = 0.45, size = 0.9) +
    scale_x_continuous(
        breaks = hhs_geo_order$x_pos,
        labels = hhs_geo_order$geo_label,
        expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_y_continuous(
        limits = c(0, NA),
        breaks = days_since_breaks(c(0, max(location_maturation_df$q75_lag, na.rm = TRUE)))
    ) +
    scale_color_manual(values = threshold_cols) +
    guides(color = guide_legend(byrow = TRUE)) +
    labs(
        title = "Stabilization time distribution by location",
        subtitle = "Lag needed to reach each stability threshold (dots = median, lines = IQR)",
        x = NULL, y = "Lag at stabilization (days)",
        color = "Threshold"
    ) +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10, face = "bold")
    )

location_maturation_plot / hhs_bracket_plot() + plot_layout(heights = c(12, 2.1))
```

---

## Are there patterns over observation time? {#time-patterns}

This section asks whether revision behavior changes across observation time. It
first summarizes the amount of revision activity over time, then repeats the
same two-part reading pattern used above: cumulative stabilization first, then
the revision-distance curve. The final plot turns back to stabilization time,
matching the quantity used in Sections 2.2 and 3.3.

Rather than using calendar years, these views use surveillance seasons. CDC
respiratory virus surveillance is commonly organized by MMWR weeks and named as
cross-year seasons. In this report, each season follows a flu-surveillance-style
MMWR season: MMWR week 40 through MMWR week 39 of the following year. Seasons are
labeled by the two calendar years they span (for example, `2022-2023`).

```{r}
#| label: season-definitions
#| echo: false
surveillance_season <- function(date) {
    date <- as.Date(date)
    mmwr_date <- MMWRweek(date)
    start_year <- if_else(mmwr_date$MMWRweek >= 40, mmwr_date$MMWRyear, mmwr_date$MMWRyear - 1L)
    paste0(start_year, "-", start_year + 1L)
}

season_levels <- stability_combined %>%
    mutate(obs_season = surveillance_season(time_value)) %>%
    distinct(obs_season) %>%
    arrange(obs_season) %>%
    pull(obs_season)

stability_season_df <- stability_combined %>%
    mutate(obs_season = factor(surveillance_season(time_value), levels = season_levels))

convergence_season_df <- convergence_df %>%
    mutate(obs_season = factor(surveillance_season(time_value), levels = season_levels))
```

### Number of revisions over observation time

```{r}
#| label: revision-count-plot
#| echo: false
#| fig-height: 4.5
rev_count_df <- archive$DT %>%
    group_by(geo_value, time_value) %>%
    summarize(n_revisions = n_distinct(version) - 1L, .groups = "drop")

rev_count_summary <- rev_count_df %>%
    group_by(time_value) %>%
    summarize(
        q25_revisions = quantile(n_revisions, 0.25, na.rm = TRUE),
        median_revisions = median(n_revisions, na.rm = TRUE),
        q75_revisions = quantile(n_revisions, 0.75, na.rm = TRUE),
        n_locations = n_distinct(geo_value),
        .groups = "drop"
    )

ggplot(rev_count_summary, aes(x = time_value)) +
    geom_ribbon(
        aes(ymin = q25_revisions, ymax = q75_revisions, fill = "25th-75th percentile"),
        alpha = 0.22,
        color = NA
    ) +
    geom_line(aes(y = median_revisions, color = "Median"), linewidth = 0.9) +
    scale_color_manual(values = c("Median" = "steelblue4")) +
    scale_fill_manual(values = c("25th-75th percentile" = "steelblue")) +
    scale_x_date(breaks = breaks_pretty()) +
    scale_y_continuous(breaks = breaks_pretty()) +
    labs(
        title = "Number of revisions by reference date",
        subtitle = "Line shows the median across locations; shaded band shows the 25th-75th percentile range.",
        x = "Reference date",
        y = "Number of revisions per location",
        color = NULL,
        fill = NULL
    )
```

The companion diagnostic below shows published-version lag by surveillance
season. Each panel is one MMWR week 40-39 season, allowing the seasonal timing pattern
to be compared directly.

```{r}
#| label: published-version-lag-by-season
#| fig-height: 7
backfill_df %>%
    mutate(obs_season = factor(surveillance_season(time_value), levels = season_levels)) %>%
    filter(!is.na(obs_season)) %>%
    ggplot(aes(x = backfill_gap, fill = is_initial_release)) +
    geom_histogram(bins = 40, position = "identity", alpha = 0.7, color = "white") +
    facet_wrap(~obs_season, ncol = 2, scales = "free_y") +
    scale_fill_manual(
        values = c("TRUE" = dis_pal[2], "FALSE" = dis_pal[1]),
        labels = c("TRUE" = "First published value", "FALSE" = "Subsequent version"),
        name = "Version type"
    ) +
    scale_x_continuous(breaks = breaks_pretty()) +
    labs(
        title = "Published-version lag distribution by observation season",
        subtitle = "MMWR week 40-39 seasons; first published value timing vs. subsequent revision depth",
        x = "Lag (days)",
        y = "Count"
    ) +
    theme(strip.text.x = element_text(size = 12, face = "bold"))
```

### Cumulative stabilization by revision-distance threshold and observation season

The curve answers the same question as Section 2.1: as lag increases, what
fraction of observations have settled within each revision-distance threshold
and never left it? Here the same view is drawn separately for each observation
season.

```{r}
#| label: cumulative-stabilization-by-season
#| fig-width: 12
#| fig-height: 7
season_stabilization_plots <- map(season_levels, function(season_value) {
    make_cumulative_stabilization_plot(
        stability_season_df %>% filter(obs_season == season_value),
        title = season_value
    )
})

wrap_plots(season_stabilization_plots, ncol = 2) +
    plot_annotation(
        title = "Cumulative stabilization by revision-distance threshold and observation season",
        subtitle = sprintf(
            "Same view as Section 2.1, shown separately for each MMWR week 40-39 surveillance season. Lags 0-%d days.",
            focus_days
        )
    )
```

The next view uses the same revision-distance formula, but instead of asking
when each observation crosses a threshold, it shows how that distance changes
with lag across observation seasons.

### Revision distance by observation season

```{r}
#| label: revision-distance-by-season
#| fig-height: 5
season_conv <- convergence_season_df %>%
    group_by(obs_season, publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_obs = n(),
        .groups = "drop"
    ) %>%
    filter(n_obs > 5) %>%
    filter(publication_lag <= convergence_focus_days)

ggplot(season_conv, aes(x = publication_lag, color = obs_season, fill = obs_season)) +
    geom_ribbon(
        aes(ymin = q25_rel_diff, ymax = q75_rel_diff),
        alpha = 0.14,
        color = NA
    ) +
    geom_line(aes(y = median_rel_diff), linewidth = 0.9) +
    scale_x_continuous(
        limits = c(0, convergence_focus_days),
        breaks = focus_day_breaks(c(0, convergence_focus_days), convergence_focus_days),
        expand = expansion(mult = 0)
    ) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    labs(
        title = "Revision distance by observation season",
        subtitle = sprintf(
            "Same revision-distance definition as Sections 2.1 and 3.2. Seasons run MMWR week 40-39; lags 0-%d days. Lines show medians and shaded bands show 25th-75th percentile ranges.",
            convergence_focus_days
        ),
        x = "Lag (days)",
        y = "Revision distance",
        color = "Observation season",
        fill = "Observation season"
    )
```

### Stabilization time distribution over observation time

This is the time-level version of the stabilization-time distribution from
Sections 2.2 and 3.3. Instead of showing the full distribution or splitting by
location, it asks whether the time to reach each threshold changes across
observation periods. Lines show the median stabilization time, and shaded bands
show the 25th-75th percentile range. Recent observations may be censored because
there has not yet been enough time to observe their full revision history.
Apparent jumps or short discontinuities can occur when the available observations
within a time bucket become sparse or compositionally different near the
censoring boundary; these are data-support effects rather than an additional
axis break.

```{r}
#| label: diagnostic-censoring
#| fig-height: 4.5
max_v <- max(archive$DT$version)

date_range_days <- as.numeric(max(stability_combined$time_value) - min(stability_combined$time_value))
agg_unit <- case_when(
    date_range_days > 365 * 2 ~ "quarter",
    date_range_days > 365 ~ "month",
    date_range_days > 100 ~ "2-week",
    date_range_days > 40 ~ "week",
    date_range_days > 20 ~ "3-day",
    TRUE ~ "day"
)

floor_time_bucket <- function(date, unit) {
    date <- as.Date(date)
    if (unit == "2-week") {
        origin <- min(date, na.rm = TRUE)
        return(origin + 14 * floor(as.numeric(date - origin) / 14))
    }
    if (unit == "3-day") {
        origin <- min(date, na.rm = TRUE)
        return(origin + 3 * floor(as.numeric(date - origin) / 3))
    }
    floor_date(date, unit = unit)
}

stability_time_summary <- stability_combined %>%
    filter(!is.na(lag_stable), !is.na(threshold)) %>%
    mutate(
        threshold = factor(threshold, levels = threshold_levels),
        time_bucket = floor_time_bucket(time_value, agg_unit)
    ) %>%
    group_by(threshold, time_bucket) %>%
    summarize(
        median_lag = median(lag_stable, na.rm = TRUE),
        q25_lag = quantile(lag_stable, 0.25, na.rm = TRUE),
        q75_lag = quantile(lag_stable, 0.75, na.rm = TRUE),
        .groups = "drop"
    )

ggplot(stability_time_summary, aes(x = time_bucket, color = threshold, fill = threshold)) +
    geom_ribbon(
        aes(ymin = q25_lag, ymax = q75_lag),
        alpha = 0.22,
        color = NA
    ) +
    geom_line(aes(y = median_lag), linewidth = 0.85) +
    geom_function(
        fun = function(x) as.numeric(max_v - as.Date(x, origin = "1970-01-01")),
        linetype = "dashed", color = "gray50", linewidth = 0.5
    ) +
    annotate(
        "text",
        x = min(stability_time_summary$time_bucket),
        y = as_days(max_v - min(stability_combined$time_value)),
        label = "Censoring Boundary", color = "gray50",
        vjust = 1, hjust = -1, size = 3
    ) +
    scale_x_date(breaks = breaks_pretty()) +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty()) +
    scale_color_manual(values = threshold_cols) +
    scale_fill_manual(values = threshold_cols) +
    labs(
        title = "Stabilization time distribution over observation time",
        subtitle = sprintf(
            "Reference dates grouped by %s. Lines show medians; shaded bands show the 25th-75th percentile range.",
            agg_unit
        ),
        x = "Reference date",
        y = "Lag at stabilization (days)",
        color = "Threshold",
        fill = "Threshold"
    )
```

---

## Are revision patterns related to other signal features? {#signal-features}

After checking first-release behavior, maturation time, geography, and
surveillance season, this section asks whether other observable features of the
signal are related to revision behavior. These views are exploratory: they help
identify where revision patterns may be driven by signal magnitude, initial
publication lag, or finer reference-date structure.

### Signal magnitude vs. relative spread

This plot asks whether low or high signal values tend to have larger relative
revisions.

```{r}
#| label: diagnostic-magnitude
#| fig-height: 4.5
rev_beh %>%
    filter(rel_spread > 0) %>%
    ggplot(aes(x = max_value, y = rel_spread)) +
    geom_point(alpha = 0.2, size = 1) +
    geom_smooth() +
    scale_x_log10(labels = label_comma()) +
    scale_y_log10(labels = label_percent()) +
    labs(
        title = "Signal magnitude vs. relative spread",
        subtitle = "Log-log scale",
        x = "Maximum observed value (log scale)", y = "Relative spread (log scale)"
    )
```

### Revision distance heatmap

This heatmap shows the revision-distance surface over daily reference date and
lag, split by surveillance season. Each cell summarizes the median revision
distance for observations with the same reference date and lag.

```{r}
#| label: revision-distance-heatmap
#| fig-height: 10
season_start_date <- function(season_label) {
    start_year <- as.integer(str_extract(as.character(season_label), "^\\d{4}"))
    as.Date(sprintf("%d-10-01", start_year))
}

season_month_starts <- seq(as.Date("2000-10-01"), by = "month", length.out = 12)
season_month_breaks <- as.numeric(season_month_starts - season_month_starts[1])
season_month_labels <- format(season_month_starts, "%m")
season_axis_end <- as.numeric(as.Date("2001-10-01") - as.Date("2000-10-01"))

heatmap_df <- convergence_season_df %>%
    filter(publication_lag >= 0, publication_lag <= focus_days) %>%
    filter(!is.na(obs_season)) %>%
    mutate(
        season_start = season_start_date(obs_season),
        season_day = as.numeric(time_value - season_start)
    ) %>%
    filter(season_day >= 0, season_day < season_axis_end) %>%
    group_by(obs_season, season_day, publication_lag) %>%
    summarize(
        median_revision_distance = median(abs_rel_diff, na.rm = TRUE),
        n_obs = n(),
        .groups = "drop"
    )

heatmap_cap <- max(
    quantile(heatmap_df$median_revision_distance, 0.95, na.rm = TRUE),
    0.01,
    na.rm = TRUE
)
heatmap_step <- 0.05
heatmap_upper <- max(0.20, ceiling(heatmap_cap / heatmap_step) * heatmap_step)
heatmap_breaks <- seq(0, heatmap_upper, by = heatmap_step)
heatmap_cols <- colorRampPalette(c(
    "#f2f2f2",
    "#bcbddc",
    "#98df8a",
    "#ffbb78",
    "#1f77b4"
)
)(length(heatmap_breaks) - 1)

ggplot(heatmap_df, aes(x = season_day, y = publication_lag, fill = median_revision_distance)) +
    geom_tile(width = 1, height = 1) +
    facet_wrap(~obs_season, ncol = 2) +
    scale_x_continuous(
        breaks = season_month_breaks,
        labels = season_month_labels,
        limits = c(0, season_axis_end),
        expand = expansion(mult = 0)
    ) +
    scale_y_continuous(
        breaks = focus_day_breaks(c(0, focus_days), focus_days),
        limits = c(0, focus_days),
        expand = expansion(mult = 0)
    ) +
    scale_fill_stepsn(
        colors = heatmap_cols,
        breaks = heatmap_breaks,
        labels = label_percent(accuracy = 1),
        limits = c(0, heatmap_upper),
        oob = squish,
        guide = guide_colorsteps(
            direction = "horizontal",
            even.steps = FALSE,
            barheight = grid::unit(4, "mm"),
            barwidth = grid::unit(110, "mm"),
            title.position = "top",
            label.position = "bottom"
        )
    ) +
    labs(
        title = "Revision distance heatmap",
        subtitle = sprintf(
            "Daily reference dates aligned within surveillance season; cells show median revision distance. Color bins use 5 percentage-point steps; values above %s are capped.",
            label_percent(accuracy = 1)(heatmap_upper)
        ),
        x = "Reference month",
        y = "Lag (days)",
        fill = "Revision distance"
    ) +
    theme(
        legend.position = "bottom",
        legend.box = "vertical",
        legend.title = element_text(hjust = 0.5)
    )
```

### Revision distance by initial lag

Early publication can be useful, but it may also leave more room for later
revision. This view groups observations by the first observed lag for each
reference date, then compares how revision distance changes across lags.

```{r}
#| label: revision-distance-by-initial-lag
#| fig-height: 5
initial_lag_df <- convergence_df %>%
    group_by(geo_value, time_value) %>%
    summarize(
        initial_publication_lag = min(publication_lag, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(
        initial_lag_group = cut(
            initial_publication_lag,
            breaks = c(-Inf, 7, 14, 30, Inf),
            labels = c("<=7 days", "8-14 days", "15-30 days", ">30 days"),
            right = TRUE
        )
    )

initial_lag_revision_df <- convergence_df %>%
    inner_join(
        initial_lag_df %>% select(geo_value, time_value, initial_lag_group),
        by = c("geo_value", "time_value")
    ) %>%
    filter(!is.na(initial_lag_group), publication_lag <= convergence_focus_days) %>%
    group_by(initial_lag_group, publication_lag) %>%
    summarize(
        q25_rel_diff = quantile(abs_rel_diff, 0.25, na.rm = TRUE),
        median_rel_diff = median(abs_rel_diff, na.rm = TRUE),
        q75_rel_diff = quantile(abs_rel_diff, 0.75, na.rm = TRUE),
        n_obs = n(),
        .groups = "drop"
    ) %>%
    filter(n_obs > 5)

ggplot(initial_lag_revision_df, aes(x = publication_lag, color = initial_lag_group, fill = initial_lag_group)) +
    geom_ribbon(aes(ymin = q25_rel_diff, ymax = q75_rel_diff), alpha = 0.14, color = NA) +
    geom_line(aes(y = median_rel_diff), linewidth = 0.95) +
    scale_x_continuous(
        limits = c(0, convergence_focus_days),
        breaks = focus_day_breaks(c(0, convergence_focus_days), convergence_focus_days),
        expand = expansion(mult = 0)
    ) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    labs(
        title = "Revision distance by initial lag",
        subtitle = "Groups are based on the first observed lag for each reference date.",
        x = "Lag (days)",
        y = "Revision distance",
        color = "Initial lag",
        fill = "Initial lag"
    )
```

## Summary {#summary}

```{r}
#| label: final-summary-narrative
#| echo: false
#| results: asis
# Statistical context for the narrative
final_rev_data <- stability_combined %>% filter(threshold == "Final revision")
med_window <- round(median(final_rev_data$lag_stable, na.rm = TRUE))
p90_window <- round(quantile(final_rev_data$lag_stable, 0.9, na.rm = TRUE))

p_no_rev <- label_percent()(no_rev_pct)
pct_under <- label_percent()(mean(bias_by_loc$median_bias < 0))

cat(sprintf(
    "For **%s** (`%s / %s`), the version history covering %s to %s\nshows the following revision and backfill profile:\n\n",
    params$signal_name, params$source, params$signal,
    params$start_day, params$end_day
))

cat(sprintf("- **%s** of observations were never revised after their first published value.\n", p_no_rev))
cat(sprintf(
    "- The **median lag to persistent stability is %d days**: half of all revised ",
    med_window
))
cat(sprintf(
    "observations reached their latest value by lag %d.\n",
    med_window
))
cat(sprintf(
    "- The 90th percentile for maturity is **%d days** (~%d weeks).\n",
    p90_window, ceiling(p90_window / 7)
))
cat(sprintf(
    "- **%s of locations** exhibit systematic under-reporting in their first published value.\n",
    pct_under
))

# Threshold maturation summary
stab_summary_meds <- stability_combined %>%
    group_by(threshold) %>%
    summarize(median = round(median(lag_stable, na.rm = TRUE)), .groups = "drop")

cat(sprintf("- **Median days to reach stability**: "))
cat(paste0(
    "**", stab_summary_meds$threshold, "**: ", stab_summary_meds$median, " days",
    collapse = " | "
))
cat(".\n\n")
```
