---
title: "Indicator Exploratory Analysis: `r paste(if (!is.null(params$name)) params$name else params$signal, collapse = ' | ')`"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
params:
  # Candidate Indicator
  source:
  signal:
  name:
  input_csv: !expr "NULL"
  # Shared settings
  geo_type:
  time_type:
  start_day:
  end_day:
  # Maximum locations to show in plots/tables
  max_locations_plot: 60
  max_archive_locs: 60
knitr:
  opts_chunk:
    comment: "#>"
    message: FALSE
    warning: FALSE
    out.width: "100%"
    dev: "png"
    fig.width: 8
    fig.height: 4.5
    fig.align: center
editor_options: 
  chunk_output_type: console
execute:
  cache: false
---

```{r}
#| label: setup
#| include: false
#| cache: false
# When running interactively, populate params from defaults
if (!exists("params")) {
    params <- list(
        source = "beta_pophive",
        signal = "epic_pct_covid_total",
        name = "PopHive: Epic % COVID Total",
        input_csv = NULL,
        # Shared settings,
        geo_type = "state",
        time_type = "week",
        start_day = "2022-01-01",
        end_day = "2025-12-31",
        max_locations_plot = 60,
        max_archive_locs = 60
    )
}

# Load libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    epidatr, epiprocess,
    tidyverse, scales, patchwork,
    here, glue, ggtext
)
# Custom HTML-friendly label wrapper for ggtext::element_markdown
label_wrap_html <- function(width = 25) {
  fun <- label_wrap_gen(width = width)
  function(labels) {
    res <- fun(labels)
    res[] <- lapply(res, function(x) {
      if (is.character(x) || is.factor(x)) {
        gsub("\n", "<br>", as.character(x), fixed = TRUE)
      } else {
        x
      }
    })
    res
  }
}

# Invalidate knitr cache when params change
knitr::opts_chunk$set(cache.extra = list(params))

# Normalize signal/name params to unique character vectors
signals_vec  <- unique(as.character(params$signal))
multi_signal <- length(signals_vec) > 1

# Build signal label 
name_vec <- {
    n <- params$name
    if (!is.null(n) && length(n) == length(signals_vec)) as.character(n)
    else signals_vec
}

# Wrap long names
name_vec <- str_wrap(name_vec, width = 15)

signal_labels <- setNames(name_vec, signals_vec)

# Set theme and colors
colours <- c(
    "#172431",
    "#ffffff",
    "#C41230",
    "#941120",
    "#007BC0",
    "#002676",
    "#FDB515",
    "#002145",
    "#00A7E1",
    "#8C338F",
    "#368f33"
)

# Palette selection
dis_pal_list <- colours[c(3, 5, 7, 10, 11, 9)]

theme_set(
    theme_minimal(base_family = "Lato", base_size = 12) +
        theme(
            plot.title = ggtext::element_textbox_simple(hjust = 0.5, halign = 0.5, size = 15, face = "bold", margin = margin(b = 6)),
            plot.subtitle = element_text(hjust = 0.5, color = "gray50", margin = margin(b = 10)),
            plot.caption = ggtext::element_textbox_simple(hjust = 0, halign = 0, color = "gray50", margin = margin(t = 10)),
            panel.grid = element_line(linetype = 2, linewidth = 0.3, color = "gray90"),
            panel.grid.minor = element_blank(),
            legend.position = "top",
            legend.text = ggtext::element_markdown(size = 10),
            legend.title = ggtext::element_markdown(size = 10, face = "bold"),
            strip.text = ggtext::element_markdown(hjust = 0.5, size = 11, face = "bold", margin = margin(t = 4, b = 4)),
            strip.background = element_rect(fill = "gray95", linetype = "blank"),
            panel.border = element_rect(color = "gray95", fill = NA)
        )
)

# Palette as global defaults for discrete and continuous scales
options(
    ggplot2.discrete.colour = dis_pal_list,
    ggplot2.discrete.fill = dis_pal_list,
    ggplot2.continuous.colour = function() scale_colour_gradientn(colours = dis_pal_list),
    ggplot2.continuous.fill = function() scale_fill_gradientn(colours = dis_pal_list)
)

# Update defaults for geoms/stats
tibble::tribble(
    ~name,      ~aes,     ~alpha, ~is_stat,
    "line",     "colour", 1.0,    FALSE,
    "point",    "colour", 0.7,    FALSE,
    "function", "colour", 1.0,    FALSE,
    "bar",      "fill",   0.9,    FALSE,
    "col",      "fill",   0.9,    FALSE,
    "boxplot",  "fill",   0.8,    FALSE,
    "bin",      "fill",   0.8,    TRUE,
    "identity", "fill",   0.7,    TRUE
) %>%
    purrr::pwalk(function(name, aes, alpha, is_stat) {
        updater <- if (is_stat) update_stat_defaults else update_geom_defaults
        updater(name, setNames(list(dis_pal_list[1], alpha), c(aes, "alpha")))
    })

# Helper: round any date to the immediately preceding Sunday (or same day if already Sunday)
round_to_sunday <- function(d) d - as.integer(format(d, "%w"))

# Helper: build epirange with correct format for the indicator's time_type
time_range <- function(start, end, time_type = params$time_type) {
    if (time_type == "week") {
        to_yw <- function(d) {
            d <- as.Date(d)
            m <- MMWRweek::MMWRweek(d)
            m$MMWRyear * 100L + m$MMWRweek
        }
        epirange(to_yw(start), to_yw(end))
    } else {
        epirange(start, end)
    }
}

# Helper: map geo codes to readable names
geo_to_name <- function(geos, geo_type = params$geo_type) {
    if (geo_type == "state") {
        names <- covidcast::abbr_to_name(toupper(geos))
    } else if (geo_type == "county") {
        names <- covidcast::county_fips_to_name(as.character(geos))
    } else if (geo_type == "msa") {
        names <- covidcast::cbsa_to_name(as.character(geos))
    } else {
        return(geos)
    }
    # Fallback to original codes for any missing mappings
    names[is.na(names)] <- geos[is.na(names)]
    return(names)
}

# Helper: consistent random sample from a vector
random_sample <- function(vec, n = 15) {
    if (length(vec) <= n) {
        return(vec)
    }
    withr::with_seed(42, sort(sample(vec, n)))
}

# Helper function to print list (up to 15 obs)
format_excluded <- function(locs, names) {
    if (length(locs) == 0) {
        return(NULL)
    }
    sampled_idx <- match(random_sample(locs), locs)
    sampled_locs <- locs[sampled_idx]
    sampled_names <- names[sampled_idx]
    entries <- ifelse(
        sampled_locs == sampled_names,
        sampled_locs,
        sprintf("%s (%s)", sampled_locs, sampled_names)
    )
    if (length(locs) > 15) {
        paste0(paste(entries, collapse = ", "), ", etc.")
    } else {
        paste(entries, collapse = ", ")
    }
}

# Set cache
cache_path <- file.path(getwd(), "epidatr_cache")
if (!dir.exists(cache_path)) dir.create(cache_path, recursive = TRUE)

set_cache(cache_dir = cache_path)
```

This notebook evaluates a **candidate indicator** to understand its characteristics, coverage, missingness, and versioning/revision behavior.

```{r}
#| label: print-config
#| echo: false
config_tbl <- tibble(
    Field = c("Source", "Signal(s)", "Label(s)", "Geography", "Period"),
    Value = c(
        params$source,
        paste(signals_vec, collapse = ", "),
        paste(str_remove(name_vec, "\\n"), collapse = ", "),
        params$geo_type,
        paste(params$start_day, "to", params$end_day)
    )
)

DT::datatable(
    config_tbl,
    caption = "Analysis configuration",
    rownames = FALSE,
    options = list(scrollX = TRUE, dom = "t", paging = FALSE)
)
```

---

## Data Loading {#data}

```{r}
#| label: load-candidate
#| results: asis
# Load candidate indicator
candidate_archive <- NULL
if (!is.null(params$input_csv)) {
    df_candidate_raw <- read.csv(params$input_csv) %>%
        mutate(time_value = as.Date(time_value)) %>%
        filter(time_value >= as.Date(params$start_day) & time_value <= as.Date(params$end_day))

    # CSV may include version/issue columns for revision analysis
    has_versions <- any(c("version", "issue") %in% colnames(df_candidate_raw))
    if (has_versions) {
        df_candidate_raw <- df_candidate_raw %>%
            {
                if ("issue" %in% names(.)) rename(., version = issue) else .
            } %>%
            mutate(version = as.Date(version))
        csv_archive_df  <- df_candidate_raw %>%
            select(geo_value, time_value, version, any_of("signal"), value) %>%
            group_by(geo_value, time_value, version, pick(any_of("signal"))) %>%
            slice_max(value, n = 1, with_ties = FALSE) %>%
            ungroup()

        candidate_archive <- csv_archive_df %>%
            as_epi_archive(signal_format = "long")
        df_candidate <- epix_as_of(candidate_archive, candidate_archive$versions_end)
        cat(sprintf(
            "- CSV contains version history: %s versions\n",
            label_comma()(n_distinct(df_candidate_raw$version))
        ))
    } else {
        df_candidate <- df_candidate_raw %>%
            as_epi_df(signal_format = "long")
    }
} else {
    # Determine if new API can be used for candidate source
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE
    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                all(signals_vec %in% meta_res[[params$source]]$signals)
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }
    if (use_new_api) {
        # epidata_snapshot does not accept multiple signals
        df_candidate <- purrr::map(signals_vec, ~ {
            epidata_snapshot(
                source = params$source,
                signals = .x,
                geo_type = params$geo_type
            ) %>%
                rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
                {
                    if ("fill_method" %in% colnames(.)) {
                        filter(., fill_method == "source" | !any(fill_method == "source"))
                    } else .
                } %>%
                # {
                #     if ("version" %in% colnames(.)) {
                #         group_by(., geo_value, time_value) %>%
                #             slice_max(version, n = 1, with_ties = FALSE) %>%
                #             ungroup()
                #     } else .
                # } %>%
                filter(
                    time_value >= as.Date(params$start_day),
                    time_value <= as.Date(params$end_day)
                ) %>%
                select(geo_value, any_of("signal"), time_value, value)
        }) %>%
            bind_rows() %>%
            as_epi_df(signal_format = "long")
    } else {
        # pub_covidcast accepts a signal vector and returns a signal column
        df_candidate <- pub_covidcast(
            params$source, signals_vec,
            params$geo_type, params$time_type,
            time_values = time_range(params$start_day, params$end_day),
            as_of = Sys.Date()
        ) %>% as_epi_df(signal_format = "long")
    }
}

cat(sprintf(
    "- **Summary**: %s rows across %s locations and %d signal(s)\n",
    label_comma()(nrow(df_candidate)),
    label_comma()(n_distinct(df_candidate$geo_value)),
    length(signals_vec)
))
```

```{r}
#| label: load-candidate-archive
#| results: asis
# Fetch revision history from API if not already built from CSV
if (is.null(candidate_archive) && is.null(params$input_csv)) {
    # Only fetch version history if the source supports revisions
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE
    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                all(signals_vec %in% meta_res[[params$source]]$signals)
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }

    # Fetch archive for revision behavior
    archive_geos <- unique(df_candidate$geo_value)
    # Limit number of locations for version history queries to avoid API limits
    if (length(archive_geos) > params$max_locations_plot) {
        archive_geos <- random_sample(archive_geos, params$max_locations_plot)
    }

    # Determine window of issues available
    archive_start <- as.Date(params$start_day) - 60

    archive_info <- list(
        n_sampled = length(archive_geos),
        n_total = length(unique(df_candidate$geo_value)),
        time_value_start = archive_start,
        time_value_end = params$end_day
    )

    if (use_new_api) {
        # epidata_archive does not accept multiple signals
        df_archive_raw <- purrr::map(signals_vec, ~ {
            epidata_archive(
                source = params$source,
                signals = .x,
                geo_type = params$geo_type
            ) %>%
                rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
                filter(
                    geo_value %in% archive_geos,
                    time_value >= as.Date(archive_start),
                    time_value <= as.Date(params$end_day)
                )
        }) %>% bind_rows()
    } else {
        # pub_covidcast accepts a signal vector and returns a signal column
        df_archive_raw <- pub_covidcast(
            source = params$source,
            signal = signals_vec,
            geo_type = params$geo_type,
            time_type = params$time_type,
            geo_values = archive_geos,
            time_values = time_range(archive_start, params$end_day),
            issues = "*",
        )
    }

    if (nrow(df_archive_raw) > 0) {
        df_archive_raw <- df_archive_raw %>%
            rename(any_of(c(time_value = "reference_time", version = "issue"))) %>%
            rename(any_of(c(version = "report_time"))) %>%
            select(geo_value, time_value, version, any_of("signal"), value) #%>%
            # Deduplicate: sources like NWSS can have multiple rows per
            # (geo, time, version) from different measurement sources.
            # group_by(geo_value, time_value, version, across(any_of("signal"))) %>%
            # slice_max(value, n = 1, with_ties = FALSE) %>%
            # ungroup()

        if ("version" %in% colnames(df_archive_raw)) {
            archive_info$version_start <- min(df_archive_raw$version)
            archive_info$version_end <- max(df_archive_raw$version)
            candidate_archive <- df_archive_raw %>%
                as_epi_archive(signal_format = "long")
        }
    }
}

# Determine time type and number of locations
env_time_type <- attributes(df_candidate)$metadata$time_type
if (is.null(env_time_type) || length(env_time_type) != 1 || is.na(env_time_type) || !is.character(env_time_type)) {
    env_time_type <- params$time_type
}

n_locations <- n_distinct(df_candidate$geo_value)
```

```{r}
#| label: quarto-metadata
#| results: asis
#| echo: false
# Dynamically set Quarto metadata so we can use it in ::: {.content-visible}
# blocks later.
cat(sprintf(
    "\n---\nmulti_location: %s\n---\n",
    tolower(n_locations > 1)
))
```

## Exploratory Data Analysis {#eda}

The data is aggregated at the `r params$geo_type` level. The dataset contains `r n_locations` valid locations measured at a `r env_time_type`-level frequency.

### Time Series Overview

This section provides a visual overview of the candidate indicator over time for a subset of observation locations. We look at the raw values for the indicator across a sample of locations.

```{r}
#| label: eda-overview
#| fig-height: 4
# Pick a diverse sample of locations for visualization
all_geos <- sort(unique(df_candidate$geo_value))
n_plot <- min(length(all_geos), params$max_locations_plot)
sample_geos <- random_sample(all_geos, n_plot)

df_sample <- df_candidate %>%
    filter(geo_value %in% sample_geos)

autoplot(df_sample, value,
        .interactive = TRUE, .facet_by = "geo_value", .facet_to_dropdown = TRUE) %>%
        plotly::toWebGL() %>%
        plotly::partial_bundle()
```

::: {.content-visible when-meta="multi_location"}

### Quantile Trends

This plot displays the distribution of both indicators over time across all included locations. The solid line represents the median, the inner ribbon shows the 25th to 75th percentile range, and the outer ribbon covers the 5th to 95th percentile range. We smooth the data with a moving average to handle day-of-week effects, and we use different scales for the top and bottom panels so each indicator's distribution is clearly visible.

By observing the width of these ribbons, you can get a sense of geographic variability. Narrow ribbons mean the values are tightly clustered across locations, while wide ribbons show more spread. This helps us assess the overall stability and systemic trends of the indicators.

```{r}
#| label: eda-quantiles
#| fig-height: !expr min(max(5, length(signals_vec)*1.5), 15)
#| eval: !expr n_locations > 1
cat("\n")

# Aggregate location values per (signal, time_value); use signal name as geo_value
# so epi_slide partitions by signal automatically.
df_quantiles_wide <- df_candidate %>%
    group_by(signal, time_value) %>%
    summarize(
        vals = list(value), 
        geo_value = "all",
        .groups = "drop"
    ) %>%
    as_epi_df(signal_format = "long")

# Handle time frequency differences for .window_size.
window_size <- switch(env_time_type %||% "unknown",
    day   = 7L, # 7-day smoothing
    week  = as.difftime(3, units = "weeks"), # 3-week smoothing
    month = as.difftime(3, units = "months"), # 3-month smoothing
    year  = as.difftime(3, units = "years"), # 3-year smoothing
    Inf # Default to cumulative for unknown types
    # (too sparse for sliding averages)
)

df_quantiles_wide <- df_quantiles_wide %>%
    group_by(geo_value, signal) %>%
    epi_slide(
        ~ tibble(
            `0.05` = quantile(unlist(.x$vals), 0.05, na.rm = TRUE),
            `0.25` = quantile(unlist(.x$vals), 0.25, na.rm = TRUE),
            `0.5`  = quantile(unlist(.x$vals), 0.5, na.rm = TRUE),
            `0.75` = quantile(unlist(.x$vals), 0.75, na.rm = TRUE),
            `0.95` = quantile(unlist(.x$vals), 0.95, na.rm = TRUE)
        ),
        .window_size = window_size
    ) %>%
    ungroup() %>%
    mutate(indicator = signal_labels[signal])

ggplot(df_quantiles_wide, aes(x = time_value)) +
    geom_ribbon(aes(ymin = `0.05`, ymax = `0.95`, fill = "5%-95%"), alpha = 0.2) +
    geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`, fill = "25%-75%"), alpha = 0.4) +
    geom_line(aes(y = `0.5`, color = "Median"), linewidth = 1) +
    facet_wrap(~indicator, ncol = 1, scales = "free_y") +
    scale_x_date(breaks = breaks_pretty(), label = label_date_short()) +
    scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_si(""))) +
    scale_fill_manual(values = c("5%-95%" = dis_pal_list[2], "25%-75%" = dis_pal_list[2])) +
    scale_color_manual(values = c("Median" = dis_pal_list[2])) +
    labs(
        title = "Smoothed Quantile Trends",
        subtitle = sprintf("Across all %d locations, smoothed window: %s", n_locations, format(window_size)),
        x = "Date", y = "Value (smoothed)",
        fill = "Range", color = "Trend"
    ) +
    theme(legend.position = "bottom")
```

#### Cross-Sectional Geographic Outliers

While the quantile plots summarize the main body of the distribution, they may hide the behavior of individual highly erratic locations. Here, we track how often individual locations jump into the extreme upper tail (above the cross-sectional 95th percentile recorded on that specific date).

::: {.callout-note}
This approach compares locations against one another on each date. If the indicator relies on raw counts rather than population-standardized rates or percentages, this table will frequently just identify locations with the largest populations. In those cases, this metric reflects size rather than outliers.
:::

```{r}
#| label: eda-aberrant-locations
#| echo: false
#| results: asis
#| eval: !expr n_locations > 1
# Calculate daily 95th percentiles per signal
daily_p95 <- df_candidate %>%
    group_by(signal, time_value) %>%
    summarize(
        cand_p95 = quantile(value, 0.95, na.rm = TRUE),
        .groups = "drop"
    )

# Find how often each location exceeds the 95th percentile, per signal
aberrant_locs <- df_candidate %>%
    left_join(daily_p95, by = c("signal", "time_value")) %>%
    group_by(signal, geo_value) %>%
    summarize(
        n_days = n(),
        cand_aberrant_days = sum(value > cand_p95, na.rm = TRUE),
        cand_aberrant_pct = (cand_aberrant_days / n_days) * 100,
        .groups = "drop"
    ) %>%
    mutate(
        geo_name     = geo_to_name(geo_value),
        signal_label = signal_labels[signal]
    ) %>%
    arrange(desc(cand_aberrant_pct))

if (nrow(aberrant_locs) > 0) {
    aberrant_display <- if (multi_signal) {
        aberrant_locs %>% select(signal_label, geo_value, geo_name, cand_aberrant_pct)
    } else {
        aberrant_locs %>% select(geo_value, geo_name, cand_aberrant_pct)
    }
    DT::datatable(
        aberrant_display,
        colnames = c(
            if (multi_signal) "Signal" else NULL,
            "Geo-level",
            "Name",
            "% Time Above 95th (candidate)"
        ),
        caption = "Summary of aberrant locations (time spent in the upper tail).",
        rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
    ) %>% DT::formatRound(columns = if (multi_signal) 4 else 3, digits = 1)
} else {
    cat("Insufficient data to calculate aberrant time for these locations.\n\n")
}
```





:::

::: {.content-visible when-meta="multi_location"}
### Coverage and Missingness

This section explores the availability of data across time and space. The area chart below tracks the number of geographic locations with valid data on any given date for the candidate indicator.

```{r}
#| label: eda-coverage
#| echo: false
#| results: asis
coverage <- df_candidate %>%
    group_by(signal, time_value) %>%
    summarize(n_available = sum(!is.na(value)), .groups = "drop") %>%
    mutate(geo_value = "all") %>%
    as_epi_df(signal_format = "long")

# Determine smoothing window based on time type
slide_window <- switch(env_time_type %||% "unknown",
    day = 7L,
    week = 4L,
    month = 3L,
    1L
)

if (env_time_type != params$time_type &&slide_window > 1) {
    coverage <- coverage %>%
        group_by(geo_value, signal) %>%
        epi_slide(
            n_available_smooth = mean(n_available, na.rm = TRUE),
            .window_size = slide_window
        ) %>%
        ungroup() %>%
        mutate(n_available = n_available_smooth)
}

coverage <- coverage %>%
    mutate(indicator = signal_labels[signal])
```

```{r}
#| label: eda-coverage-plot
#| fig-height: !expr min(max(5, length(signals_vec)*2), 12)
#| echo: true
ggplot(coverage, aes(x = time_value, y = n_available,
                     color = indicator, linetype = indicator)) +
    geom_line(linewidth = 0.8) +
    scale_x_date(breaks = breaks_pretty(), label = label_date_short()) +
    scale_y_continuous(labels = label_comma(), limits = c(0, NA)) +
    labs(
        title = "Data coverage over time",
        subtitle = sprintf("Number of locations with non-missing values per %s%s", env_time_type,
                           if (env_time_type != params$time_type && slide_window > 1) sprintf(" (%d-%s rolling average)", slide_window, env_time_type) else ""),
        x = "Date", y = "Number of locations",
        color = "Indicator", linetype = "Indicator"
    ) +
    {if (length(signals_vec) > 6) facet_wrap(~indicator, ncol = 2, labeller = label_wrap_html(width = 25)) else NULL}
```


:::


#### Summary Statistics by Location

This table summarizes the available data across all included locations. Locations with no candidate information are not considered.

```{r}
#| label: eda-summary
#| results: asis
#| echo: false
# Compute base stats per (signal, geo_value)
summary_tbl_all <- df_candidate %>%
    complete(signal, geo_value, time_value) %>%
    group_by(signal, geo_value) %>%
    summarize(
        n_days = n(),
        cand_mean = mean(value, na.rm = TRUE),
        cand_var = var(value, na.rm = TRUE),
        cand_miss_frac = mean(is.na(value)),
        cand_miss = sprintf("%.1f%%", 100 * cand_miss_frac),
        .groups = "drop"
    ) %>%
    mutate(
        geo_name     = geo_to_name(geo_value),
        signal_label = signal_labels[signal]
    )

excluded_locs_df <- summary_tbl_all %>%
    filter(is.na(cand_mean) | cand_miss_frac == 1)
if (nrow(excluded_locs_df) > 0) {
    for (sig in signals_vec) {
        ex_sig <- excluded_locs_df %>% filter(signal == sig)
        if (nrow(ex_sig) > 0) {
            cat(sprintf(
                "**Excluded locations (no data) — %s**: %s\n\n",
                signal_labels[sig],
                format_excluded(ex_sig$geo_value, ex_sig$geo_name)
            ))
        }
    }
}

zero_var_cand_locs_df <- summary_tbl_all %>%
    filter(!is.na(cand_var) & cand_var == 0 & cand_miss_frac < 1)
if (nrow(zero_var_cand_locs_df) > 0) {
    for (sig in signals_vec) {
        zv_sig <- zero_var_cand_locs_df %>% filter(signal == sig)
        if (nrow(zv_sig) > 0) {
            cat(sprintf(
                "**Zero variance locations (flatlined) — %s**: %s\n\n",
                signal_labels[sig],
                format_excluded(zv_sig$geo_value, zv_sig$geo_name)
            ))
        }
    }
}

# Summary of missingness levels per signal
miss_levels <- summary_tbl_all %>%
    mutate(
        cand_miss_group = case_when(
            cand_miss_frac == 1 ~ "100% missing",
            cand_miss_frac < 0.05 ~ "<5% missing",
            cand_miss_frac < 0.25 ~ "5-25% missing",
            cand_miss_frac < 0.50 ~ "25-50% missing",
            TRUE ~ "50-99.9% missing"
        ),
        cand_miss_group = factor(cand_miss_group, levels = c("<5% missing", "5-25% missing", "25-50% missing", "50-99.9% missing", "100% missing"))
    )

for (sig in signals_vec) {
    cat(sprintf("**Missingness Overview%s:**\n\n",
                if (multi_signal) sprintf(" — %s", signal_labels[sig]) else " (All Locations)"))
    cand_counts <- miss_levels %>%
        filter(signal == sig) %>%
        count(cand_miss_group) %>%
        complete(cand_miss_group, fill = list(n = 0))
    for (i in seq_len(nrow(cand_counts))) {
        cat(sprintf("- **%s**: %d locations\n",
                    cand_counts$cand_miss_group[i], cand_counts$n[i]))
    }
    cat("\n")
}

summary_tbl <- summary_tbl_all %>%
    filter(!is.na(cand_mean) & cand_miss_frac < 1) %>%
    arrange(desc(cand_miss_frac)) %>%
    {
        if (multi_signal) select(., signal_label, geo_value, geo_name, n_days, cand_mean, cand_miss)
        else select(., geo_value, geo_name, n_days, cand_mean, cand_miss)
    }

DT::datatable(
    summary_tbl,
    colnames = c(
        if (multi_signal) "Signal" else NULL,
        sprintf("Geo-level (%s)", params$geo_type),
        "Name",
        sprintf("Total %s(s)", str_to_title(env_time_type)),
        "Avg. Value",
        "% Missing"
    ),
    caption = sprintf(
        "Summary statistics by location. Missingness is calculated against the full analysis window (%d periods).",
        n_distinct(df_candidate$time_value)
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = if (multi_signal) 5 else 4, digits = 2)
```





::: {.content-visible when-meta="multi_location"}

### Noise Assessment

We calculate the Coefficient of Variation (CV) for each location to compare indicator stability across geographic areas.

$$\text{CV}_i = \frac{\text{SD}(X_i)}{|\bar{X}_i|}$$

```{r}
#| label: eda-noise
#| fig-height: !expr min(max(5, length(signals_vec)*0.5), 12)
#| eval: !expr n_locations > 1
# Calculate CV per (signal, location)
noise_stats <- df_candidate %>%
    arrange(signal, geo_value, time_value) %>%
    group_by(signal, geo_value) %>%
    summarize(
        cand_cv = sd(value, na.rm = TRUE) / abs(mean(value, na.rm = TRUE)),
        .groups = "drop"
    ) %>%
    filter(is.finite(cand_cv))

# Clip extreme outliers for robust visualization
cv_upper <- quantile(noise_stats$cand_cv, 0.99, na.rm = TRUE)
noise_stats_plot <- noise_stats %>%
    filter(cand_cv <= cv_upper) %>%
    mutate(signal_label = signal_labels[signal])

n_excluded <- nrow(noise_stats) - nrow(noise_stats_plot)

# Summary Table — one row per signal
noise_summary <- noise_stats %>%
    group_by(signal) %>%
    summarize(`Median CV` = median(cand_cv), .groups = "drop") %>%
    mutate(Signal = signal_labels[signal]) %>%
    select(Signal, `Median CV`)

DT::datatable(noise_summary,
    caption = sprintf(
        "CV summary across locations.%s",
        if (n_excluded > 0) sprintf(" (%d extreme outlier(s) excluded from plot.)", n_excluded) else ""
    ),
    rownames = FALSE, options = list(scrollX = TRUE, paging = FALSE, dom = "t")
) %>% DT::formatRound(columns = 2, digits = 3)

# Histogram (faceted by signal when multi_signal)
ggplot(noise_stats_plot, aes(x = cand_cv)) +
    geom_histogram(bins = 30, color = "white", fill = dis_pal_list[2], alpha = 0.7) +
    scale_x_continuous(labels = label_percent()) +
    {if (multi_signal) facet_wrap(~signal_label, scales = "free", labeller = label_wrap_html(width = 25)) else NULL} +
    labs(
        title = "CV Distribution by Location",
        subtitle = "Lower values indicate a more stable indicator",
        x = "Coefficient of Variation (CV)", y = "Count"
    ) +
    theme(legend.position = "none")
```





:::

```{r}
#| label: revision-analysis-setup
#| echo: false
#| dependson: "load-candidate-archive"
# Get versioned data
archive <- NULL
archive_has_history <- FALSE
if (!is.null(candidate_archive) &&
    n_distinct(candidate_archive$DT$version) > 1) {
    archive <- candidate_archive
    archive_span_days <- as.numeric(
        max(archive$DT$version) - min(archive$DT$version),
        units = "days"
    )
    archive_has_history <- archive_span_days >= 60
}
```





### Candidate Versioning

This section examines how published values change over time as new data arrives. Each observation may be revised multiple times after its initial release, and the statistics below summarize both how often and how much those revisions occur. Relative revisions measure the magnitude of a revision relative to the signal's own scale. A relative revision of 0.1 means the observation shifted by 10% of its own magnitude across revisions, while values near 1 indicate revisions as large as the signal itself.

```{r}
#| label: revision-analysis-text
#| echo: false
#| results: asis
if (is.null(archive)) {
    cat(sprintf(
        "No version history is available for `%s / %s` — versioning analysis is skipped.\n\n",
        params$source, paste(signals_vec, collapse = ", ")
    ))
} else if (!archive_has_history) {
    version_start <- min(archive$DT$version)
    cat(sprintf(
        "Versioning began on %s. At least 60 days of version history are needed for meaningful revision analysis; the earliest this will be possible is %s.\n\n",
        format(version_start), format(version_start + 60)
    ))
    cat(sprintf(
        "Current versions: %s.\n\n",
        paste(sort(unique(archive$DT$version)), collapse = ", ")
    ))
} else {
    cat("> The `revision_summary` is under revision and may produce misleading results in certain edge cases. Use the summary statistics below with caution while development is ongoing.\n\n")

    if (exists("archive_info") && !is.null(archive_info)) {
        cat(sprintf(
            "Revision analysis is performed on a sample of %d out of %d locations, with versions from %s to %s, and time from %s to %s.\n\n",
            archive_info$n_sampled, archive_info$n_total, archive_info$version_start, archive_info$version_end, archive_info$time_value_start, archive_info$time_value_end
        ))
    }

    # Helper: print revision stats block for one signal's rev_details
    print_rev_stats <- function(rd, archive_obj) {
        rb <- rd$revision_behavior
        cat(sprintf(
            "- **Time range**: %s to %s\n",
            rd$range_time_values[1], rd$range_time_values[2]
        ))
        lag_stats <- summary(as.numeric(rb$min_lag, units = "days"))
        cat("- **Min lag (days to first version)**: ")
        cat(paste(names(lag_stats), round(as.numeric(lag_stats), 1), sep = ": ", collapse = ", "), "\n")
        quick_threshold <- 3
        few_threshold <- 3
        rel_spread_threshold <- 0.1
        abs_spread_threshold <- 0.05 * rd$max_val
        no_rev_pct   <- mean(rb$n_revisions == 0)
        quick_rev_pct <- mean(as.numeric(rb$max_lag, units = "days") <= quick_threshold)
        few_rev_pct  <- mean(rb$n_revisions <= few_threshold)
        cat(sprintf("- **Total number of versions**: %d\n", n_distinct(archive_obj$DT$version)))
        cat(sprintf("- **No revisions**: %s\n", label_percent()(no_rev_pct)))
        cat(sprintf("- **Quick revisions (<= %d days)**: %s\n", quick_threshold, label_percent()(quick_rev_pct)))
        cat(sprintf("- **Few revisions (<= %d)**: %s\n", few_threshold, label_percent()(few_rev_pct)))
        revised_beh <- rb %>% filter(n_revisions > 0)
        if (nrow(revised_beh) > 0) {
            low_rel_spread_pct  <- mean(revised_beh$rel_spread < rel_spread_threshold, na.rm = TRUE)
            high_abs_spread_pct <- mean(revised_beh$spread > abs_spread_threshold, na.rm = TRUE)
            cat(sprintf("- **Minor relative revisions (< %.1f)**: %s\n",
                        rel_spread_threshold, label_percent()(low_rel_spread_pct)))
            cat(sprintf("- **Major absolute revisions (> %.3f)**: %s\n",
                        abs_spread_threshold, label_percent()(high_abs_spread_pct)))
        }
        cat(sprintf("- **Days until within %d%% of latest**: ", rd$within_latest * 100))
        conv_stats <- summary(as.numeric(rb$lag_near_latest, units = "days"))
        cat(paste(names(conv_stats), round(as.numeric(conv_stats), 1), sep = ": ", collapse = ", "), "\n\n")
    }

    # Run revision_summary per signal so results are always signal-specific
    rev_details_list <- list()
    rev_beh_combined <- NULL

    for (sig in signals_vec) {
        if (multi_signal) cat(sprintf("\n#### %s\n\n", signal_labels[sig]))

        # Subset archive to this signal (drop signal col for revision_summary)
        sub_archive <- archive$DT %>%
            as_tibble() %>%
            filter(signal == sig) %>%
            select(geo_value, time_value, version, value) %>%
            as_epi_archive()

        rd_sig <- revision_summary(sub_archive)
        rev_details_list[[sig]] <- rd_sig

        # Accumulate combined revision_behavior with signal label
        rev_beh_combined <- bind_rows(
            rev_beh_combined,
            rd_sig$revision_behavior %>% mutate(signal = sig)
        )

        print_rev_stats(rd_sig, sub_archive)
    }

    # Expose rev_details (first signal) for the visuals chunk; inject combined rev_behavior
    rev_details <- rev_details_list[[signals_vec[1]]]
    rev_details$revision_behavior <- rev_beh_combined
}
```

```{r}
#| label: revision-analysis-visuals-text
#| echo: false
#| results: asis
#| eval: !expr (archive_has_history && n_locations > 1)
cat(sprintf("The time series below visualize the candidate indicator(s) (%s) with line colors corresponding to the issue date, meaning the date the data was initially published or subsequently revised. Darker colors denote older versions, while lighter colors represent more recent updates.\n\n", paste(name_vec, collapse = ", ")))
cat("If the lines perfectly overlap, it means the data for that location is not subject to revisions. Where you see a vertical spread along a single date on the x-axis, the data has been revised over time.\n\n")
```

```{r}
#| label: revision-analysis-visuals
#| eval: !expr archive_has_history
#| results: asis
#| fig-height: !expr min(max(5, length(signals_vec)*1.5), 12)
rev_by_loc <- rev_details$revision_behavior %>%
    filter(n_revisions != 0) %>%
    group_by(signal, geo_value) %>%
    summarize(
        n_versions = n(),
        n_rev = mean(n_revisions, na.rm = TRUE),
        min_lag = min(min_lag, na.rm = TRUE),
        max_lag = max(max_lag, na.rm = TRUE),
        spread = mean(spread, na.rm = TRUE),
        rel_spread = mean(rel_spread, na.rm = TRUE),
        lag_near_latest = mean(lag_near_latest, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(signal_label = signal_labels[signal]) %>%
    arrange(desc(rel_spread))

if (nrow(rev_by_loc) > 0) {
    cat("The table below summarizes revision behavior per location. Column definitions:\n\n")
    cat("- Min Lag. Minimum days between the event date and its first publication.\n")
    cat("- Max Lag. Maximum days between the event date and its final revision.\n")
    cat("- Avg. Spread. Average absolute difference between the minimum and maximum published values.\n")
    cat("- Avg. Rel. Spread. Average relative difference between the minimum and maximum published values (spread divided by the maximum value).\n")
    cat(sprintf("- Lag Near Latest. Average days until the value stays within %d%% of the latest published value.\n\n", rev_details$within_latest * 100))

    rev_by_loc_display <- if (multi_signal) {
        rev_by_loc %>% select(signal_label, geo_value, n_versions:lag_near_latest)
    } else {
        rev_by_loc %>% select(geo_value, n_versions:lag_near_latest)
    }
    DT::datatable(rev_by_loc_display,
        colnames = c(
            if (multi_signal) "Signal" else NULL,
            "Location", "Versions", "Avg. Revisions", "Min Lag", "Max Lag",
            "Avg. Spread", "Avg. Rel. Spread", "Lag Near Latest"
        ),
        caption = "Candidate revision behavior by location",
        rownames = FALSE, options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
    ) %>% DT::formatRound(columns = if (multi_signal) 3:9 else 2:8, digits = 2)
}

if (nrow(rev_by_loc) > 0 && n_locations > 1) {
    # Pick top locations by max rel_spread across signals
    rev_geos <- rev_by_loc %>%
        group_by(geo_value) %>%
        summarize(rel_spread = max(rel_spread, na.rm = TRUE), .groups = "drop") %>%
        arrange(desc(rel_spread)) %>%
        #sample_geos
        head(params$max_locations_plot) %>%
        pull(geo_value)

    # Limit to last year for clarity
    # For multi-signal, loop per signal to avoid autoplot dropdown issues with
    # the signal other-key (same pattern as eda-overview).
    archive_filtered <- archive %>%
        filter(geo_value %in% rev_geos) %>%
        filter(time_value >= max(time_value) - years(1))

    if (!multi_signal) {
        n_geos_rev <- n_distinct(archive_filtered$DT$geo_value)
        autoplot(archive_filtered, .versions = "month",
            .interactive = TRUE, .facet_to_dropdown = n_geos_rev > 1) %>%
            plotly::toWebGL() %>%
            plotly::partial_bundle()
    } else {
        rev_plots <- purrr::map(signals_vec, function(sig) {
            arch_sig_dt <- archive_filtered$DT %>%
                as_tibble() %>%
                filter(signal == sig) %>%
                select(-signal)
            if (nrow(arch_sig_dt) == 0) return(NULL)
            arch_sig <- as_epi_archive(arch_sig_dt)
            n_geos_sig <- n_distinct(arch_sig$DT$geo_value)
            p <- autoplot(arch_sig, .versions = "month",
                .interactive = TRUE, .facet_to_dropdown = n_geos_sig > 1) %>%
                plotly::toWebGL() %>%
                plotly::partial_bundle()
            htmltools::tagList(
                htmltools::tags$h5(signal_labels[sig]),
                p
            )
        }) %>%
            purrr::compact()
        htmltools::tagList(rev_plots)
    }
}
```





### Signal sanity checks

This section is by its nature more difficult than others to do systematically. Things to look for:

- If it's a percentile, is it between 0 and 100?
- If it's a rate, is it roughly proportional to population? *(requires external population data — inspect manually)*
- Are the actual values reasonable? E.g. if it's a percentage, does it ever approach 100 outside of literal pandemics?
- If it's for a seasonal disease, are rates highest in the winter and lowest in the summer?
- Are geo-aggregated coarser levels plausible aggregations? *(checked below for state-level indicators vs HHS/nation)*

The histogram and table below show the distribution of all observed values and the value range per location for the candidate indicator, to assess whether the range is plausible given the indicator's expected scale.

```{r}
#| label: sanity-value-range
#| fig-height: !expr min(max(5, length(signals_vec)*0.5), 12)
df_candidate_plot <- df_candidate %>%
    mutate(indicator = signal_labels[signal])

ggplot(df_candidate_plot, aes(x = value)) +
    geom_histogram(bins = 40, alpha = 0.8, color = "white") +
    scale_x_continuous(labels = label_number(scale_cut = cut_si(""))) +
    scale_y_continuous(labels = label_number(scale_cut = cut_si(""))) +
    {if (multi_signal) facet_wrap(~indicator, scales = "free", labeller = label_wrap_html(width = 25)) else NULL} +
    labs(
        title = if (multi_signal) "Value distribution by signal"
                else sprintf("Value distribution: %s", signal_labels[signals_vec]),
        subtitle = sprintf(
            "%s observations across %d locations",
            label_comma()(nrow(df_candidate)), n_locations
        ),
        x = "Value", y = "Count"
    )
```

```{r}
#| label: sanity-value-range-table
#| echo: false
range_tbl <- df_candidate %>%
    group_by(signal, geo_value) %>%
    summarize(
        n = n(),
        min = min(value, na.rm = TRUE),
        p05 = quantile(value, 0.05, na.rm = TRUE),
        median = median(value, na.rm = TRUE),
        p95 = quantile(value, 0.95, na.rm = TRUE),
        max = max(value, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(signal_label = signal_labels[signal]) %>%
    arrange(desc(max)) %>%
    {
        if (multi_signal) select(., signal_label, geo_value, n:max)
        else select(., geo_value, n:max)
    }

DT::datatable(
    range_tbl,
    colnames = c(
        if (multi_signal) "Signal" else NULL,
        "Location", "N", "Min", "5th pct", "Median", "95th pct", "Max"
    ),
    caption = sprintf(
        "Value range per location%s. Sort by Min or Max to surface potential out-of-bounds values.",
        if (multi_signal) "" else sprintf(" for %s", signal_labels[signals_vec])
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, dom = "tp"
    )
) %>% DT::formatRound(columns = if (multi_signal) 4:8 else 3:7, digits = 3)
```

The plot below shows the seasonal pattern of the indicator, with each panel representing one epidemiological year (June–May). The line traces the monthly median across all locations, the inner band covers the interquartile range, and the outer band spans the full observed range.

```{r}
#| label: sanity-seasonality
#| fig-height: !expr min(max(5, length(signals_vec)*2.5), 15)
#| eval: !expr as.numeric(difftime(max(df_candidate$time_value), min(df_candidate$time_value), units = "days")) >= 365

season_month_labels <- function(x) {
c("Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May")[x]
}

df_seasonal <- df_candidate %>%
    mutate(
        yr = as.integer(format(time_value, "%Y")),
        month = as.integer(format(time_value, "%m")),
        season_start = if_else(month >= 6L, yr, yr - 1L),
        season = paste0(season_start, "-", substr(as.character(season_start + 1L), 3L, 4L)),
        season_month = if_else(month >= 6L, month - 5L, month + 7L)
    ) %>%
    group_by(signal, season, season_month) %>%
    summarize(
        median_val = median(value, na.rm = TRUE),
        min_val = min(value, na.rm = TRUE),
        max_val = max(value, na.rm = TRUE),
        p25 = quantile(value, 0.25, na.rm = TRUE),
        p75 = quantile(value, 0.75, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    # Keep only signal-season combos with >= 3 months of data
    semi_join(
        count(., signal, season) %>% filter(n >= 3),
        by = c("signal", "season")
    ) %>%
    mutate(signal_label = signal_labels[signal])

ggplot(df_seasonal, aes(x = season_month, color = season, fill = season, group = season)) +
    geom_ribbon(aes(ymin = min_val, ymax = max_val), alpha = 0.10, linewidth = 0.2) +
    geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.30, linewidth = 0.3) +
    geom_line(aes(y = median_val), linewidth = 0.8) +
    geom_point(aes(y = median_val), size = 1.5) +
    {
        if (multi_signal) facet_grid(signal_label ~ season, scales = "free_y")
        else facet_wrap(~season, ncol = 2, scales = "free_y")
    } +
    scale_x_continuous(labels = season_month_labels) +
    scale_y_continuous(labels = label_number(scale_cut = cut_si(""))) +
    labs(
        title = if (multi_signal) "Seasonal patterns by signal"
                else sprintf("Seasonal pattern: %s", signal_labels[signals_vec]),
        subtitle = "Median by month across all locations — inner band: IQR, outer band: full range",
        x = "Month", y = "Median value"
    ) +
    theme(legend.position = "none")
```

```{r}
#| label: sanity-geo-aggregation
#| echo: false
#| results: asis
#| eval: !expr params$geo_type == "state"
pacman::p_load(nanoparquet)

crosswalk_path <- here("tests", "test_data", "delphi_geomap", "geomapper_crosswalk_20200101.parquet")
crosswalk <- tryCatch(
    {
        nanoparquet::read_parquet(crosswalk_path) %>%
            distinct(state_geo = state_abbr, hhs_region = as.character(hhs))
    },
    error = function(e) {
        # Fallback to standard HHS region mapping if the file is missing
        tibble(
            state_geo = c(
                "CT", "ME", "MA", "NH", "RI", "VT",
                "NJ", "NY", "PR", "VI",
                "DE", "DC", "MD", "PA", "VA", "WV",
                "AL", "FL", "GA", "KY", "MS", "NC", "SC", "TN",
                "IL", "IN", "MI", "MN", "OH", "WI",
                "AR", "LA", "NM", "OK", "TX",
                "IA", "KS", "MO", "NE",
                "CO", "MT", "ND", "SD", "UT", "WY",
                "AZ", "CA", "HI", "NV", "AS", "GU", "MP",
                "AK", "ID", "OR", "WA"
            ),
            hhs_region = c(
                rep("1", 6),
                rep("2", 4),
                rep("3", 6),
                rep("4", 8),
                rep("5", 6),
                rep("6", 5),
                rep("7", 4),
                rep("8", 6),
                rep("9", 7),
                rep("10", 4)
            )
        )
    }
)

is_pct_indicator <- any(grepl("%|pct", name_vec, ignore.case = TRUE))

fetch_coarser_candidate <- function(geo_type_coarser) {
    tryCatch(
        # pub_covidcast accepts a vector of signals
        pub_covidcast(
            params$source, signals_vec,
            geo_type_coarser, params$time_type,
            time_values = time_range(params$start_day, params$end_day),
            as_of = Sys.Date()
        ) %>% select(time_value, signal, geo_value, coarser_val = value),
        error = function(e) NULL
    )
}

state_vals <- df_candidate %>%
    as_tibble() %>%
    select(signal, time_value, state_geo = geo_value, state_val = value)

check_aggregation <- function(df_coarser, group_label, state_to_coarser) {
    if (is.null(df_coarser)) {
        return(NULL)
    }

    state_bounds <- state_vals %>%
        inner_join(state_to_coarser, by = "state_geo") %>%
        group_by(signal, time_value, coarser_geo) %>%
        summarize(
            state_min = min(state_val, na.rm = TRUE),
            state_max = max(state_val, na.rm = TRUE),
            n_states  = n(),
            .groups   = "drop"
        )

    df_coarser %>%
        rename(coarser_geo = geo_value) %>%
        inner_join(state_bounds, by = c("signal", "time_value", "coarser_geo")) %>%
        filter(!is.na(coarser_val) & (coarser_val < state_min - 1e-9 | coarser_val > state_max + 1e-9)) %>%
        mutate(
            geo_level    = group_label,
            signal_label = signal_labels[signal],
            direction    = if_else(coarser_val < state_min, "below min", "above max")
        ) %>%
        select(geo_level, signal_label, time_value, coarser_geo, coarser_val, state_min, state_max, direction, n_states)
}

df_hhs_raw <- fetch_coarser_candidate("hhs")
df_nation_raw <- fetch_coarser_candidate("nation")

hhs_map <- crosswalk %>% select(state_geo, coarser_geo = hhs_region)
nation_map <- crosswalk %>%
    distinct(state_geo) %>%
    mutate(coarser_geo = "us")

violations <- bind_rows(
    check_aggregation(df_hhs_raw, "HHS", hhs_map),
    check_aggregation(df_nation_raw, "Nation", nation_map)
)

unavailable <- c(
    if (is.null(df_hhs_raw)) "HHS" else NULL,
    if (is.null(df_nation_raw)) "Nation" else NULL
)

if (length(unavailable) > 0) {
    cat(sprintf(
        "**Note**: %s-level data not available for %s / %s.\n\n",
        paste(unavailable, collapse = " and "),
        params$source, paste(signals_vec, collapse = ", ")
    ))
}

if (nrow(violations) == 0) {
    cat(sprintf(
        "No geo-aggregation violations found: all HHS/nation values are within the [min, max] of constituent state values for %s.\n",
        paste(name_vec, collapse = ", ")
    ))
} else {
    cat(sprintf(
        "**%d observations** where the aggregated value falls outside the [min, max] of its constituent states (%s indicator):\n\n",
        nrow(violations),
        if (is_pct_indicator) "percentage" else "non-percentage"
    ))
    print(htmltools::tagList(
        DT::datatable(
            violations,
            colnames = c("Level", "Signal", "Date", "Geo", "Agg Value", "State Min", "State Max", "Direction", "N States"),
            caption = sprintf(
                "Geo-aggregation bounds check for %s. Aggregated values should fall within [min, max] of constituent state values.",
                paste(name_vec, collapse = ", ")
            ),
            rownames = FALSE,
            options = list(
                scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
                paging = TRUE, pageLength = 20, dom = "tp"
            )
        ) %>% DT::formatRound(columns = 5:7, digits = 3)
    ))
}
```





The table below flags locations whose seasonal peak is at least twice the typical (median) peak across all locations in the same season.

```{r}
#| label: sanity-seasonality-anomalies
#| echo: false
#| eval: !expr as.numeric(difftime(max(df_candidate$time_value), min(df_candidate$time_value), units = "days")) >= 365
df_seasonal_loc <- df_candidate %>%
    mutate(
        yr = as.integer(format(time_value, "%Y")),
        month = as.integer(format(time_value, "%m")),
        season_start = if_else(month >= 6L, yr, yr - 1L),
        season = paste0(season_start, "-", substr(as.character(season_start + 1L), 3L, 4L))
    ) %>%
    group_by(signal, season, geo_value) %>%
    summarize(loc_max = max(value, na.rm = TRUE), .groups = "drop")

typical_max <- df_seasonal_loc %>%
    group_by(signal, season) %>%
    summarize(typical_max = median(loc_max, na.rm = TRUE), .groups = "drop")

anomalies <- df_seasonal_loc %>%
    left_join(typical_max, by = c("signal", "season")) %>%
    mutate(
        ratio        = loc_max / typical_max,
        signal_label = signal_labels[signal]
    ) %>%
    filter(ratio >= 2) %>%
    arrange(desc(ratio)) %>%
    {
        if (multi_signal) select(., signal_label, season, geo_value, loc_max, typical_max, ratio)
        else select(., season, geo_value, loc_max, typical_max, ratio)
    }

if (nrow(anomalies) == 0) {
    cat("No locations found with a seasonal max more than 2x the typical (median) seasonal max.")
} else {
    DT::datatable(
        anomalies,
        colnames = c(
            if (multi_signal) "Signal" else NULL,
            "Season", "Location", "Location Max", "Typical Max", "Ratio"
        ),
        caption = sprintf(
            "Locations where seasonal max exceeds 2x the typical (median) seasonal max%s.",
            if (multi_signal) "" else sprintf(" for %s", signal_labels[signals_vec])
        ),
        rownames = FALSE,
        options = list(
            scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
            paging = TRUE, pageLength = 20, dom = "tp"
        )
    ) %>% DT::formatRound(columns = if (multi_signal) 4:6 else 3:5, digits = 3)
}
```